### 2021-07-25 notebooks/DataCleaning.Rmd
started datacleaning file

---
title: "Lab2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r import libraries}
library(tidyverse)
library(ggplot2) 
library(sandwich)
library(stargazer)

```

```{r read in data}

d_covid <- read.csv(file = './JAL_Lab2/src/data/covid-counties.csv')
d_openings <- read.csv(file = './JAL_Lab2/src/data/COVID-19 US state policy database (CUSP) - Closures & Reopening.csv')
d_trips <- read.csv(file = './JAL_Lab2/src/data/Trips_by_Distance_-_2020.csv')
```

### 2021-07-26 notebooks/DataCleaning.Rmd

@@
 library(stargazer)
+library(plyr)
+library(dplyr)
 
@@
 
-d_covid <- read.csv(file = './JAL_Lab2/src/data/covid-counties.csv')
-d_openings <- read.csv(file = './JAL_Lab2/src/data/COVID-19 US state policy database (CUSP) - Closures & Reopening.csv')
-d_trips <- read.csv(file = './JAL_Lab2/src/data/Trips_by_Distance_-_2020.csv')
+d_covid <- read.csv(file = './../src/data/covid-counties.csv')
+d_openings <- read.csv(file = './../src/data/COVID-19 US state policy database (CUSP) - Closures & Reopening.csv')
+d_trips <- read.csv(file = './../src/data/Trips_by_Distance_-_2020.csv')
 ```
+
+```{r filter data}
+#filter by date, month of april, 2020
+d_covid <- filter(d_covid, date >= "2020-04-01", date <= "2020-04-30")
+d_trips <- filter(d_trips, Date >= "4/1/2020", Date <= "4/30/2020")
+
+#sum values for each county
+covid_col <- ddply(d_covid, "county", numcolwise(sum))
+trips_col <- ddply(d_trips, "County.Name", numcolwise(sum))
+
+#clean county name format
+d_trips$county <- gsub("([A-Za-z]+).*", "\\1", d_trips$County.Name)
+trips_col$county <- gsub("([A-Za-z]+).*", "\\1", trips_col$County.Name)
+
+#code to merge state info to d
+d <- merge(x = trips_col, y = covid_col, by = "county")
+d <- merge(x = d, y = d_trips[ , c("county", "State.Postal.Code")], by = "county", all.x=TRUE)
+
+#merge county policies to d
+d_openings$strict <- d_openings$Closed.bars.x2 != "0"
+
+d_openings$State.Postal.Code <- d_openings$State.Abbreviation
+d <- merge(x = d, y = d_openings[ , c("State.Postal.Code", "strict")], by = "State.Postal.Code", all.x = TRUE)
+
+#drop na and duplicates
+d <- d %>% drop_na()
+d <- unique(d)
+```

### 2021-07-27 notebooks/lab2_report.Rmd
Initial report commit.

---
title: 'Title!'
author: "Autumn Rains, Jonas Degnan, Lucy Wu"
date: "8/5/2021"
output:
  bookdown::pdf_document2: 
    toc: true
---

\clearpage

```{r install packages, echo=FALSE, warning=FALSE, message=FALSE}
# install.packages('tinytex')
# tinytex::install_tinytex()
```

```{r load packages, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse) 
library(magrittr)
library(knitr)
library(patchwork)
library(kableExtra)
library(tinytex)

theme_set(theme_bw())

options(tinytex.verbose = TRUE)
```

```{r load clean data, echo=FALSE, message=FALSE}

```

### 2021-07-29 src/data/DataCleaning.Rmd
updated data source paths. made covid cleaning more functional to all for changes in investigation targets.

---
title: "Lab2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r import libraries, echo=FALSE, warning=FALSE}
library(tidyverse)
library(anytime)
# source(local=TRUE, file="../util.R")
```

```{r read in data, echo=FALSE}

d_covid <- read.csv(file = './../../data/raw/covid-counties.csv')
d_openings <- read.csv(file = './../../data/raw/COVID-19 US state policy database (CUSP) - Closures & Reopening.csv')
d_trips <- read.csv(file = './../../data/raw/Trips_by_Distance_-_2020.csv')

```

```{r clean covid data, echo=FALSE}
# state(s) under investigation
state_id = c("OH","Ohio",39)

# dates under investigation
start_date = "2020-04-01"
end_date = "2020-04-30"

# COVID-19 data
# filters by date/state
d_covid <- d_covid %>%
  filter(d_covid, 
         date >= start_date, date <= end_date,
         grepl(paste(state_id, collapse = "|"), state)
         )

# aggregate county data
d_covid <- d_covid %>%
  group_by(fips) %>%
  mutate(
    sum_cases = sum(cases),
    sum_deaths = sum(deaths),
    state = state.abb[match(state, state.name)]
  ) %>%
  select(
    county, state, fips, sum_cases, sum_deaths
  ) %>%
  distinct(fips, .keep_all = TRUE)
```

```{r clean mobility data, echo=FALSE}
# clean mobility data

# filters by date/state
d_trips <- d_trips %>%
  filter(d_trips, 
         Date >= "4/1/2020", Date <= "4/2/2020",
         grepl(paste(state_id, collapse = "|"), State.Postal.Code)
         )

# cleaning
d_trips_filter <- d_trips %>%

  mutate(County.Name = gsub("([A-Za-z]+).*", "\\1", County.Name)) %>%
  group_by(County.Name) %>%
  rowwise() %>%
  mutate(
    # clean county name format
    County.Name = gsub("([A-Za-z]+).*", "\\1", County.Name),
    # total pop
    total_pop = sum(c_across(Population.Staying.at.Home:Population.Not.Staying.at.Home)),
    # define long trips per capita
    long_trip = sum(c_across(Number.of.Trips.100.250:Number.of.Trips...500)),
    long_trips_pc = long_trip/total_pop,
    # # define short trips per capita
    short_trip = sum(c_across(Number.of.Trips.1.3:Number.of.Trips.50.100)),
    short_trips_pc = short_trip/total_pop
    # #
  ) %>%
  rename(
    state = State.Postal.Code,
    county = County.Name
  ) %>%
  select(
    county, state, total_pop, long_trip, long_trips_pc, short_trip, short_trips_pc
  ) %>%
  distinct(county, .keep_all = TRUE)

  # filter(State.Postal.Code %in% c(state[1]))

  # filter(State.Postal.Code %in% c("AL"), grepl("Aut",County.Name))

trips_col <- ddply(d_trips, "County.Name", numcolwise(sum))

#clean county name format
trips_col$county <- gsub("([A-Za-z]+).*", "\\1", trips_col$County.Name)

#code to merge state info to d
d <- merge(x = trips_col, y = covid_col, by = "county")
d <- merge(x = d, y = d_trips[ , c("county", "State.Postal.Code")], by = "county", all.x=TRUE)

#merge county policies to d
d_openings$strict <- d_openings$Closed.bars.x2 != "0"

d_openings$State.Postal.Code <- d_openings$State.Abbreviation
d <- merge(x = d, y = d_openings[ , c("State.Postal.Code", "strict")], by = "State.Postal.Code", all.x = TRUE)

#drop na and duplicates
d <- d %>% drop_na()
d <- unique(d)
```

### 2021-07-29 src/data/DataCleaning.Rmd
udpated covid-19 data source. fixed covid-19 filter.

@@
 
-```{r import libraries, echo=FALSE, warning=FALSE}
+```{r import libraries, echo=FALSE, message=FALSE, warning=FALSE}
 library(tidyverse)
@@
 ```{r read in data, echo=FALSE}
-
-d_covid <- read.csv(file = './../../data/raw/covid-counties.csv')
+d_covid <- read.csv(url('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties-recent.csv'))
 d_openings <- read.csv(file = './../../data/raw/COVID-19 US state policy database (CUSP) - Closures & Reopening.csv')
@@
 
-```{r clean covid data, echo=FALSE}
+```{r filter parameters, echo=FALSE}
 # state(s) under investigation
@@
 # dates under investigation
-start_date = "2020-04-01"
-end_date = "2020-04-30"
+start_date = as.Date("2021-06-01")
+end_date = as.Date("2021-06-30")
+```
 
+```{r clean covid data, echo=FALSE}
 # COVID-19 data
@@
 d_covid <- d_covid %>%
-  filter(d_covid, 
+  mutate(
+    date = as.Date(date)
+  ) %>%
+  filter(
     date >= start_date, date <= end_date,

### 2021-07-29 src/data/DataCleaning.Rmd
cleaned trip distance

@@
 d_openings <- read.csv(file = './../../data/raw/COVID-19 US state policy database (CUSP) - Closures & Reopening.csv')
-d_trips <- read.csv(file = './../../data/raw/Trips_by_Distance_-_2020.csv')
-
+d_trips <- read.csv(file = './../../data/raw/Trips_by_Distance.csv')
 ```
@@
 # state(s) under investigation
-state_id = c("OH","Ohio",39)
+state_id = c("TX","Texas")
 
@@
 d_trips <- d_trips %>%
-  filter(d_trips, 
-         Date >= "4/1/2020", Date <= "4/2/2020",
+  mutate(
+    Date = as.Date(Date)
+  ) %>%
+  filter(
+    Date >= start_date, Date <= end_date,
     grepl(paste(state_id, collapse = "|"), State.Postal.Code)
@@
 # cleaning
-d_trips_filter <- d_trips %>%
-
-  mutate(County.Name = gsub("([A-Za-z]+).*", "\\1", County.Name)) %>%
-  group_by(County.Name) %>%
-  rowwise() %>%
+d_trips <- d_trips %>%
+  # clean county name format
   mutate(
-    # clean county name format
     County.Name = gsub("([A-Za-z]+).*", "\\1", County.Name),
-    # total pop
-    total_pop = sum(c_across(Population.Staying.at.Home:Population.Not.Staying.at.Home)),
-    # define long trips per capita
+    total_pop = select(.,Population.Staying.at.Home:Population.Not.Staying.at.Home) %>% rowSums(na.rm = TRUE)
+    ) %>%
+  group_by(County.FIPS) %>%
+  mutate(
+    # define long trips
     long_trip = sum(c_across(Number.of.Trips.100.250:Number.of.Trips...500)),
-    long_trips_pc = long_trip/total_pop,
-    # # define short trips per capita
+    long_trip_pc = long_trip/total_pop,
+    # define short trips
     short_trip = sum(c_across(Number.of.Trips.1.3:Number.of.Trips.50.100)),
-    short_trips_pc = short_trip/total_pop
-    # #
+    short_trip_pc = short_trip/total_pop
   ) %>%
@@
   select(
-    county, state, total_pop, long_trip, long_trips_pc, short_trip, short_trips_pc
+    county, County.FIPS, state, total_pop, long_trip, long_trip_pc, short_trip, short_trip_pc
   ) %>%
-  distinct(county, .keep_all = TRUE)
+  distinct(County.FIPS, .keep_all = TRUE)
+```
 
-  # filter(State.Postal.Code %in% c(state[1]))
-
-  # filter(State.Postal.Code %in% c("AL"), grepl("Aut",County.Name))
-
-trips_col <- ddply(d_trips, "County.Name", numcolwise(sum))
-
-#clean county name format
-trips_col$county <- gsub("([A-Za-z]+).*", "\\1", trips_col$County.Name)
-
+```{r merged dataframe, echo=FALSE}
 #code to merge state info to d

### 2021-07-29 src/data/DataCleaning.Rmd
merged interim dataframe

@@
 # state(s) under investigation
-state_id = c("TX","Texas")
+state_id = c("TX","Texas",48)
 
@@
     state = State.Postal.Code,
-    county = County.Name
+    county = County.Name,
+    fips = County.FIPS
   ) %>%
   select(
-    county, County.FIPS, state, total_pop, long_trip, long_trip_pc, short_trip, short_trip_pc
+    county, fips, state, total_pop, long_trip, long_trip_pc, short_trip, short_trip_pc
   ) %>%
-  distinct(County.FIPS, .keep_all = TRUE)
+  distinct(fips, .keep_all = TRUE)
 ```
@@
 ```{r merged dataframe, echo=FALSE}
-#code to merge state info to d
-d <- merge(x = trips_col, y = covid_col, by = "county")
-d <- merge(x = d, y = d_trips[ , c("county", "State.Postal.Code")], by = "county", all.x=TRUE)
-
-#merge county policies to d
-d_openings$strict <- d_openings$Closed.bars.x2 != "0"
-
-d_openings$State.Postal.Code <- d_openings$State.Abbreviation
-d <- merge(x = d, y = d_openings[ , c("State.Postal.Code", "strict")], by = "State.Postal.Code", all.x = TRUE)
+#code to merge county info to d
+d <- merge(x = d_trips, y = d_covid, by = "fips")
 
@@
 d <- unique(d)
+
+#merge county policies to d
+# d_openings$strict <- d_openings$Closed.bars.x2 != "0"
+
+# d_openings$State.Postal.Code <- d_openings$State.Abbreviation
+# d <- merge(x = d, y = d_openings[ , c("State.Postal.Code", "strict")], by = "State.Postal.Code", all.x = TRUE)
 ```
+```{r export interim dataframe as csv, echo=FALSE}
+
+```

### 2021-07-29 src/data/DataCleaning.Rmd
generated covid_trips interim csv

@@
 ```{r export interim dataframe as csv, echo=FALSE}
-
+write.csv(d, '../../data/interim/covid_trips_interim.csv')
 ```

### 2021-08-01 src/data/DataCleaning.Rmd

@@
 library(tidyverse)
-library(anytime)
 # source(local=TRUE, file="../util.R")
@@
 #code to merge county info to d
-d <- merge(x = d_trips, y = d_covid, by = "fips")
+d <- merge(x = d_trips, y = d_covid, by = c("fips", "county", "state"))
 

### 2021-08-01 reports/Draft_Report.Rmd
Adding draft report RMD

---
title: "Section_8_Team1_Lab2_Report"
author: "Degnan, Rains, Wu"
date: "8/1/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Summertime vacations in the United States are synonymous with sunny days spent with loved ones. As Ella Fitzgerald so aptly sings, “Summertime, and the living is easy.” However, travel in the recent peak-vacation months has certainly deviated from this idealistic frame of mind with the onset of the COVID-19 pandemic in early 2020. The global travel industry has experienced sharp declines. In the United States, “travel spending totaled a mere $679 billion in 2020, an unprecedented 42% annual decline (nearly $500 billion) from 2019,” according to analysis from the U.S. Travel Association. 
As infection and vaccination rates oscillate along with government policies for business operation restrictions in the United States, our team of data scientists will look to answer this research question in more detail using linear regression techniques: 
Do COVID infection rates impact vacation travel distance for U.S. residents in the state of Texas? 
How does income, population density, and covid regulations in each county impact traveling? 
We will analyze datasets pertaining to these variables for Texas counties and create several models to understand if causal relationships exist between distance traveled, population income, population density (rural or non-rural counties), and regulations.

## Operationalization

## A Model Building Process

## Linear Regression Results

## Discussion: Model Limitations

## Discussion: Omitted Variable

## Conclusion & Next Steps

## References

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2021-08-01 reports/Draft_Report.Rmd
latest report draft

@@
 date: "8/1/2021"
-output: word_document
+output:
+  html_document:
+    df_print: paged
+  word_document: default
+  pdf_document: default
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
+
 ```
@@
 As infection and vaccination rates oscillate along with government policies for business operation restrictions in the United States, our team of data scientists will look to answer this research question in more detail using linear regression techniques:
+
 Do COVID infection rates impact vacation travel distance for U.S. residents in the state of Texas? 
 How does income, population density, and covid regulations in each county impact traveling?
+
 We will analyze datasets pertaining to these variables for Texas counties and create several models to understand if causal relationships exist between distance traveled, population income, population density (rural or non-rural counties), and regulations.
@@
 ## Operationalization
+
+### Variables
+
+For our analysis, we will use datasets and variables from the four following sources:
+
+[![DataSources](lab2datatable.JPG)](https://ibb.co/zQ3yj2K)
+
+As indicated, each source will provide information on county population demographics such as income, rural vs.
+urban classification, and status of pandemic regulations.
+Together, these sources were filtered for the state of Texas and combined to create a dataframe for analysis.
+
+```{r dataframe}
+#head(dataframe)
+
+```
+
+### Exploratory Data Analysis
 

### 2021-08-01 reports/Draft_Report.Rmd
latest report draft AR

@@
 
-```{r setup, include=FALSE}
+```{r setup, message=FALSE}
+install.packages("table1")
+
+#we may not need all of these
+library(tidyverse)
+library(magrittr)
+library(ggplot2)
+library(patchwork)
+library(sandwich)
+library(lmtest)
+library(dplyr)
+library(table1)
+theme_set(theme_minimal())
 knitr::opts_chunk$set(echo = TRUE)
@@
 
-```{r dataframe}
-#head(dataframe)
+```{r dataframe, echo=FALSE}
+df <- read.csv('~/lab-2-lab2_section8_jal/data/interim/covid_trips_interim.csv')
+head(df)
 
@@
 
-### Exploratory Data Analysis
+### Accounting Tables
+```{r Accounting Table creation & output}
+```
+
+#### Additional Changes to the Data:
+
+Upon final review of the dataset, there are no additional changes required to prepare the dataset for further exploration and model creation.
+Outliers or incorrect data are not observed upon final inspection.
+The Accounting Tables capture the changes made through cleaning and filtering as indicated previously.
+
+## Exploratory Data Analysis
+
+```{r EDA}
+#Example Table
+table1::label(df$total_pop) <- "TX County Population"
+table1::label(df$sum_cases) <- "TX County Covid-19 Cases"
+
+table1::table1(~total_pop + sum_cases | state, data = df)
+
+#probably going to have to transform total_pop
+
+ggplot(df, aes(x=total_pop)) + geom_histogram(bins=100) 
+ggplot(df, aes(x=sum_cases)) + geom_histogram(bins=100)
+ggplot(df, aes(x=sum_deaths)) + geom_histogram(bins=100) 
+ggplot(df, aes(x=long_trip)) + geom_histogram(bins=100)
+
+```
 
@@
 
-## Discussion: Model Limitations
+## Discussion
+###Model Limitations
 
-## Discussion: Omitted Variable
+```{r CLM assumptions check}
+
+#run this to display the CLM grapsh needed to verify CLM assumptions met
+#par(mfrow = c(2, 2))
+#plot(model)
+
+```
+
+###Omitted Variable
+
+<!-- If the team has taken up an explanatory (i.e.
+causal) question to evaluate, then identify what you think are the most important omitted variables that bias results you care about.
+For each variable you name, you should reason about the direction of bias caused by omitting this variable.
+If you can argue whether the bias is large or small, that is even better.
+State whether you have any variables available that may proxy (even imperfectly) for the omitted variable.
+Pay particular attention to whether each omitted variable bias is towards zero or away from zero.
+You will use this information to judge whether the effects you find are likely to be real, or whether they might be entirely an artifact of omitted variable bias.
+-->
+
+#### Variable 1
+#### Variable 2
 
@@
 ## References
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2021-08-02 reports/Draft_Report.Rmd
Lastest Draft 8_2

@@
 
-Summertime vacations in the United States are synonymous with sunny days spent with loved ones.
+Summer vacations in the United States are synonymous with sunny, leisurely days spent with loved ones.
 As Ella Fitzgerald so aptly sings, “Summertime, and the living is easy.” However, travel in the recent peak-vacation months has certainly deviated from this idealistic frame of mind with the onset of the COVID-19 pandemic in early 2020.
 The global travel industry has experienced sharp declines.
-In the United States, “travel spending totaled a mere $679 billion in 2020, an unprecedented 42% annual decline (nearly $500 billion) from 2019,” according to analysis from the U.S.
+Specifically in the United States, “travel spending totaled a mere $679 billion in 2020, an unprecedented 42% annual decline (nearly $500 billion) from 2019,” according to analysis from the U.S.
 Travel Association.
-As infection and vaccination rates oscillate along with government policies for business operation restrictions in the United States, our team of data scientists will look to answer this research question in more detail using linear regression techniques:
+As infection and vaccination rates oscillate regionally along with government policies for business operation restrictions in the United States, our team of data scientists will look to answer the following research question in more detail using linear regression techniques:
 
@@
 
-We will analyze datasets pertaining to these variables for Texas counties and create several models to understand if causal relationships exist between distance traveled, population income, population density (rural or non-rural counties), and regulations.
+We will analyze datasets pertaining to these variables for Texas counties and create three models to understand if causal relationships exist between distance traveled, population income, population density (rural or urban counties), and government regulations.
+Our hypothesis is that an there is a causal relationship between """
 
@@
 ### Accounting Tables
+
+To demonstrate the operationalization of the data, Tables 1 and 2 below display the original value counts from the dataset as well as how many values were removed as a part of filtering the dataset to remove all states except Texas.
+For the analysis, $$### responses were analyzed as shown in Table 1.
+Table 2 depicts the county type breakdown for rural or urban classification as well.
+
 ```{r Accounting Table creation & output}
+
+step_num = c(1,2,3)
+
+number_of_samples = c(nrow(df),
+                      nrow(#newdf1),
+                      nrow(#newdf2))                      
+samples_removed = c(0,
+                    nrow(df)-nrow(#newdf1),
+                    nrow(#newdf1)-nrow(#newdf2))
+
+reasoning = c("Original Dataset",
+              "Removing XYZ",
+              "Removing 123")
+
+accounting_df <- data.frame(step_num,
+                            number_of_samples,
+                            samples_removed,
+                            reasoning)
+
+# knitr::kable(accounting_df)
+colnames(accounting_df) <- c('Step','Number of Samples','Samples Removed', 'Reason')
+knitr::kable(accounting_df, caption="Accounting Table", align = "c")
+
+# final_depRep_tab_df = dplyr::count(Q2_df, DemorRep, sort = TRUE)
+# colnames(final_depRep_tab_df) <- c('Voter Party', 'Number of Samples')
+# knitr::kable(final_depRep_tab_df, caption="Final Samples by Party", align = "c")
 ```
@@
 
-Upon final review of the dataset, there are no additional changes required to prepare the dataset for further exploration and model creation.
+Upon final review of the dataset, there are no additional changes required to prepare the dataset for data exploration and subsequent model creation.
 Outliers or incorrect data are not observed upon final inspection.
@@
 
+Upon transformation of the data based on the figures above, we can see that poupulation size...
+
 ## A Model Building Process
@@
 ## Linear Regression Results
+
+```{r model creation}
+
+model <- lm(, data = )
+model2 <- lm(, data = )
+model3 <- lm(, data = ) #final model
+
+```
 
@@
 
-#run this to display the CLM grapsh needed to verify CLM assumptions met
+#run this to display the CLM graphs needed to verify CLM assumptions met
 #par(mfrow = c(2, 2))
-#plot(model)
+#plot(model2)
+
+##FINAL MODEL
+#par(mfrow = c(2, 2))
+#plot(model3)
 

### 2021-08-02 src/data/DataCleaning.Rmd
housekeeping

@@
----
-title: "Lab2"
-output: html_document
----
-
 ```{r setup, include=FALSE}
@@
 ```
-
-## R Markdown
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
 

### 2021-08-02 src/data/DataCleaning.Rmd

@@
 ```{r read in data, echo=FALSE}
-d_covid <- read.csv(url('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties-recent.csv'))
+d_covid <- read.csv(url('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'))
 d_openings <- read.csv(file = './../../data/raw/COVID-19 US state policy database (CUSP) - Closures & Reopening.csv')
@@
 end_date = as.Date("2021-06-30")
+
+# start_date = as.Date("2021-06-01")
+# end_date = as.Date("2021-06-30")
 ```
@@
   filter(
-    date >= start_date, date <= end_date,
+    (date >= start_date & date <= end_date),
     grepl(paste(state_id, collapse = "|"), state)
@@
 
-# aggregate county data
+# max county data (cases and deaths are cumulative)
 d_covid <- d_covid %>%
@@
   mutate(
-    sum_cases = sum(cases),
-    sum_deaths = sum(deaths),
+    cases = max(cases) - min(cases),
+    deaths = max(deaths) - min(cases),
     state = state.abb[match(state, state.name)]
@@
   select(
-    county, state, fips, sum_cases, sum_deaths
+    county, state, fips, cases, deaths
   ) %>%
@@
   filter(
-    Date >= start_date, Date <= end_date,
+    (Date >= start_date & Date <= end_date),
     grepl(paste(state_id, collapse = "|"), State.Postal.Code)

### 2021-08-03 reports/Draft_Report.Rmd
report adds- operationalization and model updates

@@
 
+To create our dataframe, we cleaned and filtered data from multiple sources of data and joined based on texas county name or code.
+
+#### Operationalization of focus variables
+We gathered covid case and death data from NYT.
+We gathered trip distance data from the Bureau of Transportation, which they record based on data gathered from personal mobile devices.
+This data was presented as number of travelers in binned distances.
+We defined a short trip as less than 100 miles and long trips as greater than 100 miles.
+This was decided under the assumption that distances traveled of less than 100 miles are for the purpose of commuting, while greater than 100 miles indicated vacation travel.
+
+#### Operationalization of additional variables
+To enrich our model and limit OVB, we brainstormed a number of additional variables that could affect travel rates during covid.
+For one, we looked at location due to the assumption that those who live in rural settings are likely to commute farther than those in an urban setting.
+We also looked at age, gender, income demographics.
+Finally, we pulled in data for county population to normalize all the data on a per capita basis.
+
 ### Variables
@@
 ```{r dataframe, echo=FALSE}
-df <- read.csv('~/lab-2-lab2_section8_jal/data/interim/covid_trips_interim.csv')
+df <- read.csv('/home/jovyan/r_bridge/code/lab-2-lab2_section8_jal/data/interim/covid_trips_interim.csv')
 head(df)
@@
 
-ggplot(df, aes(x=total_pop)) + geom_histogram(bins=100) 
+ggplot(df, aes(x=total_pop)) + geom_histogram(bins=100))
 ggplot(df, aes(x=sum_cases)) + geom_histogram(bins=100)
@@
 
-model <- lm(, data = )
+#need to change these to per capita data
+model1 <- lm(long_trip ~ cases + deaths, data = df) #limited model
 model2 <- lm(, data = )
@@
 
+#Run F test to show statistical significance of additional variables
+anova(model1, model2, test = "F")
+anova(model2, model3, test = "F")
+
+#Run coeftest to show statistical significanve of all variables in the model
+coeftest(model3, vcov = vcovHC(model3, type="HC1"))
 ```

### 2021-08-03 reports/Draft_Report.Rmd
added OMVB draft language and GGally code.

@@
 install.packages("table1")
+install.packages("GGally")
 
@@
 library(table1)
+library("GGally")
+
 theme_set(theme_minimal())
@@
 
-To create our dataframe, we cleaned and filtered data from multiple sources of data and joined based on texas county name or code.
+To create our dataframe, we cleaned and filtered data from multiple sources of data and joined based on Texas county name or code.
 
@@
 
-# final_depRep_tab_df = dplyr::count(Q2_df, DemorRep, sort = TRUE)
-# colnames(final_depRep_tab_df) <- c('Voter Party', 'Number of Samples')
-# knitr::kable(final_depRep_tab_df, caption="Final Samples by Party", align = "c")
+# final_depRep_tab_df = dplyr::count(df, #df2, sort = TRUE)
+# colnames(final_depRep_tab_df) <- c('County Type', 'Number of Samples')
+# knitr::kable(final_depRep_tab_df, caption="Final Samples by County", align = "c")
 ```
@@
 ggplot(df, aes(x=long_trip)) + geom_histogram(bins=100)
+
+ggpairs(df, columns = 1:ncol(data), title = "EDA: Variable Correlation",  
+  axisLabels = "show", columnLabels = colnames(data[, columns]))
 
@@
 #### Variable 1
+
+One example of an omitted variable that may cause bias in our model is that of vehicle ownership among the population.
+It can certainly be expected that individuals that own a vehicle would be able to more easily travel short or long distances more frequently with relative ease.
+If an individual owned a vehicle, this would influence our model by driving the bias further from zero.
+From the data available, there would no immediate proxy for this omitted variable.
+
 #### Variable 2
+
+Another example of an omitted variable that may also cause bias in our model is fuel prices.
+Whether this is jet fuel or gasoline, fluctuations in this commodity would certainly impact travel costs.
+If fuel prices increase, it is generally observed that costs for travel increase.
+The opposite is also generally true.
+In the instance of increased fuel prices, this would influence our model by driving the bias closer to zero as distance traveled would drop.
+Conversely, if fuel prices for the time period of the sample for the population were lower, the bias would would be farther from zero.
 

### 2021-08-03 notebooks/accounting_table.Rmd
added unemployment data

---
title: "accounting table"
output: html_document
---

```{r load packages}
library(dplyr)
library(tinytex)
library(knitr)
library(kableExtra)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r accounting-data}
stage <- c("raw", "date_state", "county", "model_1", "model_2", "model_3")
covid <- c(1576181, 7620, 254,230,"","")
mobility <- c(97402, 7620, 254,230,"","")
urb_rural <- c(3220, "value_2","","","","")
econ <- c(3220, , )
income <- c("value_1", "value_2","","","","")
adjacency <- c("value_1", "value_2","","","","")

df <- data.frame(stage, covid, mobility, urb_rural, income, adjacency)

```

``` {r make population summary table, echo=FALSE}
accounting_summary <- table(
    df$stage, 
    df$covid),
    df$mobility,
    df$urb_rural,
    df$income,
    df$adjacency
)
# names(dimnames(party_summary)) <- c("Self-reported", "Sentiment-revealed")
```

``` {r accounting-table, echo=FALSE}
kbl(accounting_summary, 
    digits = 2,
    caption = 'Accounting Table for Data Source Cleaning',
    booktabs = TRUE) %>%
  kable_paper(c('hover', 'striped'), full_width = T)
```

### 2021-08-03 src/data/DataCleaning.Rmd
added unemployment data

@@
 
-```{r merged dataframe, echo=FALSE}
+```{r merged model_1 dataframe, echo=FALSE}
 #code to merge county info to d
@@
 ```
-```{r export interim dataframe as csv, echo=FALSE}
+
+# Clean other variables
+```{r read in variable data, echo=FALSE}
+density <- read.delim("./../../data/raw/2019_Gaz_counties_national.txt", header=TRUE, sep="\t")
+dp03_econ <- read.csv(file = './../../data/raw/ACSDP5Y2019.DP03_data_with_overlays_2021-07-26T143523.csv')
+dp05_age <- read.csv(file = './../../data/raw/ACSDP1Y2019.DP05_data_with_overlays_2021-07-21T215013.csv')
+s1901_income <- read.csv(file = './../../data/raw/ACSST1Y2019.S1901_data_with_overlays_2021-07-21T222906.csv') 
+urb_rur_h2 <- read.csv(file = './../../data/raw/DECENNIALSF12010.H2_data_with_overlays_2021-06-08T173848.csv')
+
+```
+
+```{r}
+library(dplyr)
+df = df %>% mutate(D=gsub("\\..*","",A))
+
+              A    B   C    D
+   awer.ttp.net Code 554 awer
+   abcd.ttp.net Code 747 abcd
+   asdf.ttp.net Part 554 asdf
+    xyz.ttp.net Part 747  xyz
+
+    names(df) <- substring(names(df), 4)
+```
+
+```{r clean dp03 data, echo=FALSE}
+# filters by date/state
+econ <- dp03_econ %>%
+  rename(fips = GEO_ID, unemp_rate = DP03_0009PE) %>%
+  select(fips, unemp_rate) %>%
+  mutate(
+    fips = substring(fips, 10)
+  )
+
+# merge into main dataframe
+d <- merge(x = econ, y = d, by = c("fips"))
+```
+
+# Clean up
+```{r remove interim dataframes}
+# remove dataframes
+rm(d_covid, d_trips, d_openings)
+rm(dp03_econ, dp05_age, density, s1901_income, urb_rur_h2)
+
+```
+
+# Write to .csv
+```{r export interim model_1 dataframe as csv, echo=FALSE}
 write.csv(d, '../../data/interim/covid_trips_interim.csv')

### 2021-08-03 reports/Draft_Report.Rmd
added more to intro, code for ggaly.

@@
 author: "Degnan, Rains, Wu"
-date: "8/1/2021"
+date: "8/5/2021"
 output:
@@
 We will analyze datasets pertaining to these variables for Texas counties and create three models to understand if causal relationships exist between distance traveled, population income, population density (rural or urban counties), and government regulations.
-Our hypothesis is that an there is a causal relationship between """
+The outcome variable will be distance traveled.
+The predictor varialbes will be income, county type (rural or urban), and policy regulations around Covid-19.
+Our hypothesis is that an there is a causal relationship between these predictor variables and the outcome variable.
 
@@
 ```{r dataframe, echo=FALSE}
-df <- read.csv('/home/jovyan/r_bridge/code/lab-2-lab2_section8_jal/data/interim/covid_trips_interim.csv')
+df <- read.csv('~/lab-2-lab2_section8_jal/data/interim/covid_trips_interim.csv')
 head(df)
@@
 To demonstrate the operationalization of the data, Tables 1 and 2 below display the original value counts from the dataset as well as how many values were removed as a part of filtering the dataset to remove all states except Texas.
-For the analysis, $$### responses were analyzed as shown in Table 1.
+For the analysis, 230 responses were analyzed as shown in Table 1.
 Table 2 depicts the county type breakdown for rural or urban classification as well.
@@
 
-step_num = c(1,2,3)
+step_num = c(1,2)
 
-number_of_samples = c(nrow(df),
-                      nrow(#newdf1),
-                      nrow(#newdf2))                      
+number_of_samples = c(nrow(df), #Need dataframe for all US
+                      nrow(df))
+
 samples_removed = c(0,
-                    nrow(df)-nrow(#newdf1),
-                    nrow(#newdf1)-nrow(#newdf2))
+                    nrow(df)-nrow(df))
 
 reasoning = c("Original Dataset",
-              "Removing XYZ",
-              "Removing 123")
+              "Filtered for Texas Only")
 
@@
 
-# knitr::kable(accounting_df)
+knitr::kable(accounting_df)
 colnames(accounting_df) <- c('Step','Number of Samples','Samples Removed', 'Reason')
@@
 
-# final_depRep_tab_df = dplyr::count(df, #df2, sort = TRUE)
-# colnames(final_depRep_tab_df) <- c('County Type', 'Number of Samples')
-# knitr::kable(final_depRep_tab_df, caption="Final Samples by County", align = "c")
+#Need to create column category of rural vs urban - will create a summary table below with this
+# final_tab_df = dplyr::count(df, df, sort = TRUE)
+# colnames(final_tab_df) <- c('County Type', 'Number of Samples')
+# knitr::kable(final_tab_df, caption="Final Samples by County", align = "c")
 ```
@@
 table1::label(df$total_pop) <- "TX County Population"
-table1::label(df$sum_cases) <- "TX County Covid-19 Cases"
+table1::label(df$cases) <- "TX County Covid-19 Cases"
 
-table1::table1(~total_pop + sum_cases | state, data = df)
+table1::table1(~total_pop + cases | state, data = df)
 
@@
 
-ggplot(df, aes(x=total_pop)) + geom_histogram(bins=100))
-ggplot(df, aes(x=sum_cases)) + geom_histogram(bins=100)
-ggplot(df, aes(x=sum_deaths)) + geom_histogram(bins=100) 
+ggplot(df, aes(x=total_pop)) + geom_histogram(bins=100)
+ggplot(df, aes(x=cases)) + geom_histogram(bins=100)
+ggplot(df, aes(x=deaths)) + geom_histogram(bins=100) 
 ggplot(df, aes(x=long_trip)) + geom_histogram(bins=100)
 
-ggpairs(df, columns = 1:ncol(data), title = "EDA: Variable Correlation",  
-  axisLabels = "show", columnLabels = colnames(data[, columns]))
+df_test <- df
+
+df_test <- df_test %>% mutate(County_Class = 
+               case_when(df_test$total_pop <= 99999 ~ "Rural", 
+                         df_test$total_pop >= 100000 ~ "Urban"))
+
+# df_test$County_Rural <- ifelse(df_test$County_Class == "Rural", 1, 0) #will be useful for model
+# df_test$County_Class <- as.factor(df_test$County_Class) #this turns
+
+ggpairs(df_test, columns = 5:(ncol(df)-1), title = "EDA: Variable Correlation",  
+  axisLabels = "show", aes(colour = County_Class, alpha = 0.4))
 
@@
 ###Model Limitations
+
+<!-- Add language for model 2 and model 3 about CLM assumptions. -->
 
@@
 
+Based upon the linear regression analysis above, it is observed that there may be a causal relationship between XX and YY.
+The hypothesis posed within the introduction is 'failed to be rejected/ rejected.' ....
+
+For future analysis, it is recommended that Covid-19 policy adherence data per county be incorporated into the model to increase the accuracy.
+While the Govenor of Texas issued detailed protocols or mandates for the state of Texas, the degree to which each county (and thus the individual residents) complied would be useful data to have if such data were to exist.
+
 ## References

### 2021-08-03 src/data/DataCleaning.Rmd
added age and income data.

@@
 d <- unique(d)
-
-#merge county policies to d
-# d_openings$strict <- d_openings$Closed.bars.x2 != "0"
-
-# d_openings$State.Postal.Code <- d_openings$State.Abbreviation
-# d <- merge(x = d, y = d_openings[ , c("State.Postal.Code", "strict")], by = "State.Postal.Code", all.x = TRUE)
 ```
@@
 
-```{r}
-library(dplyr)
-df = df %>% mutate(D=gsub("\\..*","",A))
-
-              A    B   C    D
-   awer.ttp.net Code 554 awer
-   abcd.ttp.net Code 747 abcd
-   asdf.ttp.net Part 554 asdf
-    xyz.ttp.net Part 747  xyz
-
-    names(df) <- substring(names(df), 4)
-```
-
 ```{r clean dp03 data, echo=FALSE}
@@
   mutate(
-    fips = substring(fips, 10)
+    fips = substring(fips, 10),
+    unemp_rate = as.double(unemp_rate)
   )
@@
 
+```{r clean dp03 data, echo=FALSE}
+# filters by date/state
+age <- dp05_age %>%
+  rename(fips = GEO_ID,
+         female = DP05_0003PE, 
+         pct_lt5 = DP05_0005PE, 
+         pct_5to9 = DP05_0006PE, 
+         pct_10to14 = DP05_0007PE, 
+         pct_15to19 = DP05_0008PE, 
+         pct_20to24 = DP05_0009PE, 
+         pct_25to34 = DP05_0010PE, 
+         pct_35to44 = DP05_0011PE, 
+         pct_45to54 = DP05_0012PE
+) %>% rowwise() %>%
+  mutate(
+    fips = substring(fips, 10),
+    female = as.double(female), 
+    pct_lt5 = as.double(pct_lt5), 
+    pct_5to9 = as.double(pct_5to9), 
+    pct_10to14 = as.double(pct_10to14), 
+    pct_15to19 = as.double(pct_15to19), 
+    pct_20to24 = as.double(pct_20to24), 
+    pct_25to34 = as.double(pct_25to34), 
+    pct_35to44 = as.double(pct_35to44), 
+    pct_45to54 = as.double(pct_45to54),
+    pct_gt55 = 100 - sum(c(pct_lt5, pct_5to9, pct_10to14, pct_15to19, pct_20to24, pct_25to34, pct_35to44, pct_45to54)),
+    pct_lt19 = sum(c(pct_lt5, pct_5to9, pct_10to14, pct_15to19)),
+    pct_10to54 = 100 - pct_gt55 - pct_lt19
+    ) %>%
+  select(fips, pct_lt19, pct_10to14, pct_gt55)
+
+# merge into main dataframe
+d <- merge(x = age, y = d, by = c("fips"))
+```
+
+```{r clean s1901 data, echo=FALSE}
+# filters by date/state
+income <- s1901_income %>%
+  rename(fips = GEO_ID,
+  median_income = S1901_C01_012E,  
+  pct_50to75k = S1901_C01_007E, 
+  pct_75to100k = S1901_C01_008E, 
+  pct_100to150k = S1901_C01_009E, 
+  pct_150to200k = S1901_C01_010E, 
+  pct_gt200k = S1901_C01_011E
+) %>% rowwise() %>%
+  mutate(
+    fips = substring(fips, 10),
+    median_income = as.double(median_income), 
+    pct_50to75k = as.double(pct_50to75k), 
+    pct_75to100k = as.double(pct_75to100k), 
+    pct_100to150k = as.double(pct_100to150k), 
+    pct_150to200k = as.double(pct_150to200k), 
+    pct_over_us_median = sum(c(pct_50to75k, pct_75to100k, pct_100to150k, pct_150to200k))) %>%
+  select(fips, median_income, pct_over_us_median)
+
+# merge into main dataframe
+d <- merge(x = income, y = d, by = c("fips"))
+```
+
 # Clean up
@@
 rm(d_covid, d_trips, d_openings)
-rm(dp03_econ, dp05_age, density, s1901_income, urb_rur_h2)
+rm(dp03_econ, dp05_age, density, s1901_income, urb_rur_h2, econ, income, age)
 

### 2021-08-03 src/data/DataCleaning.Rmd
fixed COVID death calculation

@@
     cases = max(cases) - min(cases),
-    deaths = max(deaths) - min(cases),
+    deaths = max(deaths) - min(deaths),
     state = state.abb[match(state, state.name)]

### 2021-08-03 src/data/DataCleaning.Rmd

@@
     pct_lt19 = sum(c(pct_lt5, pct_5to9, pct_10to14, pct_15to19)),
-    pct_10to54 = 100 - pct_gt55 - pct_lt19
+    pct_19to54 = 100 - pct_gt55 - pct_lt19
     ) %>%
-  select(fips, pct_lt19, pct_10to14, pct_gt55)
+  select(fips, pct_lt19, pct_19to54, pct_gt55)
 

### 2021-08-03 src/data/DataCleaning.Rmd

@@
 
+unemp_claims <- readxl::read_excel("./../../data/raw/weekly-claims-by-county-twc.xlsx")
+
 ```
@@
 ```{r clean dp03 data, echo=FALSE}
-# filters by date/state
 econ <- dp03_econ %>%
@@
 d <- merge(x = econ, y = d, by = c("fips"))
+```
+```{r clean twc unemployment data, echo=FALSE}
+
+uclaims <- unemp_claims[-c(1,2),] %>%
+  select(TEXAS,
+         ...6,
+         ...7,
+         ...8,
+         ...9
+         ) %>% 
+  rowwise() %>%
+  mutate(
+    ...6 = as.integer(...6),
+    ...7 = as.integer(...7),
+    ...8 = as.integer(...8),
+    ...9 = as.integer(...9),
+    total_claims = sum(c(...6, ...7, ...8, ...9))
+  ) %>%
+  rename(
+    county = TEXAS
+  ) %>% drop_na()
+
+# merge into main dataframe
+d <- merge(x = select(uclaims, c(county, total_claims)), y = d, by = c("county"))
 ```

### 2021-08-04 src/data/DataCleaning.Rmd
limited model dataframe completed

@@
 library(tidyverse)
-# source(local=TRUE, file="../util.R")
 ```
+
+# limited model data cleaning
 
@@
 d_covid <- read.csv(url('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'))
-d_openings <- read.csv(file = './../../data/raw/COVID-19 US state policy database (CUSP) - Closures & Reopening.csv')
 d_trips <- read.csv(file = './../../data/raw/Trips_by_Distance.csv')
+urb_rur_h2 <- read.csv(file = './../../data/raw/DECENNIALSF12010.H2_data_with_overlays_2021-06-08T173848.csv') 
+density <- read.csv(file='./../../data/raw/TAC_population.csv')
+income <- read.csv('./../../data/raw/TAC_income.csv')
 ```
@@
 end_date = as.Date("2021-06-30")
-
-# start_date = as.Date("2021-06-01")
-# end_date = as.Date("2021-06-30")
 ```
@@
 
-# cleaning
+# cleaning/county filter
 d_trips <- d_trips %>%
@@
     long_trip = sum(c_across(Number.of.Trips.100.250:Number.of.Trips...500)),
-    long_trip_pc = long_trip/total_pop,
     # define short trips
-    short_trip = sum(c_across(Number.of.Trips.1.3:Number.of.Trips.50.100)),
-    short_trip_pc = short_trip/total_pop
+    # short_trip = sum(c_across(Number.of.Trips.1.3:Number.of.Trips.50.100)),
   ) %>%
@@
   select(
-    county, fips, state, total_pop, long_trip, long_trip_pc, short_trip, short_trip_pc
+    county, fips, state, total_pop, long_trip
   ) %>%
   distinct(fips, .keep_all = TRUE)
+```
+
+```{r clean urban/rural data}
+rural_ind <- urb_rur_h2[-c(1),] %>%
+  rename(fips = GEO_ID,
+         total_pop = H002001,
+         urbanized = H002003,
+         rural = H002005) %>%
+  select(fips, total_pop, urbanized, rural) %>%
+  mutate(
+    fips = substring(fips, 10),
+    total_pop = as.double(total_pop),
+    urbanized = as.double(urbanized),
+    rural = as.double(rural),
+    is_rural = ifelse((total_pop < 50000 & urbanized == 0), 1, 0)
+  ) %>% 
+  select(fips, is_rural) %>% drop_na()
+```
+
+```{r clean income data}
+income_f <- income %>%
+  rename(
+    county = County,
+    median_income = Median.Household.Income,
+    povery_rate = X..of.Population.in.Poverty
+  ) %>%
+  select(
+    county, median_income, povery_rate
+  ) %>%
+  mutate(
+    median_income = as.double(gsub('[$,]', '', median_income))
+  )
+
+```
+
+```{r clean density data}
+density_f <- density %>%
+  rename(
+    county = County,
+    total_pop = County.Population,
+    density = X.Population.Density.Per.Sq.Mile
+  ) %>%
+  select(
+    county, total_pop, density
+  ) %>%
+  mutate(
+    total_pop = as.double(gsub('[,]', '', total_pop)),
+    density = as.double(gsub('[,]', '', density))
+  )
+
+```
+
+```{r merge limited model dataframe}
+d <- merge(x = d_trips, y = d_covid, by = c("fips", "county", "state")) 
+d <- merge(x = rural_ind, y = d, by = c("fips"))
+d <- merge(x = income_f, y = d, by = c("county"))
+d <- merge(x = density_f, y = d, by = c("county"))
+d <- d %>% drop_na()
+
+#rm(d)
+```
+
+# Write to .csv
+```{r export interim model_1 dataframe as csv, echo=FALSE}
+write.csv(d, '../../data/interim/covid_trips_interim.csv')
 ```
@@
 
+# Model 2 data cleaning
+
+# Model 3 data cleaning
+
 # Clean other variables
@@
 s1901_income <- read.csv(file = './../../data/raw/ACSST1Y2019.S1901_data_with_overlays_2021-07-21T222906.csv')
-urb_rur_h2 <- read.csv(file = './../../data/raw/DECENNIALSF12010.H2_data_with_overlays_2021-06-08T173848.csv')
 
 unemp_claims <- readxl::read_excel("./../../data/raw/weekly-claims-by-county-twc.xlsx")
-
+ur <- readxl::read_excel("./../../data/raw/TX_Rural_County.xlsx")
 ```
@@
 ```
+
 ```{r clean twc unemployment data, echo=FALSE}
@@
 
+```{r rural indicator, echo=FALSE}
+d$rural_ind <- ifelse(d$county %in% ur$`Rural County`, 1, 0)
+
+summary(d$rural_ind)
+
+```
+
 # Clean up

### 2021-08-04 src/data/DataCleaning.Rmd
removed duplicate columns.

@@
     county = County,
-    total_pop = County.Population,
     density = X.Population.Density.Per.Sq.Mile
+    # total_pop = County.Population,
   ) %>%
   select(
-    county, total_pop, density
+    county, 
+    density 
+    #total_pop
   ) %>%
   mutate(
-    total_pop = as.double(gsub('[,]', '', total_pop)),
+    #total_pop = as.double(gsub('[,]', '', total_pop)),
     density = as.double(gsub('[,]', '', density))
@@
 
+knitr::kable(head(d))
 #rm(d)

### 2021-08-04 src/data/DataCleaning.Rmd

@@
 ```{r export interim model_1 dataframe as csv, echo=FALSE}
-write.csv(d, '../../data/interim/covid_trips_interim.csv')
-```
-
-```{r merged model_1 dataframe, echo=FALSE}
-#code to merge county info to d
-d <- merge(x = d_trips, y = d_covid, by = c("fips", "county", "state"))
-
-#drop na and duplicates
-d <- d %>% drop_na()
-d <- unique(d)
+write.csv(d, '../../data/interim/covid_trips_interim_limited.csv')
 ```
@@
 
+```{r load unemployment data}
+unemploy <- read.csv('./../../data/raw/TAC_unemployment.csv')
+```
+
+```{r clean unemployment data}
+unemploy_f <- unemploy %>%
+  rename(
+    county = County,
+    unempoy_rate = Unemployment.Rate....
+  ) %>%
+  select(
+    county, 
+    unempoy_rate 
+  )
+
+```
+
+```{r merge model 2 dataframe}
+d <- merge(x = unemploy_f, y = d, by = c("county"))
+d <- d %>% drop_na()
+
+knitr::kable(head(d))
+#rm(d)
+```
+
+# Write to .csv
+```{r export interim model_1 dataframe as csv, echo=FALSE}
+write.csv(d, '../../data/interim/covid_trips_interim_model2.csv')
+```
+
 # Model 3 data cleaning
 
-# Clean other variables
-```{r read in variable data, echo=FALSE}
-density <- read.delim("./../../data/raw/2019_Gaz_counties_national.txt", header=TRUE, sep="\t")
-dp03_econ <- read.csv(file = './../../data/raw/ACSDP5Y2019.DP03_data_with_overlays_2021-07-26T143523.csv')
-dp05_age <- read.csv(file = './../../data/raw/ACSDP1Y2019.DP05_data_with_overlays_2021-07-21T215013.csv')
-s1901_income <- read.csv(file = './../../data/raw/ACSST1Y2019.S1901_data_with_overlays_2021-07-21T222906.csv')
-
-unemp_claims <- readxl::read_excel("./../../data/raw/weekly-claims-by-county-twc.xlsx")
-ur <- readxl::read_excel("./../../data/raw/TX_Rural_County.xlsx")
+```{r load age data}
+age <- read.csv('./../../data/raw/TAC_age.csv')
 ```
 
-```{r clean dp03 data, echo=FALSE}
-econ <- dp03_econ %>%
-  rename(fips = GEO_ID, unemp_rate = DP03_0009PE) %>%
-  select(fips, unemp_rate) %>%
+```{r clean unemployment data}
+age_f <- age %>%
+  rename(
+    county = County,
+    pct_lt17 = X..Age.17.and.Under
+  ) %>% rowwise() %>%
   mutate(
-    fips = substring(fips, 10),
-    unemp_rate = as.double(unemp_rate)
+    pct_gt65 = sum(c(X..Age.65.and.Older, X..Age.85.and.Older))
+  ) %>%
+  select(
+    county,
+    pct_lt17,
+    pct_gt65
   )
 
-# merge into main dataframe
-d <- merge(x = econ, y = d, by = c("fips"))
 ```
 
-```{r clean twc unemployment data, echo=FALSE}
+```{r merge limited model dataframe}
+d <- merge(x = age_f, y = d, by = c("county"))
+d <- d %>% drop_na()
 
-uclaims <- unemp_claims[-c(1,2),] %>%
-  select(TEXAS,
-         ...6,
-         ...7,
-         ...8,
-         ...9
-         ) %>% 
-  rowwise() %>%
-  mutate(
-    ...6 = as.integer(...6),
-    ...7 = as.integer(...7),
-    ...8 = as.integer(...8),
-    ...9 = as.integer(...9),
-    total_claims = sum(c(...6, ...7, ...8, ...9))
-  ) %>%
-  rename(
-    county = TEXAS
-  ) %>% drop_na()
-
-# merge into main dataframe
-d <- merge(x = select(uclaims, c(county, total_claims)), y = d, by = c("county"))
+knitr::kable(head(d))
+#rm(d)
 ```
 
-```{r clean dp03 data, echo=FALSE}
-# filters by date/state
-age <- dp05_age %>%
-  rename(fips = GEO_ID,
-         female = DP05_0003PE, 
-         pct_lt5 = DP05_0005PE, 
-         pct_5to9 = DP05_0006PE, 
-         pct_10to14 = DP05_0007PE, 
-         pct_15to19 = DP05_0008PE, 
-         pct_20to24 = DP05_0009PE, 
-         pct_25to34 = DP05_0010PE, 
-         pct_35to44 = DP05_0011PE, 
-         pct_45to54 = DP05_0012PE
-) %>% rowwise() %>%
-  mutate(
-    fips = substring(fips, 10),
-    female = as.double(female), 
-    pct_lt5 = as.double(pct_lt5), 
-    pct_5to9 = as.double(pct_5to9), 
-    pct_10to14 = as.double(pct_10to14), 
-    pct_15to19 = as.double(pct_15to19), 
-    pct_20to24 = as.double(pct_20to24), 
-    pct_25to34 = as.double(pct_25to34), 
-    pct_35to44 = as.double(pct_35to44), 
-    pct_45to54 = as.double(pct_45to54),
-    pct_gt55 = 100 - sum(c(pct_lt5, pct_5to9, pct_10to14, pct_15to19, pct_20to24, pct_25to34, pct_35to44, pct_45to54)),
-    pct_lt19 = sum(c(pct_lt5, pct_5to9, pct_10to14, pct_15to19)),
-    pct_19to54 = 100 - pct_gt55 - pct_lt19
-    ) %>%
-  select(fips, pct_lt19, pct_19to54, pct_gt55)
-
-# merge into main dataframe
-d <- merge(x = age, y = d, by = c("fips"))
-```
-
-```{r clean s1901 data, echo=FALSE}
-# filters by date/state
-income <- s1901_income %>%
-  rename(fips = GEO_ID,
-  median_income = S1901_C01_012E,  
-  pct_50to75k = S1901_C01_007E, 
-  pct_75to100k = S1901_C01_008E, 
-  pct_100to150k = S1901_C01_009E, 
-  pct_150to200k = S1901_C01_010E, 
-  pct_gt200k = S1901_C01_011E
-) %>% rowwise() %>%
-  mutate(
-    fips = substring(fips, 10),
-    median_income = as.double(median_income), 
-    pct_50to75k = as.double(pct_50to75k), 
-    pct_75to100k = as.double(pct_75to100k), 
-    pct_100to150k = as.double(pct_100to150k), 
-    pct_150to200k = as.double(pct_150to200k), 
-    pct_over_us_median = sum(c(pct_50to75k, pct_75to100k, pct_100to150k, pct_150to200k))) %>%
-  select(fips, median_income, pct_over_us_median)
-
-# merge into main dataframe
-d <- merge(x = income, y = d, by = c("fips"))
-```
-
-```{r rural indicator, echo=FALSE}
-d$rural_ind <- ifelse(d$county %in% ur$`Rural County`, 1, 0)
-
-summary(d$rural_ind)
-
+# Write to .csv
+```{r export interim model_1 dataframe as csv, echo=FALSE}
+write.csv(d, '../../data/interim/covid_trips_interim_model3.csv')
 ```
@@
 ```
-
-# Write to .csv
-```{r export interim model_1 dataframe as csv, echo=FALSE}
-write.csv(d, '../../data/interim/covid_trips_interim.csv')
-```

### 2021-08-04 reports/Draft_Report.Rmd
EDA and models added to report

--- 2021-08-03 reports/Draft_Report.Rmd
@@
 library(table1)
+library(ggally)
 theme_set(theme_minimal())
@@
 This was decided under the assumption that distances traveled of less than 100 miles are for the purpose of commuting, while greater than 100 miles indicated vacation travel.
+Hand in hand with this variable is our indicator variable for urban vs rural setting *** how did we decide this again, population threshold?**** under the assumption that those who live in rural settings are likely to commute further than those in an urban setting.
+Finally, we believe that median household income plays a significant role in an individual's ability to take vacations.
+We pulled this directly from the US Census Bureau.
 
@@
 To enrich our model and limit OVB, we brainstormed a number of additional variables that could affect travel rates during covid.
-For one, we looked at location due to the assumption that those who live in rural settings are likely to commute farther than those in an urban setting.
+For one, we looked at how covid affected the county, through the unemployment rate.
 We also looked at age, gender, income demographics.
-Finally, we pulled in data for county population to normalize all the data on a per capita basis.
+Then, in addition to income demographics, we focused on percent of younger folks, as defined as those younger than 17 and older folks as defined as those older than 65 because of the nature of how covid affects an individual.
 
@@
 ```{r dataframe, echo=FALSE}
-df <- read.csv('/home/jovyan/r_bridge/code/lab-2-lab2_section8_jal/data/interim/covid_trips_interim.csv')
+df <- read.csv('/home/jovyan/r_bridge/code/lab-2-lab2_section8_jal/data/interim/covid_trips_interim_model3.csv')
 head(df)
@@
 To demonstrate the operationalization of the data, Tables 1 and 2 below display the original value counts from the dataset as well as how many values were removed as a part of filtering the dataset to remove all states except Texas.
-For the analysis, $$### responses were analyzed as shown in Table 1.
+For the analysis, ### responses were analyzed as shown in Table 1.
 Table 2 depicts the county type breakdown for rural or urban classification as well.
@@
 
-#probably going to have to transform total_pop
+#Short vs long trip: rate of short trips is higher than rate of long trips in relation to # of covid cases
+ggplot(data = df) + aes(x = log(sum_cases), y = log(long_trip)) + geom_point()
 
-ggplot(df, aes(x=total_pop)) + geom_histogram(bins=100))
-ggplot(df, aes(x=sum_cases)) + geom_histogram(bins=100)
-ggplot(df, aes(x=sum_deaths)) + geom_histogram(bins=100) 
-ggplot(df, aes(x=long_trip)) + geom_histogram(bins=100)
+ggplot(data = df) + aes(x = log(sum_cases), y = log(short_trip)) + geom_point()
+
+df %>% 
+  mutate(
+    total_pop = log(total_pop),
+    cases = log(cases+1),
+    deaths = log(deaths+1),
+    long_trip = log(long_trip),
+    density = log(density)
+  ) %>%
+  select(long_trip, cases, deaths, median_income, unempoy_rate, density, pct_lt17) %>% 
+  GGally::ggpairs(.,
+               legend = 1,
+               mapping = ggplot2::aes(colour=as.factor(df$is_rural)))
 
@@
 #need to change these to per capita data
-model1 <- lm(long_trip ~ cases + deaths, data = df) #limited model
-model2 <- lm(, data = )
-model3 <- lm(, data = ) #final model
+model1 <- lm(log(long_trip+1) ~ log(cases+1) + log(deaths+1) + is_rural + median_income, data = df) #limited model
+model2 <- lm(log(long_trip+1) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unempoy_rate + log(density), data = df)
+model3 <- lm(log(long_trip+1) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unempoy_rate + log(density) + pct_lt17, data = df) #final model
 
@@
 
-#Run coeftest to show statistical significanve of all variables in the model
+#Run coeftest to show statistical significance of all variables in the model
 coeftest(model3, vcov = vcovHC(model3, type="HC1"))
@@
 #run this to display the CLM graphs needed to verify CLM assumptions met
-#par(mfrow = c(2, 2))
-#plot(model2)
+par(mfrow = c(2, 2))
+plot(model2)
 
 ##FINAL MODEL
-#par(mfrow = c(2, 2))
-#plot(model3)
+par(mfrow = c(2, 2))
+plot(model3)
 

### 2021-08-04 reports/Draft_Report.Rmd
updated EDA visuals

@@
 library(table1)
-library(ggally)
+library(GGally)
 theme_set(theme_minimal())
@@
 ```{r dataframe, echo=FALSE}
-df <- read.csv('/home/jovyan/r_bridge/code/lab-2-lab2_section8_jal/data/interim/covid_trips_interim_model3.csv')
-head(df)
+df <- read.csv('~/lab-2-lab2_section8_jal/data/interim/covid_trips_interim_model3.csv')
+#head(df)
+```
+
+## Exploratory Data Analysis
+
+Figure 1 below depicts the distribution and correlation of each variable in our dataset by county type (where '0' indicates an urban county).
+From Figure 1, it can be observed that some of the variables are not distributed normally, are skewed,  and thus require transformations.
+More specifically, these variables are: Total Population, Total Cases, Total Deaths, Population Density.
+The outcome variable, Long Distance, will also require transformation.
+Upon further analysis, logarithmic transformations were chosen to normalize the distribution and minimize skewness.
+
+```{r EDA}
+#Example Table
+df$rural_cat <- ifelse(df$is_rural == 1, "Rural Counties", "Urban Counties")
+table1::label(df$total_pop) <- "County Population"
+table1::label(df$density) <- "County Population Density"
+table1::label(df$cases) <- "County Covid-19 Cases"
+table1::label(df$unempoy_rate) <- "County Unemployment Rate"
+table1::label(df$median_income) <- "County Median Income ($)"
+table1::label(df$pct_lt17) <- "County Age < 17 years (%)"
+table1::label(df$pct_gt65) <- "County Age > 65 years (%)"
+# table1::label(df$is_rural) <- "County Type"
+table1::table1(~total_pop + density + cases + unempoy_rate + median_income + pct_lt17 + pct_gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Statistics", topclass="Rtable1-zebra", overall = "Total")
+
+#Short vs long trip: rate of short trips is higher than rate of long trips in relation to # of covid cases
+ggplot(data = df) + aes(x = log(cases), y = log(long_trip)) + geom_point()
+#ggplot(data = df) + aes(x = log(cases), y = log(short_trip)) + geom_point()
+
+Rural_County=as.factor(df$is_rural)
+
+df %>% 
+  mutate(
+    total_pop = log(total_pop),
+    cases = log(cases+1),
+    deaths = log(deaths+1),
+    long_trip = log(long_trip),
+    density = log(density)
+  ) %>%
+  select(long_trip, cases, deaths, median_income, unempoy_rate, density, pct_lt17) %>% 
+  GGally::ggpairs(.,
+               legend = 1, title = "Figure 1: Variable Statistics",
+               mapping = ggplot2::aes(colour=Rural_County))
 
 ```
+
+Upon transformation of the data based on the figures above, we can see that poupulation size...
 
@@
 To demonstrate the operationalization of the data, Tables 1 and 2 below display the original value counts from the dataset as well as how many values were removed as a part of filtering the dataset to remove all states except Texas.
-For the analysis, ### responses were analyzed as shown in Table 1.
+For the analysis, 230 responses were analyzed as shown in Table 1.
 Table 2 depicts the county type breakdown for rural or urban classification as well.
@@
 
-step_num = c(1,2,3)
+df2 <- df
+
+df2$log_pop <- log10(df2$total_pop)
+df2$log_cases <-log10(df2$cases)
+df2$log_deaths <-log10(df2$deaths)
+df2$log_long_trip <-log10(df2$long_trip)
+
+step_num = c(1,2)
 
 number_of_samples = c(nrow(df),
-                      nrow(#newdf1),
-                      nrow(#newdf2))                      
+                      nrow(df2))
 samples_removed = c(0,
-                    nrow(df)-nrow(#newdf1),
-                    nrow(#newdf1)-nrow(#newdf2))
+                    nrow(df)-nrow(df2))
 
 reasoning = c("Original Dataset",
-              "Removing XYZ",
-              "Removing 123")
+              "Logarithmic Transformations")
 
@@
 The Accounting Tables capture the changes made through cleaning and filtering as indicated previously.
-
-## Exploratory Data Analysis
-
-```{r EDA}
-#Example Table
-table1::label(df$total_pop) <- "TX County Population"
-table1::label(df$sum_cases) <- "TX County Covid-19 Cases"
-
-table1::table1(~total_pop + sum_cases | state, data = df)
-
-#Short vs long trip: rate of short trips is higher than rate of long trips in relation to # of covid cases
-ggplot(data = df) + aes(x = log(sum_cases), y = log(long_trip)) + geom_point()
-
-ggplot(data = df) + aes(x = log(sum_cases), y = log(short_trip)) + geom_point()
-
-df %>% 
-  mutate(
-    total_pop = log(total_pop),
-    cases = log(cases+1),
-    deaths = log(deaths+1),
-    long_trip = log(long_trip),
-    density = log(density)
-  ) %>%
-  select(long_trip, cases, deaths, median_income, unempoy_rate, density, pct_lt17) %>% 
-  GGally::ggpairs(.,
-               legend = 1,
-               mapping = ggplot2::aes(colour=as.factor(df$is_rural)))
-
-```
-
-Upon transformation of the data based on the figures above, we can see that poupulation size...
 

### 2021-08-04 reports/Draft_Report.Rmd
updated EDA tabless

@@
 
+Table 1 below displays some information about the mean, median, minimum and maximum values for each variable.
 Figure 1 below depicts the distribution and correlation of each variable in our dataset by county type (where '0' indicates an urban county).
@@
 
-Upon transformation of the data based on the figures above, we can see that poupulation size...
-
 ### Accounting Tables
@@
 df2$log_long_trip <-log10(df2$long_trip)
+df2$log_density <- log10(df2$density)
 
@@
 
-# knitr::kable(accounting_df)
+#knitr::kable(accounting_df)
 colnames(accounting_df) <- c('Step','Number of Samples','Samples Removed', 'Reason')
@@
 
-# final_depRep_tab_df = dplyr::count(Q2_df, DemorRep, sort = TRUE)
-# colnames(final_depRep_tab_df) <- c('Voter Party', 'Number of Samples')
-# knitr::kable(final_depRep_tab_df, caption="Final Samples by Party", align = "c")
+final_tab_df = dplyr::count(df2, rural_cat, sort = TRUE)
+colnames(final_tab_df) <- c('County Type', 'Number of Samples')
+knitr::kable(final_tab_df, caption="Table 2: Final Samples by County Type", align = "c")
 ```
@@
 Outliers or incorrect data are not observed upon final inspection.
-The Accounting Tables capture the changes made through cleaning and filtering as indicated previously.
+The Accounting Table capture the changes made through cleaning and filtering as indicated previously.
+Table 2 above also displays the subset per county type for the overall dataset.
 

### 2021-08-04 reports/Draft_Report.Rmd
updated CLM section

@@
 install.packages("GGally")
+install.packages("stargazer")
 
@@
 library(GGally)
+library(stargazer)
 theme_set(theme_minimal())
@@
 
-df2$log_pop <- log10(df2$total_pop)
-df2$log_cases <-log10(df2$cases)
-df2$log_deaths <-log10(df2$deaths)
-df2$log_long_trip <-log10(df2$long_trip)
-df2$log_density <- log10(df2$density)
+df2$log_pop <- log10(df2$total_pop+1)
+df2$log_cases <-log10(df2$cases+1)
+df2$log_deaths <-log10(df2$deaths+1)
+df2$log_long_trip <-log10(df2$long_trip+1)
+df2$log_density <- log10(df2$density+1)
+
+#summary(df2)
+df2 <- df2 %>% filter(!is.na(df2))
+#df2 <- df2 %>% filter(!is.infinite(df2))
 
@@
 
-```{r model creation}
+```{r model creation, warning=FALSE}
 
@@
 model1 <- lm(log(long_trip+1) ~ log(cases+1) + log(deaths+1) + is_rural + median_income, data = df) #limited model
-model2 <- lm(log(long_trip+1) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unempoy_rate + log(density), data = df)
-model3 <- lm(log(long_trip+1) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unempoy_rate + log(density) + pct_lt17, data = df) #final model
+model2 <- lm(log(long_trip+1) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unempoy_rate + log(density+1), data = df)
+model3 <- lm(log(long_trip+1) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unempoy_rate + log(density+1) + pct_lt17, data = df) #final model
+
+stargazer(model1, model2, model3, type='text', title="Results", align=TRUE)
 
@@
 ## Discussion
-###Model Limitations
+### Model Limitations
+
+1.
+**IID Sampling:** There isn't a direct database check for this unfortunately.
+Looking into how this data was generated, I belive this is IID generally.
+The data was gathered from independent YouTube uploaders and was a random sample.
+2.
+**No Perfect Colinearity:** Running a VIF in R in the cell below, we see the VIF score is relatively low for both predictors.
+Conceptually, we can see no perfect colinearity as we cannot perfectly describe the views variable in terms of average rating or vice versa.
+Additionally, from the last scatter plot above, we don't see a perfect linear relationship between the predictor variabls views or average rating.
+3.
+**Linear Conditional Expectation:** 'For this assumption, we are checking the residuals for linear relationships.
+We can view this by plotting the residuals, as seen below.
+We generally see a linear line as we move across predictor values with the residuals with no strong pattern.
+By transforming the average rating and views (log of), we are able to see this linear relationship across all predictor values.
+4.
+**Homoskedastic Errors:** 'We are looking for constant error variance across the entire range of the x's ( homoskedastic errors across the range of the x's).
+From the residual plot below for the final model ('Final Model' figure), we can see the even or equal spread dispersion across the x values whereas with the other two models, this is not as consistent.
+The residual plots for the other two show increases with the value of the fitted  variables, suggesting non-constant variances in the residuals errors (or heteroscedasticity).
+Therefore our final model does meet this condition.
+We are able to achieve this fianl plot through transfomrtion of views and average rating (log).
+5.
+**Normally Distributed Errors:** We are looking for a normal distribution of the residuals.
+We use the QQ plot to investigate this assumption.
+For the final model, we see a nice linear line from the QQ plot whereas the other two previou models do not show this.
+We are able to achieve this fianl plot through transfomrtion of views and average rating (log).
+In conclusion, we meet this assumption with the final model ("Final Model" plot, model = model_transformviewsrating).
 
@@
 par(mfrow = c(2, 2))
-plot(model2)
+plot(model2, col="blue") 
+title("Model 2", line = -0.75, outer = TRUE)
 
@@
 par(mfrow = c(2, 2))
-plot(model3)
+plot(model3, col="red") 
+title("Model 2", line = -0.75, outer = TRUE)
 

### 2021-08-04 reports/Draft_Report.Rmd
updated CLM section again

@@
 
-```{r EDA}
+```{r EDA, echo=FALSE, message=FALSE}
 #Example Table
@@
 ```{r Accounting Table creation & output}
-
 df2 <- df
@@
 df2 <- df2 %>% filter(!is.na(df2))
+
 #df2 <- df2 %>% filter(!is.infinite(df2))
@@
 
-#knitr::kable(accounting_df)
 colnames(accounting_df) <- c('Step','Number of Samples','Samples Removed', 'Reason')
@@
 #need to change these to per capita data
-model1 <- lm(log(long_trip+1) ~ log(cases+1) + log(deaths+1) + is_rural + median_income, data = df) #limited model
-model2 <- lm(log(long_trip+1) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unempoy_rate + log(density+1), data = df)
-model3 <- lm(log(long_trip+1) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unempoy_rate + log(density+1) + pct_lt17, data = df) #final model
+model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income, data = df) #limited model
+model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unempoy_rate + log(density), data = df)
+model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unempoy_rate + log(density) + pct_lt17 + pct_gt65, data = df) #final model
 
@@
 
-#Run coeftest to show statistical significance of all variables in the model
+#Run coeftest to show statistical significance of all variables in the model with robust standard errors
 coeftest(model3, vcov = vcovHC(model3, type="HC1"))
+
 ```
@@
 ## Discussion
+
 ### Model Limitations
@@
 **IID Sampling:** There isn't a direct database check for this unfortunately.
-Looking into how this data was generated, I belive this is IID generally.
-The data was gathered from independent YouTube uploaders and was a random sample.
+Looking into how the data was generated, we believe this is IID generally.
+The data was gathered from independent surveys within the Texas population for all models generated.
+
 2.
-**No Perfect Colinearity:** Running a VIF in R in the cell below, we see the VIF score is relatively low for both predictors.
-Conceptually, we can see no perfect colinearity as we cannot perfectly describe the views variable in terms of average rating or vice versa.
-Additionally, from the last scatter plot above, we don't see a perfect linear relationship between the predictor variabls views or average rating.
+**No Perfect Colinearity:** This can be assessed through the variance inflation factor (VIF).
+Running a VIF in R in the cell below, we see the VIF score is relatively low for most predictors excluding the transformed Population Density variable.
+Additionally, when we view the relationship
+
+```{r VIF & correlation}
+
+#create vector of VIF values
+vif_values2 <- car::vif(model2)
+vif_values2
+vif_values3 <- car::vif(model3)
+vif_values3
+
+#Put values into nicer format table
+```
+
 3.
 **Linear Conditional Expectation:** 'For this assumption, we are checking the residuals for linear relationships.
-We can view this by plotting the residuals, as seen below.
+We can view this by plotting the residuals, as seen below in the first scatter plot for both model 2 and model 3 (Residual vs Fitted).
 We generally see a linear line as we move across predictor values with the residuals with no strong pattern.
 By transforming the average rating and views (log of), we are able to see this linear relationship across all predictor values.
+
 4.
@@
 We are able to achieve this fianl plot through transfomrtion of views and average rating (log).
+
 5.
@@
 par(mfrow = c(2, 2))
-plot(model3, col="red") 
-title("Model 2", line = -0.75, outer = TRUE)
+plot(model3, col="green") 
+title("Model 3", line = -0.75, outer = TRUE)
 
@@
 
-###Omitted Variable
+### Omitted Variable
 
@@
 While the Govenor of Texas issued detailed protocols or mandates for the state of Texas, the degree to which each county (and thus the individual residents) complied would be useful data to have if such data were to exist.
+***Add verbiage around including vaccine rates for next steps
 

### 2021-08-04 reports/Draft_Report.Rmd
updated CLM and modeling sections

@@
 title: "Section_8_Team1_Lab2_Report"
-author: "Degnan, Rains, Wu"
-date: "8/1/2021"
+author: "Jonas Degnan, Autumn Rains, Lucy Wu"
+date: "8/5/2021"
 output:
@@
 
-```{r model creation, warning=FALSE}
+To investigate the causal relationships between distance traveled and variables within the dataset, a linear regression was performed in R Studio for three models.
+The first model is limited to study only key variables of interest for measurment.
+The second model expands upon the first model to also include key explanatory variables and covariates that are hypothesized to improve performance of the previous model.
+The third model is all-inclusive of the variables available to the team at the time of data gathering.
+Details of each model analyzed and subsequent results can be found below.
+
+$$
+  f(\text{Distance Traveled}) = \beta_0 + \beta_1 g(\text{rate})  + \beta_3 h(\text{length})
+$$
+
+<!-- Limited Model: The first model you include should include only the key variables you want to measure.
+These variables might be transformed, as determined by your EDA, but the model should include the absolute minimum number of covariates (perhaps one, or at most two-three, covariates if they are so crucial that it would be unreasonable to omit them).
+-->
+<!-- Model Two: One model that includes key explanatory variables and covariates that you believe advance your modeling goals without introducing too much multicollinearity or causing other issues of confounding.
+This model should strike a balance between accuracy and parsimony and reflect your best understanding of the relationships among key variables.
+-->
+<!-- Model Three: One model that includes the previous covariates, and many other covariates, erring on the side of inclusion.
+A key purpose of this model is to evaluate how parameters of interest change (if at all) when additional, potentially colinear variables are included in the model specification.
+-->
+
+```{r model creation, echo=FALSE, warning=FALSE}
 
@@
 Looking into how the data was generated, we believe this is IID generally.
-The data was gathered from independent surveys within the Texas population for all models generated.
+The data was gathered from independent surveys within the Texas population for all models generated across all counties.
 
 2.
-**No Perfect Colinearity:** This can be assessed through the variance inflation factor (VIF).
-Running a VIF in R in the cell below, we see the VIF score is relatively low for most predictors excluding the transformed Population Density variable.
-Additionally, when we view the relationship
+**No Perfect Colinearity:** This can be assessed through examination of the variance inflation factor (VIF).
+Running a VIF in R for model 2 and model 3, we see the VIF scores are relatively low (below 5) for the predictor variables.
+Therefore, no predictor variables were removed from the model.
+The results for the VIF analysis are seen in the table below:
 
@@
 3.
-**Linear Conditional Expectation:** 'For this assumption, we are checking the residuals for linear relationships.
-We can view this by plotting the residuals, as seen below in the first scatter plot for both model 2 and model 3 (Residual vs Fitted).
+**Linear Conditional Expectation:** For this assumption, the residuals for linear relationships must be evaluated.
+By plotting the residuals, as seen below in the first scatter plot for both model 2 and model 3 ('Residual vs Fitted').
 We generally see a linear line as we move across predictor values with the residuals with no strong pattern.
-By transforming the average rating and views (log of), we are able to see this linear relationship across all predictor values.
+By transforming the variables as indicated previously logarithmically, we are able to see this linear relationship across all predictor values.
 
 4.
-**Homoskedastic Errors:** 'We are looking for constant error variance across the entire range of the x's ( homoskedastic errors across the range of the x's).
-From the residual plot below for the final model ('Final Model' figure), we can see the even or equal spread dispersion across the x values whereas with the other two models, this is not as consistent.
-The residual plots for the other two show increases with the value of the fitted  variables, suggesting non-constant variances in the residuals errors (or heteroscedasticity).
-Therefore our final model does meet this condition.
-We are able to achieve this fianl plot through transfomrtion of views and average rating (log).
+**Homoskedastic Errors:** This assumption is examined by looking for constant error variance across the entire range of the x's (homoskedastic errors across the range of the x's).
+By plotting fitted values vs.
+the square root of the standardized residuals for the model 2 and model 3 ('Scale-Location' figure), there is an even or equal spread dispersion across the x values.
+Therefore model 2 and model both meet this condition when including the transformed variables.
 
 5.
-**Normally Distributed Errors:** We are looking for a normal distribution of the residuals.
-We use the QQ plot to investigate this assumption.
-For the final model, we see a nice linear line from the QQ plot whereas the other two previou models do not show this.
-We are able to achieve this fianl plot through transfomrtion of views and average rating (log).
-In conclusion, we meet this assumption with the final model ("Final Model" plot, model = model_transformviewsrating).
+**Normally Distributed Errors:** When investigating a normal distribution of the residuals, the QQ plot is helpful to investigate this assumption.
+For model 2 and model 3, a nice linear line can be seen between the theoretical quantities and the standardized residuals ('Normal Q-Q').
+In conclusion, we meet this assumption with both model 2 and model 3.
 
@@
 
-Based upon the linear regression analysis above, it is observed that there may be a causal relationship between XX and YY.
-The hypothesis posed within the introduction is 'failed to be rejected/ rejected.' ....
+The goal of this analysis was to detrmine if there existed a causal relationship between distance traveled and Covid-19 infection rates for residents of the state of Texas.
+Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model.
 
@@
 While the Govenor of Texas issued detailed protocols or mandates for the state of Texas, the degree to which each county (and thus the individual residents) complied would be useful data to have if such data were to exist.
-***Add verbiage around including vaccine rates for next steps
+Additionally, incorporating vaccination rates could aid in development of a model with even higher accuracy.
+Generally, vaccinations could increase confidence among individuals interested in traveling, especially if they have been in long periods of quarantine or government mandated lockdowns.
+Last, it is highly recommended that any model generated in the future continue to use the latest data available given the oscillation in vaccination rates, government policy, and new or known variants of the Covid-19 virus.
 

### 2021-08-04 src/data/DataCleaning.Rmd

@@
     county = County,
-    unempoy_rate = Unemployment.Rate....
+    unemp_rate = Unemployment.Rate....
   ) %>%
@@
     county, 
-    unempoy_rate 
+    unemp_rate 
   )
@@
 # Write to .csv
-```{r export interim model_1 dataframe as csv, echo=FALSE}
+```{r export interim model_2 dataframe as csv, echo=FALSE}
 write.csv(d, '../../data/interim/covid_trips_interim_model2.csv')
@@
 # Write to .csv
-```{r export interim model_1 dataframe as csv, echo=FALSE}
+```{r export interim model_3 dataframe as csv, echo=FALSE}
 write.csv(d, '../../data/interim/covid_trips_interim_model3.csv')
+```
+
+```{r clean unemployment data}
+d <- d %>%
+  select(
+    county,
+    long_trip,
+    cases,
+    deaths,
+    median_income,
+    is_rural,
+    unemp_rate,
+    density,
+    pct_lt17,
+    pct_gt65,
+    povery_rate
+  )
+knitr::kable(head(d))
+```
+
+# Write to .csv
+```{r export processed dataframe as csv, echo=FALSE}
+write.csv(d, '../../data/processed/covid_trips_processed.csv')
 ```
@@
 # remove dataframes
-rm(d_covid, d_trips, d_openings)
-rm(dp03_econ, dp05_age, density, s1901_income, urb_rur_h2, econ, income, age)
+rm(list = ls())
 

### 2021-08-04 notebooks/Draft_Report.Rmd
moved Rmd to /notebooks and set pdf report output to /report. set path to processed csv.

--- 2021-08-04 reports/Draft_Report.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
-
+knitr::knit(output = "./../reports/final_report.pdf")
 ```
@@
 ```{r dataframe, echo=FALSE}
-df <- read.csv('~/lab-2-lab2_section8_jal/data/interim/covid_trips_interim_model3.csv')
+df <- read.csv('./../data/processed/covid_trips_processed.csv')
 #head(df)
@@
 table1::label(df$cases) <- "County Covid-19 Cases"
-table1::label(df$unempoy_rate) <- "County Unemployment Rate"
+table1::label(df$unemp_rate) <- "County Unemployment Rate"
 table1::label(df$median_income) <- "County Median Income ($)"
@@
 # table1::label(df$is_rural) <- "County Type"
-table1::table1(~total_pop + density + cases + unempoy_rate + median_income + pct_lt17 + pct_gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Statistics", topclass="Rtable1-zebra", overall = "Total")
+table1::table1(~total_pop + density + cases + unemp_rate + median_income + pct_lt17 + pct_gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Statistics", topclass="Rtable1-zebra", overall = "Total")
 
@@
   ) %>%
-  select(long_trip, cases, deaths, median_income, unempoy_rate, density, pct_lt17) %>%
+  select(long_trip, cases, deaths, median_income, unemp, density, pct_lt17) %>%
   GGally::ggpairs(.,
@@
 model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income, data = df) #limited model
-model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unempoy_rate + log(density), data = df)
-model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unempoy_rate + log(density) + pct_lt17 + pct_gt65, data = df) #final model
+model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unemp + log(density), data = df)
+model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unemp + log(density) + pct_lt17 + pct_gt65, data = df) #final model
 

### 2021-08-04 notebooks/Draft_Report.Rmd
fixed typo.

@@
   ) %>%
-  select(long_trip, cases, deaths, median_income, unemp, density, pct_lt17) %>%
+  select(long_trip, cases, deaths, median_income, unemp_rate, density, pct_lt17) %>%
   GGally::ggpairs(.,
@@
 model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income, data = df) #limited model
-model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unemp + log(density), data = df)
-model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unemp + log(density) + pct_lt17 + pct_gt65, data = df) #final model
+model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unemp_rate + log(density), data = df)
+model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unemp_rate + log(density) + pct_lt17 + pct_gt65, data = df) #final model
 

### 2021-08-04 notebooks/Draft_Report.Rmd
cleaned total_pop from report.

@@
 
-```{r setup, message=FALSE}
+```{r install packages, message=FALSE}
 install.packages("table1")
@@
 install.packages("stargazer")
+```
 
+``` {r load packages, echo=FALSE}
 #we may not need all of these
@@
 knitr::opts_chunk$set(echo = TRUE)
-knitr::knit(output = "./../reports/final_report.pdf")
+#knitr::knit(input = "./../notebooks/Draft_Report.Rmd", output = "./../reports/final_report.pdf")
 ```
@@
 df <- read.csv('./../data/processed/covid_trips_processed.csv')
-#head(df)
 ```
@@
 df$rural_cat <- ifelse(df$is_rural == 1, "Rural Counties", "Urban Counties")
-table1::label(df$total_pop) <- "County Population"
 table1::label(df$density) <- "County Population Density"
@@
 # table1::label(df$is_rural) <- "County Type"
-table1::table1(~total_pop + density + cases + unemp_rate + median_income + pct_lt17 + pct_gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Statistics", topclass="Rtable1-zebra", overall = "Total")
+table1::table1(~density + cases + unemp_rate + median_income + pct_lt17 + pct_gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Statistics", topclass="Rtable1-zebra", overall = "Total")
 
@@
   mutate(
-    total_pop = log(total_pop),
     cases = log(cases+1),

### 2021-08-04 notebooks/Draft_Report.Rmd
updated modeling text

@@
 
-## Linear Regression Results
-
 To investigate the causal relationships between distance traveled and variables within the dataset, a linear regression was performed in R Studio for three models.
-The first model is limited to study only key variables of interest for measurment.
+The first model is limited to study only key variables of interest for measurement.
 The second model expands upon the first model to also include key explanatory variables and covariates that are hypothesized to improve performance of the previous model.
@@
 $$
-  f(\text{Distance Traveled}) = \beta_0 + \beta_1 g(\text{rate})  + \beta_3 h(\text{length})
+  Model 1 = f(\text{log(Distance Traveled)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})
 $$ 
-
+$$
+  Model 2 = f(\text{log(Distance Traveled)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density})
+$$
+$$
+  Model 3 = f(\text{log(Distance Traveled)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})
+  + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density}) + \beta_6 h(\text{log(% Age < 17}) + \beta_7 h(\text{log(% Age < 17})
+$$ 
 <!-- Limited Model: The first model you include should include only the key variables you want to measure.
@@
 
+## Linear Regression Results
 ```{r model creation, echo=FALSE, warning=FALSE}

### 2021-08-04 notebooks/Draft_Report.Rmd
updateg text

--- 2021-08-04 notebooks/Draft_Report.Rmd
@@
 
+-->Need to explain distance of travel with
+
 ## Operationalization
@@
 #Example Table
-df$rural_cat <- ifelse(df$is_rural == 1, "Rural Counties", "Urban Counties")
+df$rural_cat <- ifelse(df$is_rural == 1, "Rural", "Urban")
 table1::label(df$density) <- "County Population Density"
@@
 
-#Short vs long trip: rate of short trips is higher than rate of long trips in relation to # of covid cases
-ggplot(data = df) + aes(x = log(cases), y = log(long_trip)) + geom_point()
-#ggplot(data = df) + aes(x = log(cases), y = log(short_trip)) + geom_point()
-
-Rural_County=as.factor(df$is_rural)
+County <= as.factor(df$rural_cat)
 
@@
     long_trip = log(long_trip),
-    density = log(density)
+    density = log(density),
+    median_income=log(median_income+1),
+    unemp_rate = log(unemp_rate+1)
   ) %>%
-  select(long_trip, cases, deaths, median_income, unemp_rate, density, pct_lt17) %>%
+  select(long_trip, cases, deaths, median_income, unemp_rate, density, pct_lt17, pct_gt65) %>%
   GGally::ggpairs(., 
-               legend = 1, title = "Figure 1: Variable Statistics",
-               mapping = ggplot2::aes(colour=Rural_County))
+               legend = 1,
+               aes(color = County), 
+               title = "Figure 1: Variable Characterstics",
+               upper = list(continuous = wrap("cor", size = 2.5)),
+               lower = list(continuous = wrap("points", alpha = 0.5, size=0.3)),
+               diag = list(continuous = wrap("densityDiag", alpha = 0.5)),
+               columnLabels=c("Distance", "Cases", "Deaths", "Income", "Unemployment", "Pop Density", "Age < 17","Age > 65"))+
+               theme(legend.position = "bottom")
 
@@
 
-df2$log_pop <- log10(df2$total_pop+1)
 df2$log_cases <-log10(df2$cases+1)
@@
 
+To investigate the causal relationships between distance traveled and variables within the dataset, a linear regression was performed in R Studio for three models.
+The first model is limited to study only key variables of interest for measurement.
+The second model expands upon the first model to also include key explanatory variables and covariates that are hypothesized to improve performance of the previous model.
+The third model is all-inclusive of the variables available to the team at the time of data gathering.
+Details of each model analyzed can be found below.
+
+$$
+  Model 1 = f(\text{log(Distance Traveled)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})
+$$ 
+$$
+  Model 2 = f(\text{log(Distance Traveled)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density})
+$$
+$$
+  Model 3 = f(\text{log(Distance Traveled)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})
+  + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density}) + \beta_6 h(\text{log(% Age < 17}) + \beta_7 h(\text{log(% Age < 17})
+$$
+
 ## Linear Regression Results
 
-To investigate the causal relationships between distance traveled and variables within the dataset, a linear regression was performed in R Studio for three models.
-The first model is limited to study only key variables of interest for measurment.
-The second model expands upon the first model to also include key explanatory variables and covariates that are hypothesized to improve performance of the previous model.
-The third model is all-inclusive of the variables available to the team at the time of data gathering.
-Details of each model analyzed and subsequent results can be found below.
-
-$$
-  f(\text{Distance Traveled}) = \beta_0 + \beta_1 g(\text{rate})  + \beta_3 h(\text{length})
-$$
-
-<!-- Limited Model: The first model you include should include only the key variables you want to measure.
-These variables might be transformed, as determined by your EDA, but the model should include the absolute minimum number of covariates (perhaps one, or at most two-three, covariates if they are so crucial that it would be unreasonable to omit them).
--->
-<!-- Model Two: One model that includes key explanatory variables and covariates that you believe advance your modeling goals without introducing too much multicollinearity or causing other issues of confounding.
-This model should strike a balance between accuracy and parsimony and reflect your best understanding of the relationships among key variables.
--->
-<!-- Model Three: One model that includes the previous covariates, and many other covariates, erring on the side of inclusion.
-A key purpose of this model is to evaluate how parameters of interest change (if at all) when additional, potentially colinear variables are included in the model specification.
--->
-
+The results from the linear regression analysis can be seen below for each model: 
 ```{r model creation, echo=FALSE, warning=FALSE}
@@
 #need to change these to per capita data
-model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income, data = df) #limited model
+model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1), data = df) #limited model
 model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unemp_rate + log(density), data = df)
 model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unemp_rate + log(density) + pct_lt17 + pct_gt65, data = df) #final model
+stargazer(model1, model2, model3, type='text', title="Results", align=TRUE)
+```
+Additionally, to indicate if any of the models provide a better fit to the dataset than a model that contains no independent variables, an F test was performed.
+The results for each model are shown below:
+```{r F Test}
+#Run F test to show statistical significance of additional variables
+anova(model1, model2, model3, test = "F")
+```
+From the F-test,  we can see the the addition of the variables into model 2 and model 3 do contribute signficant information to our model based off of the derived p-values.
 
-stargazer(model1, model2, model3, type='text', title="Results", align=TRUE)
-
-#Run F test to show statistical significance of additional variables
-anova(model1, model2, test = "F")
-anova(model2, model3, test = "F")
-
+Last, the significance of individual regression coefficients was tested using robust standard errors in R studio for model 3. The results are shown below: 
+```{r coeftest}
 #Run coeftest to show statistical significance of all variables in the model with robust standard errors
 coeftest(model3, vcov = vcovHC(model3, type="HC1"))
-
 ```
+From the t-test, we can see that Covid-19 cases and median income are statistically significant.
+Additionally, the percentage of adults greather than 65, the population density, and the Covid-19 death variables are even more statistically significant.
 
@@
 
-<!-- If the team has taken up an explanatory (i.e.
-causal) question to evaluate, then identify what you think are the most important omitted variables that bias results you care about.
-For each variable you name, you should reason about the direction of bias caused by omitting this variable.
-If you can argue whether the bias is large or small, that is even better.
-State whether you have any variables available that may proxy (even imperfectly) for the omitted variable.
-Pay particular attention to whether each omitted variable bias is towards zero or away from zero.
-You will use this information to judge whether the effects you find are likely to be real, or whether they might be entirely an artifact of omitted variable bias.
--->
-
 #### Variable 1
@@
 
-The goal of this analysis was to detrmine if there existed a causal relationship between distance traveled and Covid-19 infection rates for residents of the state of Texas.
+The goal of this analysis was to determine if there existed a causal relationship between distance traveled and Covid-19 infection rates for residents of the state of Texas.
 Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model.

### 2021-08-05 src/data/DataCleaning.Rmd

@@
     median_income,
+    total_pop,
     is_rural,

### 2021-08-05 notebooks/Draft_Report.Rmd
updated text

@@
 
-Do COVID infection rates impact vacation travel distance for U.S. residents in the state of Texas? 
+Do COVID-19 infection rates impact vacation travel distance for U.S. residents in the state of Texas? 
 How does income, population density, and covid regulations in each county impact traveling?
@@
 
--->Need to explain distance of travel with
+-->Need to explain distance of travel with bs TRANSPORT data
+-->Need to add more to causal theory to justify variable selection
 
@@
 table1::table1(~density + cases + unemp_rate + median_income + pct_lt17 + pct_gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Statistics", topclass="Rtable1-zebra", overall = "Total")
+```
+Below are histograms of variables that did not follow normal distribution and required transformation. Each of these variables was logarithmically transformed.
+```{r scatterplots, message=FALSE}
+attach(df)
+par(mfrow=c(3,2))
+hist(cases, main="Histogram of Covid-19 Cases")
+hist(deaths, main="Histogram of Covid-19 Deaths")
+hist(long_trip, main="Histogram of Distance")
+hist(density, main="Histogram of Population Density")
+hist(unemp_rate, main="Histogram of Unemployment Rate")
+hist(median_income, main="Histogram of Median Income")
 
-County <= as.factor(df$rural_cat)
+```
 
-df %>%
-  mutate(
-    cases = log(cases+1),
-    deaths = log(deaths+1),
-    long_trip = log(long_trip),
-    density = log(density),
-    median_income=log(median_income+1),
-    unemp_rate = log(unemp_rate+1)
-  ) %>%
-  select(long_trip, cases, deaths, median_income, unemp_rate, density, pct_lt17, pct_gt65) %>%
-  GGally::ggpairs(., 
-               legend = 1,
-               aes(color = County), 
-               title = "Figure 1: Variable Characterstics",
-               upper = list(continuous = wrap("cor", size = 2.5)),
-               lower = list(continuous = wrap("points", alpha = 0.5, size=0.3)),
-               diag = list(continuous = wrap("densityDiag", alpha = 0.5)),
-               columnLabels=c("Distance", "Cases", "Deaths", "Income", "Unemployment", "Pop Density", "Age < 17","Age > 65"))+
-               theme(legend.position = "bottom")
+```{r ggally, echo=FALSE, message=FALSE}
+County = as.factor(df$rural_cat)
+
+# df %>%
+#   mutate(
+#     cases = log(cases+1),
+#     deaths = log(deaths+1),
+#     long_trip = log(long_trip),
+#     density = log(density),
+#     median_income=log(median_income+1),
+#     unemp_rate = log(unemp_rate+1)
+#   ) %>%
+#   select(long_trip, cases, deaths, median_income, unemp_rate, density, pct_lt17, pct_gt65) %>%
+#   GGally::ggpairs(.,
+#                legend = 1,
+#                aes(color = County),
+#                title = "Figure 1: Variable Characterstics",
+#                upper = list(continuous = wrap("cor", size = 2.5)),
+#                lower = list(continuous = wrap("points", alpha = 0.5, size=0.3)),
+#                diag = list(continuous = wrap("densityDiag", alpha = 0.5)),
+#                columnLabels=c("Distance", "Cases", "Deaths", "Income", "Unemployment", "Pop Density", "Age < 17","Age > 65"))+
+#                theme(legend.position = "bottom")
+
+ggpairs(df, columns = 3:ncol(df), legend = 1, aes(color = County), title = "Figure 1: Variable Characterstics")
+#                upper = list(continuous = wrap("cor", size = 2.5)),
+#                lower = list(continuous = wrap("points", alpha = 0.5, size=0.3)),
+#                diag = list(continuous = wrap("densityDiag", alpha = 0.5)))
 
@@
 $$
+
 $$
@@
 #Run coeftest to show statistical significance of all variables in the model with robust standard errors
-coeftest(model3, vcov = vcovHC(model3, type="HC1"))
+coeftest(model2, vcov = vcovHC(model3, type="HC1"))
 ```
@@
 
-##FINAL MODEL
-par(mfrow = c(2, 2))
-plot(model3, col="green") 
-title("Model 3", line = -0.75, outer = TRUE)
+# Model 3
+# par(mfrow = c(2, 2))
+# plot(model3, col="green") 
+# title("Model 3", line = -0.75, outer = TRUE)
 

### 2021-08-05 notebooks/Draft_Report.Rmd
adding table sources and latest draft

@@
 
-We will analyze datasets pertaining to these variables for Texas counties and create three models to understand if causal relationships exist between distance traveled, population income, population density (rural or urban counties), and government regulations.
+We will analyze datasets pertaining to these variables for Texas counties and create three models to understand if causal relationships exist between the number of long distance trips, population income, population density (rural or urban counties).
 Our hypothesis is that an there is a causal relationship between """
@@
 df <- read.csv('./../data/processed/covid_trips_processed.csv')
+head(df)
 ```
@@
 
-# df %>%
-#   mutate(
-#     cases = log(cases+1),
-#     deaths = log(deaths+1),
-#     long_trip = log(long_trip),
-#     density = log(density),
-#     median_income=log(median_income+1),
-#     unemp_rate = log(unemp_rate+1)
-#   ) %>%
-#   select(long_trip, cases, deaths, median_income, unemp_rate, density, pct_lt17, pct_gt65) %>%
-#   GGally::ggpairs(.,
-#                legend = 1,
-#                aes(color = County),
-#                title = "Figure 1: Variable Characterstics",
-#                upper = list(continuous = wrap("cor", size = 2.5)),
-#                lower = list(continuous = wrap("points", alpha = 0.5, size=0.3)),
-#                diag = list(continuous = wrap("densityDiag", alpha = 0.5)),
-#                columnLabels=c("Distance", "Cases", "Deaths", "Income", "Unemployment", "Pop Density", "Age < 17","Age > 65"))+
-#                theme(legend.position = "bottom")
+df %>%
+  mutate(
+    cases = log(cases+1),
+    deaths = log(deaths+1),
+    long_trip = log(long_trip),
+    density = log(density),
+    median_income=log(median_income+1),
+    unemp_rate = log(unemp_rate+1)
+  ) %>%
+  select(long_trip, cases, deaths, median_income, unemp_rate, density, pct_lt17, pct_gt65) %>%
+  GGally::ggpairs(.,
+               legend = 1,
+               aes(color = County),
+               title = "Figure 1: Variable Characterstics",
+               upper = list(continuous = wrap("cor", size = 2.5)),
+               lower = list(continuous = wrap("points", alpha = 0.5, size=0.3)),
+               diag = list(continuous = wrap("densityDiag", alpha = 0.5)),
+               columnLabels=c("Distance", "Cases", "Deaths", "Income", "Unemployment", "Pop Density", "Age < 17","Age > 65"))+
+               theme(legend.position = "bottom")
 
-ggpairs(df, columns = 3:ncol(df), legend = 1, aes(color = County), title = "Figure 1: Variable Characterstics")
-#                upper = list(continuous = wrap("cor", size = 2.5)),
-#                lower = list(continuous = wrap("points", alpha = 0.5, size=0.3)),
-#                diag = list(continuous = wrap("densityDiag", alpha = 0.5)))
+# ggpairs(df, columns = 3:ncol(df), legend = 1, aes(color = County), title = "Figure 1: Variable Characterstics")
+# #                upper = list(continuous = wrap("cor", size = 2.5)),
+# #                lower = list(continuous = wrap("points", alpha = 0.5, size=0.3)),
+# #                diag = list(continuous = wrap("densityDiag", alpha = 0.5)))
 
@@
 $$
-  Model 1 = f(\text{log(Distance Traveled)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})
+  Model 1 = f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})
 $$
@@
 $$
-  Model 2 = f(\text{log(Distance Traveled)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density})
+  Model 2 = f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density})
 $$
@@
 $$
-  Model 3 = f(\text{log(Distance Traveled)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})
+  Model 3 = f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})
   + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density}) + \beta_6 h(\text{log(% Age < 17}) + \beta_7 h(\text{log(% Age < 17})

### 2021-08-05 src/data/DataCleaning.Rmd
raw unemployed and in_poverty variables mutated

--- 2021-08-04 src/data/DataCleaning.Rmd
@@
 d <- merge(x = income_f, y = d, by = c("county"))
+d <- d %>% rowwise() %>%
+  mutate(
+    in_poverty = as.integer(0.01 * povery_rate * total_pop)
+  )
 d <- merge(x = density_f, y = d, by = c("county"))
@@
 d <- merge(x = unemploy_f, y = d, by = c("county"))
+d <- d %>% rowwise() %>%
+  mutate(
+    unemployed = as.integer(0.01 * unemp_rate * total_pop)
+  )
 d <- d %>% drop_na()
@@
     median_income,
+    total_pop,
     is_rural,
-    unemp_rate,
+    unemployed,
     density,
@@
     pct_gt65,
-    povery_rate
+    in_poverty
   )

### 2021-08-05 notebooks/Draft_Report.Rmd
report updates and modeling

@@
 Do COVID-19 infection rates impact vacation travel distance for U.S. residents in the state of Texas? 
-How does income, population density, and covid regulations in each county impact traveling?
+Does income, population density, unemployment rates, and age in Texas counties impact traveling?
 
-We will analyze datasets pertaining to these variables for Texas counties and create three models to understand if causal relationships exist between the number of long distance trips, population income, population density (rural or urban counties).
+We will analyze datasets from June, 2021 pertaining to these variables for Texas counties and create three models to understand if causal relationships exist between the number of long distance trips, population income, population density (rural or urban counties).
 Our hypothesis is that an there is a causal relationship between """
@@
 
-To create our dataframe, we cleaned and filtered data from multiple sources of data and joined based on texas county name or code.
+To create our dataframe, we cleaned and filtered data from multiple sources of data and joined based on texas county name or code from June, 2021.
 
@@
   ) %>%
-  select(long_trip, cases, deaths, median_income, unemp_rate, density, pct_lt17, pct_gt65) %>%
+  select(long_trip, cases, deaths, median_income, povery_rate, density, pct_lt17, pct_gt65) %>%
   GGally::ggpairs(.,
@@
 To investigate the causal relationships between distance traveled and variables within the dataset, a linear regression was performed in R Studio for three models.
-The first model is limited to study only key variables of interest for measurement.
-The second model expands upon the first model to also include key explanatory variables and covariates that are hypothesized to improve performance of the previous model.
 The third model is all-inclusive of the variables available to the team at the time of data gathering.
@@
 
+Here is the first model, which is limited to study only key variables of interest for measurement based on our hypotheses.
+Our key input variable is covid cases and key output is number of trips greater than 100 miles.
+We also chose to add an indicator for rural counties because it could be directly related to trip distance, as well as median income which plays a role in a person's ability to take vacations in the first place.
 $$
@@
 $$ 
-
+The second model expands upon the first model to also include key explanatory variables and covariates that are hypothesized to improve performance of the previous model.
+Here, we added in unemployment rate to assess if the impact of covid to that county affected travel.
 $$
@@
 $$
-
+Finally in the third model, we added in age demographics as percent under 17/over 65 as another potential source that influences travel midst covid cases due to the nature of how covid affects those at different ages.
 $$
@@
 model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1), data = df) #limited model
-model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unemp_rate + log(density), data = df)
-model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unemp_rate + log(density) + pct_lt17 + pct_gt65, data = df) #final model
+model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1) + povery_rate+ log(density), data = df)
+model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1) + povery_rate+ log(density) + pct_lt17 + pct_gt65, data = df) #final model
 stargazer(model1, model2, model3, type='text', title="Results", align=TRUE)
 ```
+From this output table, we can see most of our hypotheses about the variables we included in our model were correct and do have a statistically significant contribution to our model already.
+
 Additionally, to indicate if any of the models provide a better fit to the dataset than a model that contains no independent variables, an F test was performed.
@@
 ```
-From the F-test,  we can see the the addition of the variables into model 2 and model 3 do contribute signficant information to our model based off of the derived p-values.
+From the F-test,  we can see the the addition of the variables into model 2 and model 3 do contribute significant information to our model based off of the derived p-values.
 
@@
 #Run coeftest to show statistical significance of all variables in the model with robust standard errors
-coeftest(model2, vcov = vcovHC(model3, type="HC1"))
+coeftest(model3, vcov = vcovHC(model3, type="HC1"))
 ```
 From the t-test, we can see that Covid-19 cases and median income are statistically significant.
-Additionally, the percentage of adults greather than 65, the population density, and the Covid-19 death variables are even more statistically significant.
+Additionally, the percentage of adults greater than 65, the population density, and the Covid-19 death variables are even more statistically significant.
 

### 2021-08-05 notebooks/Draft_Report.Rmd
model explanation

@@
 table1::label(df$cases) <- "County Covid-19 Cases"
-table1::label(df$unemp_rate) <- "County Unemployment Rate"
+table1::label(df$unemployed) <- "County Unemployment Rate"
 table1::label(df$median_income) <- "County Median Income ($)"
-table1::label(df$pct_lt17) <- "County Age < 17 years (%)"
-table1::label(df$pct_gt65) <- "County Age > 65 years (%)"
+table1::label(df$lt17) <- "County Age < 17 years"
+table1::label(df$gt65) <- "County Age > 65 years"
 # table1::label(df$is_rural) <- "County Type"
-table1::table1(~density + cases + unemp_rate + median_income + pct_lt17 + pct_gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Statistics", topclass="Rtable1-zebra", overall = "Total")
+table1::table1(~density + cases + unemployed + median_income + lt17 + gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Statistics", topclass="Rtable1-zebra", overall = "Total")
 ```
@@
 hist(density, main="Histogram of Population Density")
-hist(unemp_rate, main="Histogram of Unemployment Rate")
+hist(unemployed, main="Histogram of Unemployment")
 hist(median_income, main="Histogram of Median Income")
@@
     median_income=log(median_income+1),
-    unemp_rate = log(unemp_rate+1)
+    unemployed = log(unemployed+1),
+    lt17 = log(lt17),
+    gt65 = log(gt65)
   ) %>%
-  select(long_trip, cases, deaths, median_income, povery_rate, density, pct_lt17, pct_gt65) %>%
+  select(long_trip, cases, deaths, median_income, unemployed, density, lt17, gt65) %>%
   GGally::ggpairs(.,
@@
 #need to change these to per capita data
-model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1), data = df) #limited model
-model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1) + povery_rate+ log(density), data = df)
-model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1) + povery_rate+ log(density) + pct_lt17 + pct_gt65, data = df) #final model
+model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural, data = df) #limited model
+model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1) + log(density), data = df)
+model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1) + log(density) + log(lt17) + log(gt65), data = df) #final model
 stargazer(model1, model2, model3, type='text', title="Results", align=TRUE)

### 2021-08-05 notebooks/Draft_Report.Rmd
model explanation

@@
 To investigate the causal relationships between distance traveled and variables within the dataset, a linear regression was performed in R Studio for three models.
-The third model is all-inclusive of the variables available to the team at the time of data gathering.
-Details of each model analyzed can be found below.
+The first model includes only our base variables.
+In the second model, we added additional county characteristics.
+In the final model, we added in another covid related factor.
 
@@
 Our key input variable is covid cases and key output is number of trips greater than 100 miles.
-We also chose to add an indicator for rural counties because it could be directly related to trip distance, as well as median income which plays a role in a person's ability to take vacations in the first place.
+We also chose to add an indicator for rural counties because it could be directly related to trip distance
 $$
-  Model 1 = f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})
+  Model 1 = f(\text{log(#Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)}) + \beta_2 g(\text{log(Covid-19 Deaths)})  + \beta_3 h(\text{Rural County})
 $$ 
 The second model expands upon the first model to also include key explanatory variables and covariates that are hypothesized to improve performance of the previous model.
-Here, we added in unemployment rate to assess if the impact of covid to that county affected travel.
+Here, we added in median income and density of the county as additional descriptors of the county.
 $$
-  Model 2 = f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density})
+  Model 2 = f(\text{log(#Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)}) + \beta_2 g(\text{log(Covid-19 Deaths)})  + \beta_3 h(\text{Rural County}) + \beta_4 h(\text{Median Income}) + \beta_5 h(\text{County Density})
 $$
@@
 $$
-  Model 3 = f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})
-  + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density}) + \beta_6 h(\text{log(% Age < 17}) + \beta_7 h(\text{log(% Age < 17})
+  Model 3 = f(\text{log(#Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)}) + \beta_2 g(\text{log(Covid-19 Deaths)})  + \beta_3 h(\text{Rural County}) + \beta_4 h(\text{Median Income}) + \beta_5 h(\text{County Density}) + \beta_6 h(\text{Population under 17}) + \beta_7 h(\text{Population over 65})
 $$
@@
 ```
-From this output table, we can see most of our hypotheses about the variables we included in our model were correct and do have a statistically significant contribution to our model already.
+From this output table, we can discern quite a bit of information.
+
+First, we consider the models holistically. Here are some of our observations:
+* In model1, all of our variables are extremely significant, as expected from my hypotheses, though in the opposite direction.
+We would expect an inverse relationship between covid and travel, though it seems that this is estimating a positive relationship
+* In model2, when county density is factored in, the presence of covid seems to have less positive of an effect on travel and that higher density counties see more travel. 
+* In model3, with additionally age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17. 
+* In terms of R^2 and RSE, we are improving with each model. 
+* The addition of our chosen parameters for each following model is also statistically better than the previous, as indicated by the F statistic.
+
+We will also interpret the coefficient values and what these mean in the context of our model.
+Almost all of our variables needed a log transform, including the output, making this a log-log regression.
+These regressions are most intuitive to understand in terms of percent increases.
+* First to note is our persistently significant constant in each model.
+Regardless of any variables, we are getting between ~300 (as with model3) and over 8000 (as with model2) trips over 100 miles.
+This is calculated as e^(Constant).
+* In model1, for every percent increase in both cases and deaths, we are getting almost half a percent increase in trips.
+The positive relationship doesnt make much sense, so we will look to the other models.
+* In model2, we see that density has a more positive relationship with trips than do covid cases and deaths. 
+* In model3, covid cases and deaths become insiginifcant, though its lookign
 

### 2021-08-05 notebooks/Draft_Report.Rmd
data description added. pdf formatting.

--- 2021-08-05 notebooks/Draft_Report.Rmd
@@
 output:
+  pdf_document: 
+    toc: yes
+    fig_caption: yes
+    fig_crop: no
+  word_document: default
   html_document: 
-    df_print: paged
-  word_document: default
-  pdf_document: default
+    df_print: kable
+    toc: yes
+    theme: readable
 ---
-
-```{r install packages, message=FALSE}
-install.packages("table1")
-install.packages("GGally")
-install.packages("stargazer")
+\clearpage
+```{r install packages, echo=FALSE, message=FALSE}
+# install.packages("table1")
+# install.packages("GGally")
+# install.packages("stargazer")
+# install.packages("tinytex")
 ```
 
-``` {r load packages, echo=FALSE}
+``` {r load packages, echo=FALSE, message=FALSE}
 #we may not need all of these
@@
 theme_set(theme_minimal())
+options(tinytex.verbose = TRUE)
 knitr::opts_chunk$set(echo = TRUE)
+# tinytex::install_tinytex()
 #knitr::knit(input = "./../notebooks/Draft_Report.Rmd", output = "./../reports/final_report.pdf")
@@
 
-## Introduction
+# Introduction
 
 Summer vacations in the United States are synonymous with sunny, leisurely days spent with loved ones.
-As Ella Fitzgerald so aptly sings, “Summertime, and the living is easy.” However, travel in the recent peak-vacation months has certainly deviated from this idealistic frame of mind with the onset of the COVID-19 pandemic in early 2020.
+As Ella Fitzgerald so aptly sings, “Summertime, and the living is easy.” However, travel in the recent peak-vacation months has certainly deviated from this idealistic frame of mind with the onset of the SARS-CoV-2 outbreak (COVID-19) in early 2020.
 The global travel industry has experienced sharp declines.
-Specifically in the United States, “travel spending totaled a mere $679 billion in 2020, an unprecedented 42% annual decline (nearly $500 billion) from 2019,” according to analysis from the U.S.
+Specifically in the United States (US), “travel spending totaled a mere $679 billion in 2020, an unprecedented 42% annual decline (nearly $500 billion) from 2019,” according to analysis from the U.S.
 Travel Association.
@@
 Do COVID-19 infection rates impact vacation travel distance for U.S. residents in the state of Texas? 
-Does income, population density, unemployment rates, and age in Texas counties impact traveling?
+How does income, population density, and covid regulations in each county impact traveling?
 
-We will analyze datasets from June, 2021 pertaining to these variables for Texas counties and create three models to understand if causal relationships exist between the number of long distance trips, population income, population density (rural or urban counties).
-Our hypothesis is that an there is a causal relationship between """
+We will analyze data pertaining to these variables for Texas counties and create three models to understand if causal relationships exist between the number of long distance trips, population income, population density (rural or urban counties).
+Our hypothesis is that an there is a causal relationship between """ [summer travel]
 
--->Need to explain distance of travel with bs TRANSPORT data
+-->Need to explain distance of travel with bs TRANSPORT data[^bts01] [survey](https://www.bts.gov/statistical-products/surveys/national-household-travel-survey-long-distance-travel-quick-facts)
 -->Need to add more to causal theory to justify variable selection
 
-## Operationalization
+**Causal theory needed to identify why these particular variables below were chosen**
 
-To create our dataframe, we cleaned and filtered data from multiple sources of data and joined based on texas county name or code from June, 2021.
+# Data Description
 
+Operationalization of our research question required data from four sources: the New York Times (NYT)' _Coronavirus (Covid-19) Data in the United States_ dataset[^nyt], the United States Census Bureau's 2010 decennial survey[^uscb], the Texas Association of Counties' _County Information Program_[^tac] database, and the United States Bureau of Transportation Statistics' _Trips by Distance_[^bts] dataset.
+
+The NYT has maintained a robust public dataset of county-level, time-series COVID data on GitHub and holds records dating back to the beginning of COVID-19.
+The dataset aggregates state and municipal government health department reports of cumulative cases and deaths for each US county daily from COVID-19.
+NYT data does not match deaths or cases for patients to their home state (e.g., vacationers), nor does it identify if a case occurred in one county, but a subsequent death occurred in another.
+Therefore, there is a likely mismatch between this data and those reported by individual states and counties.
+
+For our investigation, we selected on cases and deaths in the month of June 2021, to best operationalize summer.
+The case and death counts were aggregated by county to mitigate time-series affects on modeling.
+The death counts were identified as being directly related to the COVID-19 infection and not from other instances (e.g., homicide), or unrelated comorbidities (e.g., cancer).
+
+The Texas Association of Counties proved to be a rich source of demographic and economic data for our target population.
+This database aggregates national and state data sources (e.g., Census Bureau, Texas Workforce Commission, United States Bureau of Labor Statistics, United States Bureau of Economic Analysis, and the Texas Demographic Center).
+The lineage of the TAC data was challenging to individually trace.
+All of the data originated from primary sources at the Federal or State level, but there was no readily available documentation to map columnar data back to its primary source.
+We selected five measurements to support our causal theory: county population and density, median household income, number of individuals in poverty, and the unemployment rate.
+The economic measurements of poverty and employment were hypothesized to have an inverse relationship with enabling people to take long vacation trips, where median income should have a more direct relationship.
+The county population served as a control for the economic variables and density was hypothesized to encourage longer trips to more open spaces.
+
+The decennial survey conducted by the Census results in an incredibly rich national database of economic, geographic, demographic, and social measures.
+While the vast database has many of the measurements useful to support our causal theory, as the TAC database queries held a majority of these, the decennial survey did proffer a means to generate an indicator variable identifying rural or urban counties.
+This variable was important since we hypothesized that rural counties may experience higher long trip counts due to their geographies.
+
+Our final course were data produced by the US Department of Transportation's Bureau of Transportation Statistics (USBTS).
+These data are experimental *estimates* of mobility gathrered from aggregated mobile device data.
+USBTS states that these data are experimental, and but significant effort was made to ensure proper population weighting, and data multi-sourcing (to reduce variance in spacial and temporal measurements common from a single source) were implemented before any of the statistics were calculated.
+Trips are defined as a longer than 10 minute stop at a location that is not home.
+Trips are not distiguish by mode of transporation (i.e., air, train, bus, car, or bike).
+For our analyis, we were interested in only trips longer than 100 miles in Texas in June of 2021 as this best operationalized the concept of a vacation[^bts01].
+
+[^bts01]: https://www.bts.gov/statistical-products/surveys/national-household-travel-survey-long-distance-travel-quick-facts
+[^nyt]: The New York Times.
+(2021, 08 04).
+Coronavirus (Covid-19) Data in the United States.
+https://github.com/nytimes/covid-19-data.
+Retrieved 07 21, 2021, from https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv
+[^uscb]: United States Census Bureau. (2011, 11 22). Decennial Census, 2010. Explore Census Data. Retrieved 07 21, 2021, from https://data.census.gov/
+[^tac]: Texas Association of Counties. (2021, 08 04). QueriesCIP. County Information Program. Retrieved 08 02, 2021, from https://imis.county.org/iMIS/CountyInformationProgram/
+[^bts]: Department of Transportation.
+(2021, 08 02).
+Trips by Distance.
+Bureau of Transportation Statistics.
+Retrieved 07 19, 2021, from https://data.bts.gov/Research-and-Statistics/Trips-by-Distance/w96p-f2qv
+
+---
 #### Operationalization of focus variables
+
+__integrating below still__
 We gathered covid case and death data from NYT.
@@
 
-### Variables
-
-For our analysis, we will use datasets and variables from the four following sources:
-
-[![DataSources](lab2datatable.JPG)](https://ibb.co/zQ3yj2K)
-
 As indicated, each source will provide information on county population demographics such as income, rural vs.
@@
 Together, these sources were filtered for the state of Texas and combined to create a dataframe for analysis.
+---
 
@@
 df <- read.csv('./../data/processed/covid_trips_processed.csv')
-head(df)
+# head(df)
 ```
@@
 table1::label(df$median_income) <- "County Median Income ($)"
-table1::label(df$lt17) <- "County Age < 17 years"
-table1::label(df$gt65) <- "County Age > 65 years"
+table1::label(df$lt17) <- "County Age < 17 years (%)"
+table1::label(df$gt65) <- "County Age > 65 years (%)"
 # table1::label(df$is_rural) <- "County Type"
@@
 Below are histograms of variables that did not follow normal distribution and required transformation. Each of these variables was logarithmically transformed.
-```{r scatterplots, message=FALSE}
+```{r scatterplots, message=FALSE, echo=FALSE}
 attach(df)
@@
 hist(density, main="Histogram of Population Density")
-hist(unemployed, main="Histogram of Unemployment")
+hist(unemployed, main="Histogram of Unemployment Rate")
 hist(median_income, main="Histogram of Median Income")
@@
     median_income=log(median_income+1),
-    unemployed = log(unemployed+1),
-    lt17 = log(lt17),
-    gt65 = log(gt65)
+    unemployed = log(unemployed+1)
   ) %>%
@@
 
-### Accounting Tables
+## Accounting Tables
 
@@
 
-```{r Accounting Table creation & output}
+```{r Accounting Table creation & output, echo=FALSE}
 df2 <- df
@@
 
-#### Additional Changes to the Data:
+## Additional Changes to the Data:
 
@@
 
-## A Model Building Process
+# A Model Building Process
 
 To investigate the causal relationships between distance traveled and variables within the dataset, a linear regression was performed in R Studio for three models.
+The first model is limited to study only key variables of interest for measurement.
+The second model expands upon the first model to also include key explanatory variables and covariates that are hypothesized to improve performance of the previous model.
 The third model is all-inclusive of the variables available to the team at the time of data gathering.
@@
 
-Here is the first model, which is limited to study only key variables of interest for measurement based on our hypotheses.
-Our key input variable is covid cases and key output is number of trips greater than 100 miles.
-We also chose to add an indicator for rural counties because it could be directly related to trip distance, as well as median income which plays a role in a person's ability to take vacations in the first place.
-$$
-  Model 1 = f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})
-$$ 
-The second model expands upon the first model to also include key explanatory variables and covariates that are hypothesized to improve performance of the previous model.
-Here, we added in unemployment rate to assess if the impact of covid to that county affected travel.
-$$
-  Model 2 = f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density})
-$$
-Finally in the third model, we added in age demographics as percent under 17/over 65 as another potential source that influences travel midst covid cases due to the nature of how covid affects those at different ages.
-$$
-  Model 3 = f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})
-  + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density}) + \beta_6 h(\text{log(% Age < 17}) + \beta_7 h(\text{log(% Age < 17})
-$$
+$Model_1: f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})$
 
-## Linear Regression Results
+$Model_2: f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density})$
+
+$Model_3: f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{Unemploymed}) + \beta_5 h(\text{log(Population Density}) + \beta_6 h(\text{log(Age < 17}) + \beta_7 h(\text{log(Age > 65}) + \beta_7 h(\text{log(In Poverty})$
+
+# Linear Regression Results
 
@@
 #need to change these to per capita data
-model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural, data = df) #limited model
-model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1) + log(density), data = df)
-model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1) + log(density) + log(lt17) + log(gt65), data = df) #final model
+model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1), data = df) #limited model
+model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unemployed + log(density), data = df)
+model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unemployed + log(density) + lt17 + gt65, data = df) #final model
 stargazer(model1, model2, model3, type='text', title="Results", align=TRUE)
 ```
-From this output table, we can see most of our hypotheses about the variables we included in our model were correct and do have a statistically significant contribution to our model already.
-
 Additionally, to indicate if any of the models provide a better fit to the dataset than a model that contains no independent variables, an F test was performed.
 The results for each model are shown below:
-```{r F Test}
+```{r F Test, echo=FALSE}
 #Run F test to show statistical significance of additional variables
@@
 ```
-From the F-test,  we can see the the addition of the variables into model 2 and model 3 do contribute significant information to our model based off of the derived p-values.
+From the F-test,  we can see the the addition of the variables into model 2 and model 3 do contribute signficant information to our model based off of the derived p-values.
 
 Last, the significance of individual regression coefficients was tested using robust standard errors in R studio for model 3. The results are shown below:
-```{r coeftest}
+```{r coeftest, echo=FALSE}
 #Run coeftest to show statistical significance of all variables in the model with robust standard errors
-coeftest(model3, vcov = vcovHC(model3, type="HC1"))
+coeftest(model2, vcov = vcovHC(model3, type="HC1"))
 ```
 From the t-test, we can see that Covid-19 cases and median income are statistically significant.
-Additionally, the percentage of adults greater than 65, the population density, and the Covid-19 death variables are even more statistically significant.
+Additionally, the percentage of adults greather than 65, the population density, and the Covid-19 death variables are even more statistically significant.
 
@@
 
-### Model Limitations
+# Model Limitations
 
@@
 
-```{r VIF & correlation}
+```{r VIF & correlation, echo=FALSE}
 
@@
 
-```{r CLM assumptions check}
+```{r CLM assumptions check, echo=FALSE}
 
@@
 
-### Omitted Variable
+# Omitted Variable
 
-#### Variable 1
+### Vehicle Ownership
 
@@
 
-#### Variable 2
+### Fuel Prices
 
@@
 
-## Conclusion & Next Steps
+# Conclusion
 
@@
 Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model.
+
+### Next Steps
 
@@
 Last, it is highly recommended that any model generated in the future continue to use the latest data available given the oscillation in vaccination rates, government policy, and new or known variants of the Covid-19 virus.
-
-## References

### 2021-08-05 notebooks/Draft_Report.Rmd

@@
 Do COVID-19 infection rates impact vacation travel distance for U.S. residents in the state of Texas? 
-How does income, population density, and covid regulations in each county impact traveling?
+Does income, population density, unemployment rates, and age in Texas counties impact traveling?
 
-We will analyze data pertaining to these variables for Texas counties and create three models to understand if causal relationships exist between the number of long distance trips, population income, population density (rural or urban counties).
-Our hypothesis is that an there is a causal relationship between """ [summer travel]
+We will analyze datasets from June, 2021 pertaining to these variables for Texas counties and create three models to understand if causal relationships exist between the number of long distance trips, population income, population density (rural or urban counties).
+Our hypothesis is that an there is a causal relationship between """
 
@@
 table1::label(df$median_income) <- "County Median Income ($)"
-table1::label(df$lt17) <- "County Age < 17 years (%)"
-table1::label(df$gt65) <- "County Age > 65 years (%)"
+table1::label(df$lt17) <- "County Age < 17 years"
+table1::label(df$gt65) <- "County Age > 65 years"
 # table1::label(df$is_rural) <- "County Type"
@@
 hist(density, main="Histogram of Population Density")
-hist(unemployed, main="Histogram of Unemployment Rate")
+hist(unemployed, main="Histogram of Unemployment")
 hist(median_income, main="Histogram of Median Income")
@@
     median_income=log(median_income+1),
-    unemployed = log(unemployed+1)
+    unemployed = log(unemployed+1),
+    lt17 = log(lt17),
+    gt65 = log(gt65)
   ) %>%
@@
 To investigate the causal relationships between distance traveled and variables within the dataset, a linear regression was performed in R Studio for three models.
-The first model is limited to study only key variables of interest for measurement.
+The first model includes only our base variables.
+In the second model, we added additional county characteristics.
+In the final model, we added in another covid related factor.
+
+Here is the first model, which is limited to study only key variables of interest for measurement based on our hypotheses.
+Our key input variable is covid cases and key output is number of trips greater than 100 miles.
+We also chose to add an indicator for rural counties because it could be directly related to trip distance
+$Model_1: f(\text{log(#Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)}) + \beta_2 g(\text{log(Covid-19 Deaths)})  + \beta_3 h(\text{Rural County})$
 The second model expands upon the first model to also include key explanatory variables and covariates that are hypothesized to improve performance of the previous model.
-The third model is all-inclusive of the variables available to the team at the time of data gathering.
-Details of each model analyzed can be found below.
-
-$Model_1: f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})$
-
-$Model_2: f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density})$
-
-$Model_3: f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{Unemploymed}) + \beta_5 h(\text{log(Population Density}) + \beta_6 h(\text{log(Age < 17}) + \beta_7 h(\text{log(Age > 65}) + \beta_7 h(\text{log(In Poverty})$
+Here, we added in median income and density of the county as additional descriptors of the county.
+$Model_2: f(\text{log(#Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)}) + \beta_2 g(\text{log(Covid-19 Deaths)})  + \beta_3 h(\text{Rural County}) + \beta_4 h(\text{Median Income}) + \beta_5 h(\text{County Density})$
+Finally in the third model, we added in age demographics as percent under 17/over 65 as another potential source that influences travel midst covid cases due to the nature of how covid affects those at different ages.
+$Model_3: f(\text{log(#Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)}) + \beta_2 g(\text{log(Covid-19 Deaths)})  + \beta_3 h(\text{Rural County}) + \beta_4 h(\text{Median Income}) + \beta_5 h(\text{County Density}) + \beta_6 h(\text{Population under 17}) + \beta_7 h(\text{Population over 65})$
 
@@
 ```{r model creation, echo=FALSE, warning=FALSE}
+model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural, data = df) #limited model
+model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1) + log(density), data = df)
+model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1) + log(density) + log(lt17) + log(gt65), data = df) #final model
 
-#need to change these to per capita data
-model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1), data = df) #limited model
-model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unemployed + log(density), data = df)
-model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income + unemployed + log(density) + lt17 + gt65, data = df) #final model
 stargazer(model1, model2, model3, type='text', title="Results", align=TRUE)
 ```
+From this output table, we can discern quite a bit of information.
+
+First, we consider the models holistically. Here are some of our observations:
+* In model1, all of our variables are extremely significant, as expected from my hypotheses, though in the opposite direction.
+We would expect an inverse relationship between covid and travel, though it seems that this is estimating a positive relationship
+* In model2, when county density is factored in, the presence of covid seems to have less positive of an effect on travel and that higher density counties see more travel. 
+* In model3, with additionally age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17. 
+* In terms of R^2 and RSE, we are improving with each model. 
+* The addition of our chosen parameters for each following model is also statistically better than the previous, as indicated by the F statistic.
+
+We will also interpret the coefficient values and what these mean in the context of our model.
+Almost all of our variables needed a log transform, including the output, making this a log-log regression.
+These regressions are most intuitive to understand in terms of percent increases.
+* First to note is our persistently significant constant in each model.
+Regardless of any variables, we are getting between ~300 (as with model3) and over 8000 (as with model2) trips over 100 miles.
+This is calculated as e^(Constant).
+* In model1, for every percent increase in both cases and deaths, we are getting almost half a percent increase in trips.
+The positive relationship doesnt make much sense, so we will look to the other models.
+* In model2, we see that density has a more positive relationship with trips than do covid cases and deaths. 
+* In model3, covid cases and deaths become insiginifcant, though its lookign
+
 Additionally, to indicate if any of the models provide a better fit to the dataset than a model that contains no independent variables, an F test was performed.
@@
 #Run coeftest to show statistical significance of all variables in the model with robust standard errors
-coeftest(model2, vcov = vcovHC(model3, type="HC1"))
+coeftest(model3, vcov = vcovHC(model3, type="HC1"))
 ```
 From the t-test, we can see that Covid-19 cases and median income are statistically significant.
-Additionally, the percentage of adults greather than 65, the population density, and the Covid-19 death variables are even more statistically significant.
-
-## Discussion
+Additionally, the percentage of adults greater than 65, the population density, and the Covid-19 death variables are even more statistically significant.
 

### 2021-08-05 notebooks/Draft_Report.Rmd
model explanation complete

--- 2021-08-05 notebooks/Draft_Report.Rmd
@@
 ```
+
+## Discussion
 From this output table, we can discern quite a bit of information.
@@
 First, we consider the models holistically. Here are some of our observations:
-* In model1, all of our variables are extremely significant, as expected from my hypotheses, though in the opposite direction.
-We would expect an inverse relationship between covid and travel, though it seems that this is estimating a positive relationship
+* In model1, all of our variables are extremely significant, though in the opposite direction as expected.
+We would expect an inverse relationship between covid and travel, though it seems that this model is producing a positive relationship.
 * In model2, when county density is factored in, the presence of covid seems to have less positive of an effect on travel and that higher density counties see more travel. 
-* In model3, with additionally age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17. 
+* In model3 with age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17. 
 * In terms of R^2 and RSE, we are improving with each model. 
@@
 The positive relationship doesnt make much sense, so we will look to the other models.
-* In model2, we see that density has a more positive relationship with trips than do covid cases and deaths. 
-* In model3, covid cases and deaths become insiginifcant, though its lookign
-
-Additionally, to indicate if any of the models provide a better fit to the dataset than a model that contains no independent variables, an F test was performed.
-The results for each model are shown below:
-```{r F Test}
-#Run F test to show statistical significance of additional variables
-anova(model1, model2, model3, test = "F")
-```
-From the F-test,  we can see the the addition of the variables into model 2 and model 3 do contribute significant information to our model based off of the derived p-values.
-
-Last, the significance of individual regression coefficients was tested using robust standard errors in R studio for model 3. The results are shown below: 
-```{r coeftest}
-#Run coeftest to show statistical significance of all variables in the model with robust standard errors
-coeftest(model3, vcov = vcovHC(model3, type="HC1"))
-```
-From the t-test, we can see that Covid-19 cases and median income are statistically significant.
-Additionally, the percentage of adults greater than 65, the population density, and the Covid-19 death variables are even more statistically significant.
-
-## Discussion
+* In model2, we see that density has a more positive relationship with trips, with every percent increase in county density modeled to a half percent increase in trips.
+Also the positive relationship between covid cases/ deaths and trips has decreased to resulting in 0.15 and 0.25% increase in trips respectively.
+* In model3, covid cases and deaths become insignificant.
+The coefficients have become slightly negative as in increase in cases and deaths decreases trips, which is in line with our original hypothesis, but is not significant.
+Instead, for every percent median income of the county increases by, the amount of long trips decrease by 0.3%.
+This could indicate that more affluent counties trust science and err on the safe side in the midst of covid.
+Finally, in our 3rd model, counties with more people under 17 years of age are more likely to take long trips.
+Due to the time frame of our analysis, this also makes sense because june is the start of summer break when many families like to take vacations together.
 
 ### Model Limitations
+
+DELTE*** Not CLM since >100 samples. Large sample assumptions = IID and unique BLP (which is related to colinearity...)
 
@@
 The goal of this analysis was to determine if there existed a causal relationship between distance traveled and Covid-19 infection rates for residents of the state of Texas.
-Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model.
+Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model, though with all variables considered, our results do not indicate vacation and covid cases are necessarily related.
 
 For future analysis, it is recommended that Covid-19 policy adherence data per county be incorporated into the model to increase the accuracy.
-While the Govenor of Texas issued detailed protocols or mandates for the state of Texas, the degree to which each county (and thus the individual residents) complied would be useful data to have if such data were to exist.
+While the Governor of Texas issued detailed protocols or mandates for the state of Texas, the degree to which each county (and thus the individual residents) complied would be useful data to have if such data were to exist.
 Additionally, incorporating vaccination rates could aid in development of a model with even higher accuracy.

### 2021-08-05 notebooks/Draft_Report.Rmd
updated intro

@@
 As Ella Fitzgerald so aptly sings, “Summertime, and the living is easy.” However, travel in the recent peak-vacation months has certainly deviated from this idealistic frame of mind with the onset of the COVID-19 pandemic in early 2020.
-The global travel industry has experienced sharp declines.
+The global travel industry has experienced sharp declines in Revenue.
 Specifically in the United States, “travel spending totaled a mere $679 billion in 2020, an unprecedented 42% annual decline (nearly $500 billion) from 2019,” according to analysis from the U.S.
 Travel Association.
-As infection and vaccination rates oscillate regionally along with government policies for business operation restrictions in the United States, our team of data scientists will look to answer the following research question in more detail using linear regression techniques:
+Our hypothesis as to why this decline has occurred is due directly to the presence of the Covid-19 virus and the number of cases.
+As virus infection oscillate regionally in the United States, our team of data scientists will look to answer the following primary research question using linear regression statistical techniques:
 
-Do COVID-19 infection rates impact vacation travel distance for U.S. residents in the state of Texas? 
-Does income, population density, unemployment rates, and age in Texas counties impact traveling?
+Do COVID-19 cases impact vacation travel trips for Texas for June, 2021?
 
-We will analyze datasets from June, 2021 pertaining to these variables for Texas counties and create three models to understand if causal relationships exist between the number of long distance trips, population income, population density (rural or urban counties).
-Our hypothesis is that an there is a causal relationship between """
+Our primary hypothesis is that there is a causal relationship between the number of Covid-19 cases and the number of vacation trips; when Covid-19 cases increase, the number of vacation travel trips decrease.
+To investigate, a dataset comprising of Texas county data will be analyzed.
+Based on the data, our team will also investigate the following secondary question:
 
--->Need to explain distance of travel with bs TRANSPORT data
--->Need to add more to causal theory to justify variable selection
+Do income, population density, age, and/or county type (rural or urban) per county impact traveling for Texas residents?
+
+For the analysis, we will create three models to understand if causal relationships exist between the outcome variable, the number of long distance trips, and the predictor variables per Texas county: income, population density, age, Covid-19 cases and deaths .
+We will use an indicator variable, rural counties, to compare performance of our models for rural or urban counties.
 
@@
 
-To create our dataframe, we cleaned and filtered data from multiple sources of data and joined based on texas county name or code from June, 2021.
+To create our dataframe, we cleaned and filtered data from multiple sources of data and joined based on texas county name or code.
 
@@
 table1::label(df$cases) <- "County Covid-19 Cases"
-table1::label(df$unemployed) <- "County Unemployment Rate"
+table1::label(df$unemp_rate) <- "County Unemployment Rate"
 table1::label(df$median_income) <- "County Median Income ($)"
-table1::label(df$lt17) <- "County Age < 17 years"
-table1::label(df$gt65) <- "County Age > 65 years"
+table1::label(df$pct_lt17) <- "County Age < 17 years (%)"
+table1::label(df$pct_gt65) <- "County Age > 65 years (%)"
 # table1::label(df$is_rural) <- "County Type"
-table1::table1(~density + cases + unemployed + median_income + lt17 + gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Statistics", topclass="Rtable1-zebra", overall = "Total")
+table1::table1(~density + cases + unemp_rate + median_income + pct_lt17 + pct_gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Statistics", topclass="Rtable1-zebra", overall = "Total")
 ```
@@
 hist(deaths, main="Histogram of Covid-19 Deaths")
-hist(long_trip, main="Histogram of Distance")
+hist(long_trip, main="Histogram of Trip Count")
 hist(density, main="Histogram of Population Density")
-hist(unemployed, main="Histogram of Unemployment")
+hist(unemp_rate, main="Histogram of Unemployment Rate")
+hist(log(median_income), main="Histogram of Median Income")
 hist(median_income, main="Histogram of Median Income")
+summary(df$median_income)
+summary(df$long_trip)
 
@@
     median_income=log(median_income+1),
-    unemployed = log(unemployed+1),
-    lt17 = log(lt17),
-    gt65 = log(gt65)
+    unemp_rate = log(unemp_rate+1)
   ) %>%
-  select(long_trip, cases, deaths, median_income, unemployed, density, lt17, gt65) %>%
+  select(long_trip, cases, deaths, median_income, unemp_rate, density, pct_lt17, pct_gt65) %>%
   GGally::ggpairs(.,
@@
 To investigate the causal relationships between distance traveled and variables within the dataset, a linear regression was performed in R Studio for three models.
-The first model includes only our base variables.
-In the second model, we added additional county characteristics.
-In the final model, we added in another covid related factor.
+The first model is limited to study only key variables of interest for measurement.
+The second model expands upon the first model to also include key explanatory variables and covariates that are hypothesized to improve performance of the previous model.
+The third model is all-inclusive of the variables available to the team at the time of data gathering.
+Details of each model analyzed can be found below.
 
-Here is the first model, which is limited to study only key variables of interest for measurement based on our hypotheses.
-Our key input variable is covid cases and key output is number of trips greater than 100 miles.
-We also chose to add an indicator for rural counties because it could be directly related to trip distance
 $$
-  Model 1 = f(\text{log(#Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)}) + \beta_2 g(\text{log(Covid-19 Deaths)})  + \beta_3 h(\text{Rural County})
+  Model 1 = f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})
 $$
-The second model expands upon the first model to also include key explanatory variables and covariates that are hypothesized to improve performance of the previous model.
-Here, we added in median income and density of the county as additional descriptors of the county.
+
 $$
-  Model 2 = f(\text{log(#Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)}) + \beta_2 g(\text{log(Covid-19 Deaths)})  + \beta_3 h(\text{Rural County}) + \beta_4 h(\text{Median Income}) + \beta_5 h(\text{County Density})
+  Model 2 = f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density})
 $$
-Finally in the third model, we added in age demographics as percent under 17/over 65 as another potential source that influences travel midst covid cases due to the nature of how covid affects those at different ages.
+
 $$
-  Model 3 = f(\text{log(#Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)}) + \beta_2 g(\text{log(Covid-19 Deaths)})  + \beta_3 h(\text{Rural County}) + \beta_4 h(\text{Median Income}) + \beta_5 h(\text{County Density}) + \beta_6 h(\text{Population under 17}) + \beta_7 h(\text{Population over 65})
+  Model 3 = f(\text{log(Number of Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)})  + \beta_2 h(\text{Rural County}) + \beta_3 h(\text{Median Income})
+  + \beta_4 h(\text{Unemployment Rate}) + \beta_5 h(\text{log(Population Density}) + \beta_6 h(\text{log(% Age < 17}) + \beta_7 h(\text{log(% Age < 17})
 $$
@@
 #need to change these to per capita data
-model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural, data = df) #limited model
-model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1) + log(density), data = df)
-model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1) + log(density) + log(lt17) + log(gt65), data = df) #final model
-stargazer(model1, model2, model3, type='text', title="Results", align=TRUE)
+model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1), data = df) #limited model
+summary(model1)
+model1b <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + median_income, data = df) #limited model
+summary(model1b)
+summary(df$median_income)
+
+model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income) + unemp_rate + log(density), data = df)
+model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income)+ unemp_rate + log(density) + pct_lt17 + pct_gt65, data = df) #final model
+stargazer(model1, model1b, model2, model3, type='text', title="Results", align=TRUE)
 ```
+Additionally, to indicate if any of the models provide a better fit to the dataset than a model that contains no independent variables, an F test was performed.
+The results for each model are shown below:
+```{r F Test}
+#Run F test to show statistical significance of additional variables
+anova(model1, model2, model3, test = "F")
+```
+From the F-test,  we can see the the addition of the variables into model 2 and model 3 do contribute signficant information to our model based off of the derived p-values.
+
+Last, the significance of individual regression coefficients was tested using robust standard errors in R studio for model 3. The results are shown below: 
+```{r coeftest}
+#Run coeftest to show statistical significance of all variables in the model with robust standard errors
+coeftest(model2, vcov = vcovHC(model3, type="HC1"))
+```
+From the t-test, we can see that Covid-19 cases and median income are statistically significant.
+Additionally, the percentage of adults greather than 65, the population density, and the Covid-19 death variables are even more statistically significant.
 
 ## Discussion
-From this output table, we can discern quite a bit of information.
-
-First, we consider the models holistically. Here are some of our observations:
-* In model1, all of our variables are extremely significant, though in the opposite direction as expected.
-We would expect an inverse relationship between covid and travel, though it seems that this model is producing a positive relationship.
-* In model2, when county density is factored in, the presence of covid seems to have less positive of an effect on travel and that higher density counties see more travel. 
-* In model3 with age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17. 
-* In terms of R^2 and RSE, we are improving with each model. 
-* The addition of our chosen parameters for each following model is also statistically better than the previous, as indicated by the F statistic.
-
-We will also interpret the coefficient values and what these mean in the context of our model.
-Almost all of our variables needed a log transform, including the output, making this a log-log regression.
-These regressions are most intuitive to understand in terms of percent increases.
-* First to note is our persistently significant constant in each model.
-Regardless of any variables, we are getting between ~300 (as with model3) and over 8000 (as with model2) trips over 100 miles.
-This is calculated as e^(Constant).
-* In model1, for every percent increase in both cases and deaths, we are getting almost half a percent increase in trips.
-The positive relationship doesnt make much sense, so we will look to the other models.
-* In model2, we see that density has a more positive relationship with trips, with every percent increase in county density modeled to a half percent increase in trips.
-Also the positive relationship between covid cases/ deaths and trips has decreased to resulting in 0.15 and 0.25% increase in trips respectively.
-* In model3, covid cases and deaths become insignificant.
-The coefficients have become slightly negative as in increase in cases and deaths decreases trips, which is in line with our original hypothesis, but is not significant.
-Instead, for every percent median income of the county increases by, the amount of long trips decrease by 0.3%.
-This could indicate that more affluent counties trust science and err on the safe side in the midst of covid.
-Finally, in our 3rd model, counties with more people under 17 years of age are more likely to take long trips.
-Due to the time frame of our analysis, this also makes sense because june is the start of summer break when many families like to take vacations together.
 
 ### Model Limitations
-
-DELTE*** Not CLM since >100 samples. Large sample assumptions = IID and unique BLP (which is related to colinearity...)
 
@@
 The goal of this analysis was to determine if there existed a causal relationship between distance traveled and Covid-19 infection rates for residents of the state of Texas.
-Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model, though with all variables considered, our results do not indicate vacation and covid cases are necessarily related.
+Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model.
 
 For future analysis, it is recommended that Covid-19 policy adherence data per county be incorporated into the model to increase the accuracy.
-While the Governor of Texas issued detailed protocols or mandates for the state of Texas, the degree to which each county (and thus the individual residents) complied would be useful data to have if such data were to exist.
+While the Govenor of Texas issued detailed protocols or mandates for the state of Texas, the degree to which each county (and thus the individual residents) complied would be useful data to have if such data were to exist.
 Additionally, incorporating vaccination rates could aid in development of a model with even higher accuracy.

### 2021-08-05 notebooks/Draft_Report.Rmd
model revision

--- 2021-08-05 notebooks/Draft_Report.Rmd
@@
 Summer vacations in the United States are synonymous with sunny, leisurely days spent with loved ones.
-As Ella Fitzgerald so aptly sings, “Summertime, and the living is easy.” However, travel in the recent peak-vacation months has certainly deviated from this idealistic frame of mind with the onset of the SARS-CoV-2 outbreak (COVID-19) in early 2020.
-The global travel industry has experienced sharp declines.
-Specifically in the United States (US), “travel spending totaled a mere $679 billion in 2020, an unprecedented 42% annual decline (nearly $500 billion) from 2019,” according to analysis from the U.S.
+As Ella Fitzgerald so aptly sings, “Summertime, and the living is easy.” However, travel in the recent peak-vacation months has certainly deviated from this idealistic frame of mind with the onset of the COVID-19 pandemic in early 2020.
+The global travel industry has experienced sharp declines in Revenue.
+Specifically in the United States, “travel spending totaled a mere $679 billion in 2020, an unprecedented 42% annual decline (nearly $500 billion) from 2019,” according to analysis from the U.S.
 Travel Association.
-As infection and vaccination rates oscillate regionally along with government policies for business operation restrictions in the United States, our team of data scientists will look to answer the following research question in more detail using linear regression techniques:
+Our hypothesis as to why this decline has occurred is due directly to the presence of the COVID-19 virus and the number of cases.
+As virus infection oscillate regionally in the United States, our team of data scientists will look to answer the following primary research question using linear regression statistical techniques:
 
-Do COVID-19 infection rates impact vacation travel distance for U.S. residents in the state of Texas? 
-Does income, population density, unemployment rates, and age in Texas counties impact traveling?
+Do COVID-19 cases impact vacation travel trips for Texas for June, 2021?
 
-We will analyze datasets from June, 2021 pertaining to these variables for Texas counties and create three models to understand if causal relationships exist between the number of long distance trips, population income, population density (rural or urban counties).
-Our hypothesis is that an there is a causal relationship between """
+Our primary hypothesis is that there is a causal relationship between the number of COVID-19 cases and the number of vacation trips; when COVID-19 cases increase, the number of vacation travel trips decrease.
+To investigate, a dataset comprising of Texas county data will be analyzed.
+Based on the data, our team will also investigate the following secondary question:
 
--->Need to explain distance of travel with bs TRANSPORT data[^bts01] [survey](https://www.bts.gov/statistical-products/surveys/national-household-travel-survey-long-distance-travel-quick-facts)
--->Need to add more to causal theory to justify variable selection
+Do income, population density, age, and/or county type (rural or urban) per county impact traveling for Texas residents?
 
-**Causal theory needed to identify why these particular variables below were chosen**
+For the analysis, we will create three models to understand if causal relationships exist between the outcome variable, the number of long distance trips, and the control variables per Texas county: income, population density, age, COVID-19 cases and deaths .
+We will use an indicator variable, rural counties, to compare performance of our models for both rural or urban counties.
+By creating models with these variables, we intend to answer our primary and secondary research questions to provide insight about how vacation travel has been impacted so that interested stakeholders can gain an understanding of how to improve the revenue declines in the travel industry.
 
@@
 
-Operationalization of our research question required data from four sources: the New York Times (NYT)' _Coronavirus (Covid-19) Data in the United States_ dataset[^nyt], the United States Census Bureau's 2010 decennial survey[^uscb], the Texas Association of Counties' _County Information Program_[^tac] database, and the United States Bureau of Transportation Statistics' _Trips by Distance_[^bts] dataset.
+Operationalization of our research question required data from four sources: the New York Times (NYT)' _Coronavirus (COVID-19) Data in the United States_ dataset[^nyt], the United States Census Bureau's 2010 decennial survey[^uscb], the Texas Association of Counties' _County Information Program_[^tac] database, and the United States Bureau of Transportation Statistics' _Trips by Distance_[^bts] dataset.
 
@@
 Trips are not distiguish by mode of transporation (i.e., air, train, bus, car, or bike).
-For our analyis, we were interested in only trips longer than 100 miles in Texas in June of 2021 as this best operationalized the concept of a vacation[^bts01].
+For our analyis, we were interested in only trips longer than 100 miles in Texas in June of 2021 as this best operationalize the concept of a vacation[^bts01].
 
@@
 (2021, 08 04).
-Coronavirus (Covid-19) Data in the United States.
-https://github.com/nytimes/covid-19-data.
-Retrieved 07 21, 2021, from https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv
+Coronavirus (COVID-19) Data in the United States.
+https://github.com/nytimes/COVID-19-data.
+Retrieved 07 21, 2021, from https://raw.githubusercontent.com/nytimes/COVID-19-data/master/us-counties.csv
 [^uscb]: United States Census Bureau. (2011, 11 22). Decennial Census, 2010. Explore Census Data. Retrieved 07 21, 2021, from https://data.census.gov/
@@
 __integrating below still__
-We gathered covid case and death data from NYT.
+We gathered COVID case and death data from NYT.
 We gathered trip distance data from the Bureau of Transportation, which they record based on data gathered from personal mobile devices.
@@
 #### Operationalization of additional variables
-To enrich our model and limit OVB, we brainstormed a number of additional variables that could affect travel rates during covid.
-For one, we looked at how covid affected the county, through the unemployment rate.
+To enrich our model and limit OVB, we brainstormed a number of additional variables that could affect travel rates during COVID.
+For one, we looked at how COVID affected the county, through the unemployment rate.
 We also looked at age, gender, income demographics.
-Then, in addition to income demographics, we focused on percent of younger folks, as defined as those younger than 17 and older folks as defined as those older than 65 because of the nature of how covid affects an individual.
+Then, in addition to income demographics, we focused on percent of younger folks, as defined as those younger than 17 and older folks as defined as those older than 65 because of the nature of how COVID affects an individual.
 
 As indicated, each source will provide information on county population demographics such as income, rural vs.
-urban classification, and status of pandemic regulations.
+urban classification, etc.
 Together, these sources were filtered for the state of Texas and combined to create a dataframe for analysis.
@@
 df$rural_cat <- ifelse(df$is_rural == 1, "Rural", "Urban")
-table1::label(df$density) <- "County Population Density"
-table1::label(df$cases) <- "County Covid-19 Cases"
-table1::label(df$unemployed) <- "County Unemployment Rate"
-table1::label(df$median_income) <- "County Median Income ($)"
-table1::label(df$lt17) <- "County Age < 17 years"
-table1::label(df$gt65) <- "County Age > 65 years"
-# table1::label(df$is_rural) <- "County Type"
-table1::table1(~density + cases + unemployed + median_income + lt17 + gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Statistics", topclass="Rtable1-zebra", overall = "Total")
+table1::label(df$density) <- "Population Density"
+table1::label(df$cases) <- "COVID-19 Cases"
+table1::label(df$deaths) <- "COVID-19 Deaths"
+# table1::label(df$unemployed) <- "Unemployment Count"
+table1::label(df$median_income) <- "Median Income ($)"
+# table1::label(df$in_poverty) <- "Poverty Count"
+table1::label(df$lt17) <- "Age < 17 years"
+table1::label(df$gt65) <- "Age > 65 years"
+# table1::label(df$is_rural) <- "Type"
+table1::table1(~density + cases + deaths + median_income + lt17 + gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Characteristics", topclass="Rtable1-zebra", overall = "Total")
+# table1::table1(~density + cases + deaths + median_income + unemployed + in_poverty + lt17 + gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Characteristics", topclass="Rtable1-zebra", overall = "Total")
 ```
-Below are histograms of variables that did not follow normal distribution and required transformation. Each of these variables was logarithmically transformed.
-```{r scatterplots, message=FALSE, echo=FALSE}
+
+Below are histograms of variables that did not follow normal distribution and required transformation. Each of these variables were logarithmically transformed.
+
+```{r var plots, fig.align='center', message=FALSE, echo=FALSE}
 attach(df)
 par(mfrow=c(3,2))
-hist(cases, main="Histogram of Covid-19 Cases")
-hist(deaths, main="Histogram of Covid-19 Deaths")
-hist(long_trip, main="Histogram of Distance")
+hist(cases, main="Histogram of COVID-19 Cases")
+hist(deaths, main="Histogram of COVID-19 Deaths")
+hist(long_trip, main="Histogram of Trip Count")
 hist(density, main="Histogram of Population Density")
@@
 hist(median_income, main="Histogram of Median Income")
+# summary((deaths+1)^(1/2))
+```
 
-```
+Upon transformation, below is a condensed view of the statistics of each variable:
 
@@
     cases = log(cases+1),
-    deaths = log(deaths+1),
+    # deaths = log(deaths+1),
     long_trip = log(long_trip),
@@
     median_income=log(median_income+1),
-    unemployed = log(unemployed+1),
+    # unemployed = log(unemployed+1),
     lt17 = log(lt17),
@@
   ) %>%
-  select(long_trip, cases, deaths, median_income, unemployed, density, lt17, gt65) %>%
+  select(long_trip, cases, median_income, density, lt17, gt65) %>%
   GGally::ggpairs(.,
@@
                diag = list(continuous = wrap("densityDiag", alpha = 0.5)),
-               columnLabels=c("Distance", "Cases", "Deaths", "Income", "Unemployment", "Pop Density", "Age < 17","Age > 65"))+
+               columnLabels=c("Distance", "Cases", "Income", "Pop Density", "Age < 17","Age > 65"))+
                theme(legend.position = "bottom")
@@
 In the second model, we added additional county characteristics.
-In the final model, we added in another covid related factor.
+In the final model, we added in another COVID related factor.
 
 Here is the first model, which is limited to study only key variables of interest for measurement based on our hypotheses.
-Our key input variable is covid cases and key output is number of trips greater than 100 miles.
+Our key input variable is COVID cases and key output is number of trips greater than 100 miles.
 We also chose to add an indicator for rural counties because it could be directly related to trip distance
-$Model_1: f(\text{log(#Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)}) + \beta_2 g(\text{log(Covid-19 Deaths)})  + \beta_3 h(\text{Rural County})$
+
+$Model_1: f(\text{log(Long Trip Count)}) = \beta_0 + \beta_1 g(\text{log(COVID-19 Cases)}) + \beta_2 g(\text{log(COVID-19 Deaths)})  + \beta_3 h(\text{Rural County})$
+
 The second model expands upon the first model to also include key explanatory variables and covariates that are hypothesized to improve performance of the previous model.
 Here, we added in median income and density of the county as additional descriptors of the county.
-$Model_2: f(\text{log(#Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)}) + \beta_2 g(\text{log(Covid-19 Deaths)})  + \beta_3 h(\text{Rural County}) + \beta_4 h(\text{Median Income}) + \beta_5 h(\text{County Density})$
-Finally in the third model, we added in age demographics as percent under 17/over 65 as another potential source that influences travel midst covid cases due to the nature of how covid affects those at different ages.
-$Model_3: f(\text{log(#Long Trips)}) = \beta_0 + \beta_1 g(\text{log(Covid-19 Cases)}) + \beta_2 g(\text{log(Covid-19 Deaths)})  + \beta_3 h(\text{Rural County}) + \beta_4 h(\text{Median Income}) + \beta_5 h(\text{County Density}) + \beta_6 h(\text{Population under 17}) + \beta_7 h(\text{Population over 65})$
+
+$Model_2: f(\text{log(Long Trip Count)}) = \beta_0 + \beta_1 g(\text{log(COVID-19 Cases)}) + \beta_2 g(\text{log(COVID-19 Deaths)})  + \beta_3 h(\text{Rural County}) + \beta_4 h(\text{Median Income}) + \beta_5 h(\text{County Density})$
+
+Finally in the third model, we added in age demographics as percent under 17/over 65 as another potential source that influences travel midst COVID cases due to the nature of how COVID affects those at different ages.
+
+$Model_3: f(\text{log(Long Trip Count)}) = \beta_0 + \beta_1 g(\text{log(COVID-19 Cases)}) + \beta_2 g(\text{log(COVID-19 Deaths)})  + \beta_3 h(\text{Rural County}) + \beta_4 h(\text{Median Income}) + \beta_5 h(\text{County Density}) + \beta_6 h(\text{Population under 17}) + \beta_7 h(\text{Population over 65})$
 
@@
 ```{r model creation, echo=FALSE, warning=FALSE}
-model1 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural, data = df) #limited model
-model2 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1) + log(density), data = df)
-model3 <- lm(log(long_trip) ~ log(cases+1) + log(deaths+1) + is_rural + log(median_income+1) + log(density) + log(lt17) + log(gt65), data = df) #final model
+model1 <- lm(log(long_trip) ~ log(cases+1) + is_rural, data = df) #limited model
+model2 <- lm(log(long_trip) ~ log(cases+1) + is_rural + log(median_income+1) + log(density), data = df)
+model3 <- lm(log(long_trip) ~ log(cases+1) + is_rural + log(median_income+1) + log(density) + log(lt17) + log(gt65), data = df) #final model
 
 stargazer(model1, model2, model3, type='text', title="Results", align=TRUE)
+
 ```
+
+## Discussion
 From this output table, we can discern quite a bit of information.
@@
 First, we consider the models holistically. Here are some of our observations:
-* In model1, all of our variables are extremely significant, as expected from my hypotheses, though in the opposite direction.
-We would expect an inverse relationship between covid and travel, though it seems that this is estimating a positive relationship
-* In model2, when county density is factored in, the presence of covid seems to have less positive of an effect on travel and that higher density counties see more travel. 
-* In model3, with additionally age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17. 
-* In terms of R^2 and RSE, we are improving with each model. 
+* In model1, all of our variables are extremely significant, though in the opposite direction as expected.
+We would expect an inverse relationship between COVID and travel, though it seems that this model is producing a positive relationship.
+* In model2, when county density is factored in, the presence of COVID seems to have less positive of an effect on travel and that higher density counties see more travel.
+* In model3 with age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17.
+* In terms of $R^2$ and RSE, we are improving with each model.
 * The addition of our chosen parameters for each following model is also statistically better than the previous, as indicated by the F statistic.
@@
 This is calculated as e^(Constant).
-* In model1, for every percent increase in both cases and deaths, we are getting almost half a percent increase in trips.
+* In $Model_1$, for every percent increase in both cases and deaths, we are getting almost half a percent increase in trips.
 The positive relationship doesnt make much sense, so we will look to the other models.
-* In model2, we see that density has a more positive relationship with trips than do covid cases and deaths. 
-* In model3, covid cases and deaths become insiginifcant, though its lookign
-
-Additionally, to indicate if any of the models provide a better fit to the dataset than a model that contains no independent variables, an F test was performed.
-The results for each model are shown below:
-```{r F Test, echo=FALSE}
-#Run F test to show statistical significance of additional variables
-anova(model1, model2, model3, test = "F")
-```
-From the F-test,  we can see the the addition of the variables into model 2 and model 3 do contribute significant information to our model based off of the derived p-values.
-
-Last, the significance of individual regression coefficients was tested using robust standard errors in R studio for model 3. The results are shown below:
-```{r coeftest, echo=FALSE}
-#Run coeftest to show statistical significance of all variables in the model with robust standard errors
-coeftest(model3, vcov = vcovHC(model3, type="HC1"))
-```
-From the t-test, we can see that Covid-19 cases and median income are statistically significant.
-Additionally, the percentage of adults greater than 65, the population density, and the Covid-19 death variables are even more statistically significant.
+* In $Model_2$, we see that density has a more positive relationship with trips, with every percent increase in county density modeled to a half percent increase in trips.
+Also the positive relationship between COVID cases/ deaths and trips has decreased to resulting in 0.15 and 0.25% increase in trips respectively.
+* In $Model_3$, COVID cases and deaths become insignificant.
+The coefficients have become slightly negative as in increase in cases and deaths decreases trips, which is in line with our original hypothesis, but is not significant.
+Instead, for every percent median income of the county increases by, the amount of long trips decrease by 0.3%.
+This could indicate that more affluent counties trust science and err on the safe side in the midst of COVID.
+Finally, in our 3rd model, counties with more people under 17 years of age are more likely to take long trips.
+Due to the time frame of our analysis, this also makes sense because june is the start of summer break when many families like to take vacations together.
 
 # Model Limitations
+
+DELTE*** Not CLM since >100 samples. Large sample assumptions = IID and unique BLP (which is related to colinearity...)
 
@@
 
-3.
+<!-- 3.
 **Linear Conditional Expectation:** For this assumption, the residuals for linear relationships must be evaluated.
@@
 By transforming the variables as indicated previously logarithmically, we are able to see this linear relationship across all predictor values.
+-->
 
-4.
+<!-- 4.
 **Homoskedastic Errors:** This assumption is examined by looking for constant error variance across the entire range of the x's (homoskedastic errors across the range of the x's).
@@
 Therefore model 2 and model both meet this condition when including the transformed variables.
+-->
 
-5.
+<!-- 5.
 **Normally Distributed Errors:** When investigating a normal distribution of the residuals, the QQ plot is helpful to investigate this assumption.
@@
 In conclusion, we meet this assumption with both model 2 and model 3.
+-->
 
-```{r CLM assumptions check, echo=FALSE}
+<!-- ```{r CLM assumptions check, echo=FALSE} -->
 
-#run this to display the CLM graphs needed to verify CLM assumptions met
-par(mfrow = c(2, 2))
-plot(model2, col="blue")
-title("Model 2", line = -0.75, outer = TRUE)
+<!-- #run this to display the CLM graphs needed to verify CLM assumptions met -->
+<!-- par(mfrow = c(2, 2)) -->
+<!-- plot(model2, col="blue") -->
+<!-- title("Model 2", line = -0.75, outer = TRUE) -->
 
-# Model 3
-# par(mfrow = c(2, 2))
-# plot(model3, col="green")
-# title("Model 3", line = -0.75, outer = TRUE)
+<!-- # Model 3 -->
+<!-- # par(mfrow = c(2, 2)) -->
+<!-- # plot(model3, col="green") -->
+<!-- # title("Model 3", line = -0.75, outer = TRUE) -->
 
-```
+<!-- ``` -->
 
-# Omitted Variable
+<!-- # Omitted Variable -->
 
-### Vehicle Ownership
+<!-- ### Vehicle Ownership -->
 
-One example of an omitted variable that may cause bias in our model is that of vehicle ownership among the population.
+<!-- One example of an omitted variable that may cause bias in our model is that of vehicle ownership among the population.
 It can certainly be expected that individuals that own a vehicle would be able to more easily travel short or long distances more frequently with relative ease.
@@
 From the data available, there would no immediate proxy for this omitted variable.
+-->
 
-### Fuel Prices
+<!-- ### Fuel Prices -->
 
-Another example of an omitted variable that may also cause bias in our model is fuel prices.
+<!-- Another example of an omitted variable that may also cause bias in our model is fuel prices.
 Whether this is jet fuel or gasoline, fluctuations in this commodity would certainly impact travel costs.
@@
 Conversely, if fuel prices for the time period of the sample for the population were lower, the bias would would be farther from zero.
+-->
 
-# Conclusion
+<!-- # Conclusion -->
 
-The goal of this analysis was to determine if there existed a causal relationship between distance traveled and Covid-19 infection rates for residents of the state of Texas.
-Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model.
+<!-- The goal of this analysis was to determine if there existed a causal relationship between distance traveled and COVID-19 infection rates for residents of the state of Texas.
+Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model, though with all variables considered, our results do not indicate vacation and COVID cases are necessarily related.
+-->
 
-### Next Steps
+<!-- <<<<<<< HEAD -->
+<!-- The goal of this analysis was to determine if there existed a causal relationship between distance traveled and COVID-19 infection rates for residents of the state of Texas.
+Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model, though with all variables considered, our results do not indicate vacation and COVID cases are necessarily related.
+-->
+<!-- ======= -->
+<!-- ### Next Steps -->
+<!-- >>>>>>> 6ca246330f02d2b1f157dc724cc9cf099142c8b0 -->
 
-For future analysis, it is recommended that Covid-19 policy adherence data per county be incorporated into the model to increase the accuracy.
-While the Govenor of Texas issued detailed protocols or mandates for the state of Texas, the degree to which each county (and thus the individual residents) complied would be useful data to have if such data were to exist.
+<!-- For future analysis, it is recommended that COVID-19 policy adherence data per county be incorporated into the model to increase the accuracy.
+While the Governor of Texas issued detailed protocols or mandates for the state of Texas, the degree to which each county (and thus the individual residents) complied would be useful data to have if such data were to exist.
 Additionally, incorporating vaccination rates could aid in development of a model with even higher accuracy.
 Generally, vaccinations could increase confidence among individuals interested in traveling, especially if they have been in long periods of quarantine or government mandated lockdowns.
-Last, it is highly recommended that any model generated in the future continue to use the latest data available given the oscillation in vaccination rates, government policy, and new or known variants of the Covid-19 virus.
+Last, it is highly recommended that any model generated in the future continue to use the latest data available given the oscillation in vaccination rates, government policy, and new or known variants of the COVID-19 virus.
+-->

### 2021-08-05 notebooks/Draft_Report.Rmd
edits removing death and unemp language

@@
 
-For the analysis, we will create three models to understand if causal relationships exist between the outcome variable, the number of long distance trips, and the control variables per Texas county: income, population density, age, COVID-19 cases and deaths .
-We will use an indicator variable, rural counties, to compare performance of our models for both rural or urban counties.
-By creating models with these variables, we intend to answer our primary and secondary research questions to provide insight about how vacation travel has been impacted so that interested stakeholders can gain an understanding of how to improve the revenue declines in the travel industry.
+Our team also theorizes that individuals at higher risk of Covid-19 infection would travel less (e.g.
+older populations).
+For the analysis, we will create three models to understand if causal relationships exist between the outcome variable, the number of long distance trips, and the control variables per Texas county: income, population density, age, and COVID-19 cases .
+We will use an indicator variable, rural counties, to understand performance of our models for both rural or urban counties subsets.
+By creating causal models with these variables, we intend to answer our research questions to provide insight about how vacation travel has been impacted so that interested stakeholders can gain an understanding of how to improve the revenue declines in the travel industry in the near future.
 
@@
 All of the data originated from primary sources at the Federal or State level, but there was no readily available documentation to map columnar data back to its primary source.
-We selected five measurements to support our causal theory: county population and density, median household income, number of individuals in poverty, and the unemployment rate.
-The economic measurements of poverty and employment were hypothesized to have an inverse relationship with enabling people to take long vacation trips, where median income should have a more direct relationship.
+We selected five measurements to support our causal theory: county population and density, median household income, $$number of individuals in poverty, and the unemployment rate.
+The economic measurements of poverty and employment were hypothesized to have an inverse relationship with enabling people to take long vacation trips, where median income should have a more direct relationship$$.
 The county population served as a control for the economic variables and density was hypothesized to encourage longer trips to more open spaces.
@@
 To enrich our model and limit OVB, we brainstormed a number of additional variables that could affect travel rates during COVID.
-For one, we looked at how COVID affected the county, through the unemployment rate.
+For one, we looked at how COVID affected the county, $$through the unemployment rate$$.
 We also looked at age, gender, income demographics.
@@
 From Figure 1, it can be observed that some of the variables are not distributed normally, are skewed,  and thus require transformations.
-More specifically, these variables are: Total Population, Total Cases, Total Deaths, Population Density.
-The outcome variable, Long Distance, will also require transformation.
+More specifically, these variables are: Total Population, Total Cases, Population Density.
+The outcome variable, long trips, will also require transformation.
 Upon further analysis, logarithmic transformations were chosen to normalize the distribution and minimize skewness.
@@
 table1::label(df$cases) <- "COVID-19 Cases"
-table1::label(df$deaths) <- "COVID-19 Deaths"
+#table1::label(df$deaths) <- "COVID-19 Deaths"
 # table1::label(df$unemployed) <- "Unemployment Count"
@@
 # table1::label(df$in_poverty) <- "Poverty Count"
-table1::label(df$lt17) <- "Age < 17 years"
-table1::label(df$gt65) <- "Age > 65 years"
+table1::label(df$lt17) <- "Age < 17 Years"
+table1::label(df$gt65) <- "Age > 65 Years"
 # table1::label(df$is_rural) <- "Type"
-table1::table1(~density + cases + deaths + median_income + lt17 + gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Characteristics", topclass="Rtable1-zebra", overall = "Total")
+table1::table1(~density + cases + median_income + lt17 + gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Characteristics", topclass="Rtable1-zebra", overall = "Total")
 # table1::table1(~density + cases + deaths + median_income + unemployed + in_poverty + lt17 + gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Characteristics", topclass="Rtable1-zebra", overall = "Total")
@@
 
-Below are histograms of variables that did not follow normal distribution and required transformation. Each of these variables were logarithmically transformed.
+Below are histograms of variables that did not follow normal distribution and required transformation.
+Each of these variables were logarithmically transformed, which can be seen in the collective figure (GGally).
 
@@
 attach(df)
-par(mfrow=c(3,2))
-hist(cases, main="Histogram of COVID-19 Cases")
-hist(deaths, main="Histogram of COVID-19 Deaths")
-hist(long_trip, main="Histogram of Trip Count")
-hist(density, main="Histogram of Population Density")
-hist(unemployed, main="Histogram of Unemployment")
-hist(median_income, main="Histogram of Median Income")
+par(mfrow=c(2,2))
+hist(cases, main="Figure 1: Histogram of COVID-19 Cases")
+#hist(deaths, main="Histogram of COVID-19 Deaths")
+hist(long_trip, main="Figure 2: Histogram of Trip Count")
+hist(density, main="Figure 3: Histogram of Population Density")
+#hist(unemployed, main="Histogram of Unemployment")
+hist(median_income, main="Figure 4: Histogram of Median Income ($)")
 # summary((deaths+1)^(1/2))
@@
 
-Upon transformation, below is a condensed view of the statistics of each variable:
+Upon logarithmic transformation, below is a condensed view of the statistics of each variable:
 
@@
                aes(color = County),
-               title = "Figure 1: Variable Characterstics",
+               title = "Figure 5: Variable Characterstics",
                upper = list(continuous = wrap("cor", size = 2.5)),
@@
 df2$log_cases <-log10(df2$cases+1)
-df2$log_deaths <-log10(df2$deaths+1)
+#df2$log_deaths <-log10(df2$deaths+1)
 df2$log_long_trip <-log10(df2$long_trip+1)
@@
 colnames(accounting_df) <- c('Step','Number of Samples','Samples Removed', 'Reason')
-knitr::kable(accounting_df, caption="Accounting Table", align = "c")
+knitr::kable(accounting_df, caption="Table 2: Accounting Table", align = "c")
 
@@
 colnames(final_tab_df) <- c('County Type', 'Number of Samples')
-knitr::kable(final_tab_df, caption="Table 2: Final Samples by County Type", align = "c")
+knitr::kable(final_tab_df, caption="Table 3: Final Samples by County Type", align = "c")
 ```
@@
 
-$Model_1: f(\text{log(Long Trip Count)}) = \beta_0 + \beta_1 g(\text{log(COVID-19 Cases)}) + \beta_2 g(\text{log(COVID-19 Deaths)})  + \beta_3 h(\text{Rural County})$
+$Model_1: f(\text{log(Long Trip Count)}) = \beta_0 + \beta_1 g(\text{log(COVID-19 Cases)}) + \beta_2 g(\text{Rural County}))$
 
@@
 
-$Model_2: f(\text{log(Long Trip Count)}) = \beta_0 + \beta_1 g(\text{log(COVID-19 Cases)}) + \beta_2 g(\text{log(COVID-19 Deaths)})  + \beta_3 h(\text{Rural County}) + \beta_4 h(\text{Median Income}) + \beta_5 h(\text{County Density})$
+$Model_2: f(\text{log(Long Trip Count)}) = \beta_0 + \beta_1 g(\text{log(COVID-19 Cases)}) + \beta_2 g(\text{Rural County})) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{County Density})$
 
@@
 
-$Model_3: f(\text{log(Long Trip Count)}) = \beta_0 + \beta_1 g(\text{log(COVID-19 Cases)}) + \beta_2 g(\text{log(COVID-19 Deaths)})  + \beta_3 h(\text{Rural County}) + \beta_4 h(\text{Median Income}) + \beta_5 h(\text{County Density}) + \beta_6 h(\text{Population under 17}) + \beta_7 h(\text{Population over 65})$
+$Model_3: f(\text{log(Long Trip Count)}) = \beta_0 + \beta_1 g(\text{log(COVID-19 Cases)}) + \beta_2 g(\text{Rural County})) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{County Density}) + \beta_5 h(\text{Population under 17}) + \beta_6 h(\text{Population over 65})$
 
@@
 This is calculated as e^(Constant).
-* In $Model_1$, for every percent increase in both cases and deaths, we are getting almost half a percent increase in trips.
-The positive relationship doesnt make much sense, so we will look to the other models.
+* In $Model_1$, for every percent increase in cases, we are getting almost half a percent increase in trips. The positive relationship doesnt make much sense, so we will look to the other models.
 * In $Model_2$, we see that density has a more positive relationship with trips, with every percent increase in county density modeled to a half percent increase in trips.
-Also the positive relationship between COVID cases/ deaths and trips has decreased to resulting in 0.15 and 0.25% increase in trips respectively.
-* In $Model_3$, COVID cases and deaths become insignificant.
-The coefficients have become slightly negative as in increase in cases and deaths decreases trips, which is in line with our original hypothesis, but is not significant.
+Also the positive relationship between COVID cases and trips has decreased to resulting in 0.15 and 0.25% increase in trips respectively.
+* In $Model_3$, COVID cases become insignificant.
+The coefficients have become slightly negative as in increase in cases decreases trips, which is in line with our original hypothesis, but is not significant.
 Instead, for every percent median income of the county increases by, the amount of long trips decrease by 0.3%.
@@
 
-DELTE*** Not CLM since >100 samples. Large sample assumptions = IID and unique BLP (which is related to colinearity...)
+Below is an analysis of the Classin Linear Model assumptions for $Model_2$ and $Model_3$:
 
@@
 The data was gathered from independent surveys within the Texas population for all models generated across all counties.
+There may be instances of sample clustering given that individuals in counties may have influenced individuals in neighboring counties.
+However, the sample size for the analysis is sufficient to minimize this possible impact.
 
@@
 **No Perfect Colinearity:** This can be assessed through examination of the variance inflation factor (VIF).
-Running a VIF in R for model 2 and model 3, we see the VIF scores are relatively low (below 5) for the predictor variables.
-Therefore, no predictor variables were removed from the model.
-The results for the VIF analysis are seen in the table below:
+The results for the VIF analysis are seen in the table below.
+Running a VIF in R for $Model_2$and $Model_3$, we see the VIF scores are relatively low (below 5) for the predictor variables in $Model_2$.
+However, upon review of $Model_3$, this is not the case.
+With VIF factors greater than 5, the both age variables would be candidates for removal outright.
+When reviewing the population density variable in $Model_3$, it remains significant for $Model_2$.
+Additionally, when viewing the variable relationships between the two age variables and population in Figure 5 (GGally), strong linear relationships can be viewed which would also indicate almost near perfect colinearity and thus should be removed from the model.
 

### 2021-08-05 notebooks/Draft_Report.Rmd
edits on CLM, formattinge

@@
 **Linear Conditional Expectation:** For this assumption, the residuals for linear relationships must be evaluated.
-By plotting the residuals, as seen below in the first scatter plot for both model 2 and model 3 ('Residual vs Fitted').
+By plotting the residuals, as seen below in the first scatter plot for both $Model_2$ and $Model_3$ ('Residual vs Fitted').
 We generally see a linear line as we move across predictor values with the residuals with no strong pattern.
@@
 By plotting fitted values vs.
-the square root of the standardized residuals for the model 2 and model 3 ('Scale-Location' figure), there is an even or equal spread dispersion across the x values.
-Therefore model 2 and model both meet this condition when including the transformed variables.
+the square root of the standardized residuals for the $Model_2$ and model 3 ('Scale-Location' figure), there is an even or equal spread dispersion across the x values.
+Therefore $Model_2$ and $Model_3$ both meet this condition when including the transformed variables.
 -->
@@
 **Normally Distributed Errors:** When investigating a normal distribution of the residuals, the QQ plot is helpful to investigate this assumption.
-For model 2 and model 3, a nice linear line can be seen between the theoretical quantities and the standardized residuals ('Normal Q-Q').
-In conclusion, we meet this assumption with both model 2 and model 3.
+For $Model_2$ and $Model_3$, a nice linear line can be seen between the theoretical quantities and the standardized residuals ('Normal Q-Q').
+In conclusion, we meet this assumption with both $Model_2$ and $Model_3$.
 -->
@@
 
-<!-- <<<<<<< HEAD -->
-<!-- The goal of this analysis was to determine if there existed a causal relationship between distance traveled and COVID-19 infection rates for residents of the state of Texas.
-Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model, though with all variables considered, our results do not indicate vacation and COVID cases are necessarily related.
--->
-<!-- ======= -->
 <!-- ### Next Steps -->
-<!-- >>>>>>> 6ca246330f02d2b1f157dc724cc9cf099142c8b0 -->
 

### 2021-08-05 notebooks/Draft_Report.Rmd
edits on formatting

@@
 
-Our team also theorizes that individuals at higher risk of Covid-19 infection would travel less (e.g.
+Our team also theorizes that individuals at higher risk of COVID-19 infection would travel less (e.g.
 older populations).
@@
 
-The NYT has maintained a robust public dataset of county-level, time-series COVID data on GitHub and holds records dating back to the beginning of COVID-19.
+The NYT has maintained a robust public dataset of county-level, time-series COVID-19 data on GitHub and holds records dating back to the beginning of COVID-19.
 The dataset aggregates state and municipal government health department reports of cumulative cases and deaths for each US county daily from COVID-19.
@@
 __integrating below still__
-We gathered COVID case and death data from NYT.
+We gathered COVID-19 case and death data from NYT.
 We gathered trip distance data from the Bureau of Transportation, which they record based on data gathered from personal mobile devices.
@@
 #### Operationalization of additional variables
-To enrich our model and limit OVB, we brainstormed a number of additional variables that could affect travel rates during COVID.
-For one, we looked at how COVID affected the county, $$through the unemployment rate$$.
+To enrich our model and limit OVB, we brainstormed a number of additional variables that could affect travel rates during COVID-19.
+For one, we looked at how COVID-19 affected the county, $$through the unemployment rate$$.
 We also looked at age, gender, income demographics.
-Then, in addition to income demographics, we focused on percent of younger folks, as defined as those younger than 17 and older folks as defined as those older than 65 because of the nature of how COVID affects an individual.
+Then, in addition to income demographics, we focused on percent of younger folks, as defined as those younger than 17 and older folks as defined as those older than 65 because of the nature of how COVID-19 affects an individual.
 
@@
 In the second model, we added additional county characteristics.
-In the final model, we added in another COVID related factor.
+In the final model, we added in another COVID-19 related factor.
 
 Here is the first model, which is limited to study only key variables of interest for measurement based on our hypotheses.
-Our key input variable is COVID cases and key output is number of trips greater than 100 miles.
+Our key input variable is COVID-19 cases and key output is number of trips greater than 100 miles.
 We also chose to add an indicator for rural counties because it could be directly related to trip distance
@@
 
-Finally in the third model, we added in age demographics as percent under 17/over 65 as another potential source that influences travel midst COVID cases due to the nature of how COVID affects those at different ages.
+Finally in the third model, we added in age demographics as percent under 17/over 65 as another potential source that influences travel midst COVID-19 cases due to the nature of how the virus affects those at different ages.
 
@@
 First, we consider the models holistically. Here are some of our observations:
-* In model1, all of our variables are extremely significant, though in the opposite direction as expected.
-We would expect an inverse relationship between COVID and travel, though it seems that this model is producing a positive relationship.
-* In model2, when county density is factored in, the presence of COVID seems to have less positive of an effect on travel and that higher density counties see more travel.
-* In model3 with age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17.
+* In $Model_1$, all of our variables are extremely significant, though in the opposite direction as expected.
+We would expect an inverse relationship between COVID-19 cases and travel, though it seems that this model is producing a positive relationship.
+* In $Model_2$, when county density is factored in, the presence of COVID-19 seems to have less positive of an effect on travel and that higher density counties see more travel.
+* In $Model_3$ with age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17.
 * In terms of $R^2$ and RSE, we are improving with each model.
@@
 
+$$RECHECK THIS SECTION FOR VALUES$$
 We will also interpret the coefficient values and what these mean in the context of our model.
@@
 * In $Model_2$, we see that density has a more positive relationship with trips, with every percent increase in county density modeled to a half percent increase in trips.
-Also the positive relationship between COVID cases and trips has decreased to resulting in 0.15 and 0.25% increase in trips respectively.
-* In $Model_3$, COVID cases become insignificant.
+Also the positive relationship between COVID-19 cases and trips has decreased to resulting in 0.15 and 0.25% increase in trips respectively.
+* In $Model_3$, COVID-19 cases become insignificant.
 The coefficients have become slightly negative as in increase in cases decreases trips, which is in line with our original hypothesis, but is not significant.
 Instead, for every percent median income of the county increases by, the amount of long trips decrease by 0.3%.
-This could indicate that more affluent counties trust science and err on the safe side in the midst of COVID.
+This could indicate that more affluent counties trust science and err on the safe side in the midst of COVID-19.
 Finally, in our 3rd model, counties with more people under 17 years of age are more likely to take long trips.
@@
 <!-- The goal of this analysis was to determine if there existed a causal relationship between distance traveled and COVID-19 infection rates for residents of the state of Texas.
-Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model, though with all variables considered, our results do not indicate vacation and COVID cases are necessarily related.
+Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model, though with all variables considered, our results do not indicate vacation and COVID-19 cases are necessarily related.
 -->

### 2021-08-05 notebooks/Draft_Report.Rmd
added r code for % coeff calc for Lucy

@@
 $$RECHECK THIS SECTION FOR VALUES$$
+```{r calc. coeff percents}
+#Model 1
+#when you compare a coeff for non-log x to a log y:
+(1-exp(coef(model1)["is_rural"]))*100
+#For every percent increase in x 'is rural', you get a #% outcome in number of long trips.
+
+#Model 2
+(1-exp(coef(model2)["is_rural"]))*100
+#Model 3
+(1-exp(coef(model3)["is_rural"]))*100
+
+#Same for income
+
+#Model 2
+(1-exp(coef(model2)["log(median_income + 1)"]))*100
+#Model 3
+(1-exp(coef(model3)["log(median_income + 1)"]))*100
+
+#Same for density
+
+#Model 2
+(1-exp(coef(model2)["log(density)"]))*100
+#Model 3
+(1-exp(coef(model3)["log(density)"]))*100
+
+#Same for age 17 & 65
+#Model 3
+(1-exp(coef(model3)["log(lt17)"]))*100
+(1-exp(coef(model3)["log(gt65)"]))*100
+```
+
 We will also interpret the coefficient values and what these mean in the context of our model.

### 2021-08-05 notebooks/Draft_Report.Rmd
decorations and minor fixes

--- 2021-08-05 notebooks/Draft_Report.Rmd
@@
 
-Table 1 below displays some information about the mean, median, minimum and maximum values for each variable.
-Figure 1 below depicts the distribution and correlation of each variable in our dataset by county type (where '0' indicates an urban county).
-From Figure 1, it can be observed that some of the variables are not distributed normally, are skewed,  and thus require transformations.
+Table \@ref(tab:var-summary) below displays some information about the mean, median, minimum and maximum values for each variable.
+Figure \@ref(fig:ggally) below depicts the distribution and correlation of each variable in our dataset by county type (where '0' indicates an urban county).
+From Figure \@ref(fig:ggally), it can be observed that some of the variables are not distributed normally, are skewed,  and thus require transformations.
 More specifically, these variables are: Total Population, Total Cases, Population Density.
-The outcome variable, long trips, will also require transformation.
+The outcome variable, `long_trips`, will also require the same transformation.
 Upon further analysis, logarithmic transformations were chosen to normalize the distribution and minimize skewness.
 
-```{r EDA, echo=FALSE, message=FALSE}
+```{r var-summary, echo=FALSE, message=FALSE, fig.cap='Texas County Variable Characteristics', fig.pos='!b'}
 #Example Table
@@
 # table1::label(df$is_rural) <- "Type"
-table1::table1(~density + cases + median_income + lt17 + gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Characteristics", topclass="Rtable1-zebra", overall = "Total")
+table1::table1(~density + cases + median_income + lt17 + gt65 | rural_cat, data = df, topclass="Rtable1-zebra", overall = "Total")
 # table1::table1(~density + cases + deaths + median_income + unemployed + in_poverty + lt17 + gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Characteristics", topclass="Rtable1-zebra", overall = "Total")
@@
 Below are histograms of variables that did not follow normal distribution and required transformation.
-Each of these variables were logarithmically transformed, which can be seen in the collective figure (GGally).
+We applied a logarithmic transform to each of these variables, which can be seen in the collective Figure \@ref(tab:var-plots).
+*why logs*
 
-```{r var plots, fig.align='center', message=FALSE, echo=FALSE}
+```{r var-plots, fig.align='center', message=FALSE, echo=FALSE}
 attach(df)
@@
 
-```{r ggally, echo=FALSE, message=FALSE}
+```{r ggally, echo=FALSE, message=FALSE, fig.cap='Variable Characterstics', fig.pos='!b'}
 County = as.factor(df$rural_cat)
@@
   mutate(
+    # log(x+1) to preserve negative and zero valued samples
     cases = log(cases+1),
@@
                aes(color = County),
-               title = "Figure 5: Variable Characterstics",
+               # title = "Figure 5: Variable Characterstics",
                upper = list(continuous = wrap("cor", size = 2.5)),
@@
 
-To demonstrate the operationalization of the data, Tables 1 and 2 below display the original value counts from the dataset as well as how many values were removed as a part of filtering the dataset to remove all states except Texas.
+To demonstrate the operationalization of the data, Tables \@ref(tab:accounting-table) and \@ref(tab:number-of-samples) below display the original value counts from the dataset as well as how many values were removed as a part of filtering the dataset to remove all states except Texas.
 For the analysis, 230 responses were analyzed as shown in Table 1.
@@
 
-```{r Accounting Table creation & output, echo=FALSE}
+```{r accounting-table, echo=FALSE, fig.cap='Variable Characterstics', fig.pos='!b'}
 df2 <- df
@@
 colnames(accounting_df) <- c('Step','Number of Samples','Samples Removed', 'Reason')
-knitr::kable(accounting_df, caption="Table 2: Accounting Table", align = "c")
+knitr::kable(accounting_df, align = "c")
+```
 
+```{r, number-of-samples, fig.cap='Number of Samples', fig.pos='!b'}
 final_tab_df = dplyr::count(df2, rural_cat, sort = TRUE)
 colnames(final_tab_df) <- c('County Type', 'Number of Samples')
-knitr::kable(final_tab_df, caption="Table 3: Final Samples by County Type", align = "c")
+knitr::kable(final_tab_df, align = "c")
 ```
@@
 The data was gathered from independent surveys within the Texas population for all models generated across all counties.
-There may be instances of sample clustering given that individuals in counties may have influenced individuals in neighboring counties.
+There may be instances of sample clustering given that individuals in counties may have influenced individuals in neighboring counties as well as being subject to statewide restrictions on implementation of COVID transmission mitigration strategies (e.g., mask mandates)[^no_mask].
 However, the sample size for the analysis is sufficient to minimize this possible impact.
+
+[^no_mask]: https://gov.texas.gov/news/post/governor-abbott-issues-executive-order-prohibiting-government-entities-from-mandating-masks
 
@@
 The results for the VIF analysis are seen in the table below.
-Running a VIF in R for $Model_2$and $Model_3$, we see the VIF scores are relatively low (below 5) for the predictor variables in $Model_2$.
+Running a VIF in R for $Model_2$ and $Model_3$, we see the VIF scores are relatively low (below 5) for the predictor variables in $Model_2$.
 However, upon review of $Model_3$, this is not the case.

### 2021-08-05 notebooks/Draft_Report.Rmd

@@
 First, we consider the models holistically. Here are some of our observations:
-* In model1, all of our variables are extremely significant, though in the opposite direction as expected.
-We would expect an inverse relationship between COVID and travel, though it seems that this model is producing a positive relationship.
-* In model2, when county density is factored in, the presence of COVID seems to have less positive of an effect on travel and that higher density counties see more travel.
-* In model3 with age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17.
+* In $Model_1$, all of our variables are extremely significant, though in the opposite direction as expected.
+We would expect an inverse relationship between COVID-19 cases and travel, though it seems that this model is producing a positive relationship.
+* In $Model_2$, when county density is factored in, the presence of COVID-19 seems to have less positive of an effect on travel and that higher density counties see more travel.
+* In $Model_3$ with age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17.
 * In terms of $R^2$ and RSE, we are improving with each model.
 * The addition of our chosen parameters for each following model is also statistically better than the previous, as indicated by the F statistic.
+
+$$RECHECK THIS SECTION FOR VALUES$$
+```{r calc. coeff percents}
+#Model 1
+#when you compare a coeff for non-log x to a log y:
+(1-exp(coef(model1)["is_rural"]))*100
+#For every percent increase in x 'is rural', you get a #% outcome in number of long trips.
+
+#Model 2
+(1-exp(coef(model2)["is_rural"]))*100
+#Model 3
+(1-exp(coef(model3)["is_rural"]))*100
+
+#Same for income
+
+#Model 2
+(1-exp(coef(model2)["log(median_income + 1)"]))*100
+#Model 3
+(1-exp(coef(model3)["log(median_income + 1)"]))*100
+
+#Same for density
+
+#Model 2
+(1-exp(coef(model2)["log(density)"]))*100
+#Model 3
+(1-exp(coef(model3)["log(density)"]))*100
+
+#Same for age 17 & 65
+#Model 3
+(1-exp(coef(model3)["log(lt17)"]))*100
+(1-exp(coef(model3)["log(gt65)"]))*100
+```
 

### 2021-08-05 notebooks/Draft_Report.Rmd
more edits

--- 2021-08-05 notebooks/Draft_Report.Rmd
@@
 Summer vacations in the United States are synonymous with sunny, leisurely days spent with loved ones.
-As Ella Fitzgerald so aptly sings, “Summertime, and the living is easy.” However, travel in the recent peak-vacation months has certainly deviated from this idealistic frame of mind with the onset of the COVID-19 pandemic in early 2020.
+As Ella Fitzgerald so aptly sings, “Summertime, and the living is easy.” However, travel in the recent peak-vacation months has certainly deviated from this idealistic frame of mind with the onset of the SARS-CoV-2 (COVID-19) pandemic in early 2020.
 The global travel industry has experienced sharp declines in Revenue.
-Specifically in the United States, “travel spending totaled a mere $679 billion in 2020, an unprecedented 42% annual decline (nearly $500 billion) from 2019,” according to analysis from the U.S.
-Travel Association.
-Our hypothesis as to why this decline has occurred is due directly to the presence of the COVID-19 virus and the number of cases.
+Specifically in the United States, “travel spending totaled a mere \$679 billion in 2020, an unprecedented 42% annual decline (nearly \$500 billion) from 2019,” according to analysis from the U.S.
+Travel Association[^usta].
+Our hypothesis as to why this decline has occurred is due directly to the volume of COVID-19 number of cases.
 As virus infection oscillate regionally in the United States, our team of data scientists will look to answer the following primary research question using linear regression statistical techniques:
@@
 Our primary hypothesis is that there is a causal relationship between the number of COVID-19 cases and the number of vacation trips; when COVID-19 cases increase, the number of vacation travel trips decrease.
-To investigate, a dataset comprising of Texas county data will be analyzed.
+To investigate, several demographic, mobility, and COVID-19 data sources will be analyzed.
 Based on the data, our team will also investigate the following secondary question:
 
-Do income, population density, age, and/or county type (rural or urban) per county impact traveling for Texas residents?
+Does the median income, the population density, and/or the age by county type (rural or urban) impact vacation travel for Texas residents in June, 2021?
 
-Our team also theorizes that individuals at higher risk of Covid-19 infection would travel less (e.g.
-older populations).
-For the analysis, we will create three models to understand if causal relationships exist between the outcome variable, the number of long distance trips, and the control variables per Texas county: income, population density, age, and COVID-19 cases .
+Our team also theorizes that individuals at higher risk of COVID-19 (e.g.
+older populations) infection would travel less than younger populations.
+For the analysis, we will create three models to understand if causal relationships exist between the outcome variable, the number of long distance leisure trips, and the control variables for each Texas county: income, population density, age, and COVID-19 cases .
 We will use an indicator variable, rural counties, to understand performance of our models for both rural or urban counties subsets.
 By creating causal models with these variables, we intend to answer our research questions to provide insight about how vacation travel has been impacted so that interested stakeholders can gain an understanding of how to improve the revenue declines in the travel industry in the near future.
+
+[^usta]:https://www.ustravel.org/toolkit/covid-19-travel-industry-research
 
@@
 
-For our investigation, we selected on cases and deaths in the month of June 2021, to best operationalize summer.
-The case and death counts were aggregated by county to mitigate time-series affects on modeling.
-The death counts were identified as being directly related to the COVID-19 infection and not from other instances (e.g., homicide), or unrelated comorbidities (e.g., cancer).
+For our investigation, we selected on cases, aggregated by county, in the month of June 2021 to mitigate time-series effects on modeling.
+The death count measurements were not selected because of the high correlation with the case count measurements: deaths counts are a subset of case counts.
 
@@
 All of the data originated from primary sources at the Federal or State level, but there was no readily available documentation to map columnar data back to its primary source.
-We selected five measurements to support our causal theory: county population and density, median household income, $$number of individuals in poverty, and the unemployment rate.
-The economic measurements of poverty and employment were hypothesized to have an inverse relationship with enabling people to take long vacation trips, where median income should have a more direct relationship$$.
+We selected four measurements to support our causal theory: county density, median household income, and populations aged over 65 years and under 17 years.
+The economic measurement of median income should have a significant role in an individual's ability to take vacations.
 The county population served as a control for the economic variables and density was hypothesized to encourage longer trips to more open spaces.
@@
 While the vast database has many of the measurements useful to support our causal theory, as the TAC database queries held a majority of these, the decennial survey did proffer a means to generate an indicator variable identifying rural or urban counties.
-This variable was important since we hypothesized that rural counties may experience higher long trip counts due to their geographies.
 
-Our final course were data produced by the US Department of Transportation's Bureau of Transportation Statistics (USBTS).
-These data are experimental *estimates* of mobility gathrered from aggregated mobile device data.
+Our final final data source was produced by the US Department of Transportation's Bureau of Transportation Statistics (USBTS).
+These data are experimental *estimates* of mobility gathered from aggregated mobile device data.
 USBTS states that these data are experimental, and but significant effort was made to ensure proper population weighting, and data multi-sourcing (to reduce variance in spacial and temporal measurements common from a single source) were implemented before any of the statistics were calculated.
@@
 Trips are not distiguish by mode of transporation (i.e., air, train, bus, car, or bike).
-For our analyis, we were interested in only trips longer than 100 miles in Texas in June of 2021 as this best operationalize the concept of a vacation[^bts01].
+For our analysis, we were interested in only trips longer than 100 miles in Texas in June of 2021 as this best operationalized the concept of vacation travel[^bts01].
+
+USBTS conducted a survey[^bts01] that characaterized long distance travel as trips over 50 miles from home.
+The distriubtion of trips in that survey included leisure, business, and commuter travel.
+To operationalize out outcome variable more concretely, we focused the dataset on trips over 100 miles to reduced noise from non-leisure travel[^brookings].
 
@@
 Retrieved 07 19, 2021, from https://data.bts.gov/Research-and-Statistics/Trips-by-Distance/w96p-f2qv
+[^brookings]: https://www.brookings.edu/wp-content/uploads/2016/07/Srvy_JobsProximity.pdf
 
 ---
-#### Operationalization of focus variables
-
-__integrating below still__
-We gathered COVID case and death data from NYT.
-We gathered trip distance data from the Bureau of Transportation, which they record based on data gathered from personal mobile devices.
-This data was presented as number of travelers in binned distances.
-We defined a short trip as less than 100 miles and long trips as greater than 100 miles.
-This was decided under the assumption that distances traveled of less than 100 miles are for the purpose of commuting, while greater than 100 miles indicated vacation travel.
-Hand in hand with this variable is our indicator variable for urban vs rural setting *** how did we decide this again, population threshold?**** under the assumption that those who live in rural settings are likely to commute further than those in an urban setting.
-Finally, we believe that median household income plays a significant role in an individual's ability to take vacations.
-We pulled this directly from the US Census Bureau.
 
@@
 
+```{r var-description}
+
+table(
+  source = c(1, 2, 3),
+  description = c("a", "b", "c"),
+  variable = c("A", "B", "C")
+)
+```
+
 ```{r var-summary, echo=FALSE, message=FALSE, fig.cap='Texas County Variable Characteristics', fig.pos='!b'}
-#Example Table
 df$rural_cat <- ifelse(df$is_rural == 1, "Rural", "Urban")
+table1::label(df$long_trip) <- "Vacation Trip Count"
+table1::label(df$cases) <- "COVID-19 Cases"
+table1::label(df$median_income) <- "Median Income ($)"
 table1::label(df$density) <- "Population Density"
-table1::label(df$cases) <- "COVID-19 Cases"
-#table1::label(df$deaths) <- "COVID-19 Deaths"
-# table1::label(df$unemployed) <- "Unemployment Count"
-table1::label(df$median_income) <- "Median Income ($)"
-# table1::label(df$in_poverty) <- "Poverty Count"
 table1::label(df$lt17) <- "Age < 17 Years"
 table1::label(df$gt65) <- "Age > 65 Years"
-# table1::label(df$is_rural) <- "Type"
-table1::table1(~density + cases + median_income + lt17 + gt65 | rural_cat, data = df, topclass="Rtable1-zebra", overall = "Total")
-# table1::table1(~density + cases + deaths + median_income + unemployed + in_poverty + lt17 + gt65 | rural_cat, data = df, caption = "Table 1: Texas County Variable Characteristics", topclass="Rtable1-zebra", overall = "Total")
+table1::table1(~long_trip + density + cases + median_income + lt17 + gt65 | rural_cat, data = df, topclass="Rtable1-zebra", overall = "Total")
 ```
@@
 We applied a logarithmic transform to each of these variables, which can be seen in the collective Figure \@ref(tab:var-plots).
-*why logs*
 
@@
 attach(df)
-par(mfrow=c(2,2))
-hist(cases, main="Figure 1: Histogram of COVID-19 Cases")
-#hist(deaths, main="Histogram of COVID-19 Deaths")
-hist(long_trip, main="Figure 2: Histogram of Trip Count")
-hist(density, main="Figure 3: Histogram of Population Density")
-#hist(unemployed, main="Histogram of Unemployment")
-hist(median_income, main="Figure 4: Histogram of Median Income ($)")
-# summary((deaths+1)^(1/2))
+par(mfrow=c(3,2))
+hist(long_trip, main="Vacation Trip Count")
+hist(cases, main="COVID-19 Cases")
+hist(median_income, main="Median Income ($)")
+hist(density, main="Population Density")
+hist(lt17, main="Age < 17 Years")
+hist(gt65, main="Age > 65 Years")
 ```
@@
     cases = log(cases+1),
-    # deaths = log(deaths+1),
     long_trip = log(long_trip),
@@
     median_income=log(median_income+1),
-    # unemployed = log(unemployed+1),
     lt17 = log(lt17),
@@
                diag = list(continuous = wrap("densityDiag", alpha = 0.5)),
-               columnLabels=c("Distance", "Cases", "Income", "Pop Density", "Age < 17","Age > 65"))+
+               columnLabels=c("Distance", "Cases", "Income", "Density", "Age < 17","Age > 65"))+
                theme(legend.position = "bottom")
-
-# ggpairs(df, columns = 3:ncol(df), legend = 1, aes(color = County), title = "Figure 1: Variable Characterstics")
-# #                upper = list(continuous = wrap("cor", size = 2.5)),
-# #                lower = list(continuous = wrap("points", alpha = 0.5, size=0.3)),
-# #                diag = list(continuous = wrap("densityDiag", alpha = 0.5)))
 
@@
 
-To demonstrate the operationalization of the data, Tables \@ref(tab:accounting-table) and \@ref(tab:number-of-samples) below display the original value counts from the dataset as well as how many values were removed as a part of filtering the dataset to remove all states except Texas.
-For the analysis, 230 responses were analyzed as shown in Table 1.
-Table 2 depicts the county type breakdown for rural or urban classification as well.
+To demonstrate the operationalization of the data, Table \@ref(tab:accounting-table) below display the original value counts from the dataset as well as how many values were removed as a part of filtering the dataset to remove all states except Texas.
+For the analysis, 230 responses were analyzed as shown in Table \@ref(tab:accounting-table).
 
@@
 
-```{r, number-of-samples, fig.cap='Number of Samples', fig.pos='!b'}
-final_tab_df = dplyr::count(df2, rural_cat, sort = TRUE)
-colnames(final_tab_df) <- c('County Type', 'Number of Samples')
-knitr::kable(final_tab_df, align = "c")
-```
-
-## Additional Changes to the Data:
+## Additional Modifications to the Data:
 
@@
 Outliers or incorrect data are not observed upon final inspection.
-The Accounting Table capture the changes made through cleaning and filtering as indicated previously.
-Table 2 above also displays the subset per county type for the overall dataset.
+The Accounting Table captures the changes made through cleaning, filtering, and transforming as indicated previously.
+With a sample size over 100, a large-sample linear model can be built.
 
@@
 To investigate the causal relationships between distance traveled and variables within the dataset, a linear regression was performed in R Studio for three models.
-The first model includes only our base variables.
-In the second model, we added additional county characteristics.
-In the final model, we added in another COVID related factor.
+The first model includes only our the key variables associated with our primary hypothesis: that COVID-19 cases decrease vacation trip counts.
+In the second model, we added additional county econometric and demographic characteristics.
+In the final model, we added additional demographic variables for age groups.
 
-Here is the first model, which is limited to study only key variables of interest for measurement based on our hypotheses.
-Our key input variable is COVID cases and key output is number of trips greater than 100 miles.
-We also chose to add an indicator for rural counties because it could be directly related to trip distance
+## Model Equations
 
-$Model_1: f(\text{log(Long Trip Count)}) = \beta_0 + \beta_1 g(\text{log(COVID-19 Cases)}) + \beta_2 g(\text{Rural County}))$
+**Model 1: **
 
-The second model expands upon the first model to also include key explanatory variables and covariates that are hypothesized to improve performance of the previous model.
-Here, we added in median income and density of the county as additional descriptors of the county.
+$(\text{log(Long Trip Count)}) = \beta_0 + \beta_1(\text{log(COVID-19 Cases)}) + \beta_2(\text{Rural County}))$
 
-$Model_2: f(\text{log(Long Trip Count)}) = \beta_0 + \beta_1 g(\text{log(COVID-19 Cases)}) + \beta_2 g(\text{Rural County})) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{County Density})$
+**Model 2:**
 
-Finally in the third model, we added in age demographics as percent under 17/over 65 as another potential source that influences travel midst COVID cases due to the nature of how COVID affects those at different ages.
+$(\text{log(Long Trip Count)}) = \beta_0 + \beta_1(\text{log(COVID-19 Cases)}) + \beta_2(\text{Rural County})) + \beta_3(\text{log(Median Income)}) + \beta_4(\text{log(County Density)})$
 
-$Model_3: f(\text{log(Long Trip Count)}) = \beta_0 + \beta_1 g(\text{log(COVID-19 Cases)}) + \beta_2 g(\text{Rural County})) + \beta_3 h(\text{Median Income}) + \beta_4 h(\text{County Density}) + \beta_5 h(\text{Population under 17}) + \beta_6 h(\text{Population over 65})$
+**Model 3:**
+
+$(\text{log(Long Trip Count)}) = \beta_0 + \beta_1(\text{log(COVID-19 Cases)}) + \beta_2(\text{Rural County})) + \beta_3(\text{log(Median Income)}) + \beta_4(\text{log(County Density)}) + \beta_5(\text{log(Population under 17)}) + \beta_6(\text{log(Population over 65)})$
+
+\clearpage
 
@@
 The results from the linear regression analysis can be seen below for each model:
+
 ```{r model creation, echo=FALSE, warning=FALSE}
@@
 
-## Discussion
+## Results Discussion
+
 From this output table, we can discern quite a bit of information.
@@
 First, we consider the models holistically. Here are some of our observations:
-* In model1, all of our variables are extremely significant, though in the opposite direction as expected.
+* In `Model 1`, all of our variables are extremely significant, though in the opposite direction as expected.
 We would expect an inverse relationship between COVID and travel, though it seems that this model is producing a positive relationship.
-* In model2, when county density is factored in, the presence of COVID seems to have less positive of an effect on travel and that higher density counties see more travel.
-* In model3 with age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17.
+* In `Model 2`, when county density is factored in, the presence of COVID seems to have less positive of an effect on travel and that higher density counties see more travel.
+* In `Model 3` with age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17.
 * In terms of $R^2$ and RSE, we are improving with each model.
@@
 Regardless of any variables, we are getting between ~300 (as with model3) and over 8000 (as with model2) trips over 100 miles.
-This is calculated as e^(Constant).
-* In $Model_1$, for every percent increase in cases, we are getting almost half a percent increase in trips. The positive relationship doesnt make much sense, so we will look to the other models.
-* In $Model_2$, we see that density has a more positive relationship with trips, with every percent increase in county density modeled to a half percent increase in trips.
+This is calculated as $e^(Constant)$.
+* In `Model 1`, for every percent increase in cases, we are getting almost half a percent increase in trips. The positive relationship doesnt make much sense, so we will look to the other models.
+* In `Model 2`, we see that density has a more positive relationship with trips, with every percent increase in county density modeled to a half percent increase in trips.
 Also the positive relationship between COVID cases and trips has decreased to resulting in 0.15 and 0.25% increase in trips respectively.
-* In $Model_3$, COVID cases become insignificant.
+* In `Model 3`, COVID cases become insignificant.
 The coefficients have become slightly negative as in increase in cases decreases trips, which is in line with our original hypothesis, but is not significant.
@@
 
-Below is an analysis of the Classin Linear Model assumptions for $Model_2$ and $Model_3$:
+Below is an analysis of the Classin Linear Model assumptions for $Model 2$ and $Model 3$:
 
@@
 The results for the VIF analysis are seen in the table below.
-Running a VIF in R for $Model_2$ and $Model_3$, we see the VIF scores are relatively low (below 5) for the predictor variables in $Model_2$.
-However, upon review of $Model_3$, this is not the case.
+Running a VIF in R for $Model 2$ and $Model 3$, we see the VIF scores are relatively low (below 5) for the predictor variables in $Model 2$.
+However, upon review of $Model 3$, this is not the case.
 With VIF factors greater than 5, the both age variables would be candidates for removal outright.
-When reviewing the population density variable in $Model_3$, it remains significant for $Model_2$.
+When reviewing the population density variable in $Model 3$, it remains significant for $Model 2$.
 Additionally, when viewing the variable relationships between the two age variables and population in Figure 5 (GGally), strong linear relationships can be viewed which would also indicate almost near perfect colinearity and thus should be removed from the model.
@@
 
-<!-- 3.
+3.
 **Linear Conditional Expectation:** For this assumption, the residuals for linear relationships must be evaluated.
-By plotting the residuals, as seen below in the first scatter plot for both $Model_2$ and $Model_3$ ('Residual vs Fitted').
+By plotting the residuals, as seen below in the first scatter plot for both $Model 2$ and $Model 3$ ('Residual vs Fitted').
 We generally see a linear line as we move across predictor values with the residuals with no strong pattern.
 By transforming the variables as indicated previously logarithmically, we are able to see this linear relationship across all predictor values.
--->
 
-<!-- 4.
+4.
 **Homoskedastic Errors:** This assumption is examined by looking for constant error variance across the entire range of the x's (homoskedastic errors across the range of the x's).
 By plotting fitted values vs.
-the square root of the standardized residuals for the $Model_2$ and model 3 ('Scale-Location' figure), there is an even or equal spread dispersion across the x values.
-Therefore $Model_2$ and $Model_3$ both meet this condition when including the transformed variables.
--->
+the square root of the standardized residuals for the $Model 2$ and model 3 ('Scale-Location' figure), there is an even or equal spread dispersion across the x values.
+Therefore $Model 2$ and $Model 3$ both meet this condition when including the transformed variables.
 
-<!-- 5.
+5.
 **Normally Distributed Errors:** When investigating a normal distribution of the residuals, the QQ plot is helpful to investigate this assumption.
-For $Model_2$ and $Model_3$, a nice linear line can be seen between the theoretical quantities and the standardized residuals ('Normal Q-Q').
-In conclusion, we meet this assumption with both $Model_2$ and $Model_3$.
--->
+For $Model 2$ and $Model 3$, a nice linear line can be seen between the theoretical quantities and the standardized residuals ('Normal Q-Q').
+In conclusion, we meet this assumption with both $Model 2$ and $Model 3$.
 
-<!-- ```{r CLM assumptions check, echo=FALSE} -->
+```{r CLM assumptions check, echo=FALSE}
 
-<!-- #run this to display the CLM graphs needed to verify CLM assumptions met -->
-<!-- par(mfrow = c(2, 2)) -->
-<!-- plot(model2, col="blue") -->
-<!-- title("Model 2", line = -0.75, outer = TRUE) -->
+#run this to display the CLM graphs needed to verify CLM assumptions met
+par(mfrow = c(2, 2))
+plot(model2, col="blue")
+title("Model 2", line = -0.75, outer = TRUE)
 
-<!-- # Model 3 -->
-<!-- # par(mfrow = c(2, 2)) -->
-<!-- # plot(model3, col="green") -->
-<!-- # title("Model 3", line = -0.75, outer = TRUE) -->
+# Model 3
+# par(mfrow = c(2, 2))
+# plot(model3, col="green")
+# title("Model 3", line = -0.75, outer = TRUE)
 
-<!-- ``` -->
+```
 

### 2021-08-06 notebooks/Draft_Report.Rmd

@@
 
----
-
 #### Operationalization of additional variables
-To enrich our model and limit OVB, we brainstormed a number of additional variables that could affect travel rates during COVID.
-For one, we looked at how COVID affected the county, $$through the unemployment rate$$.
+To enrich our model and limit OVB, we brainstormed a number of additional variables that could affect travel rates during COVID-19.
+For one, we looked at how COVID-19 affected the county, $$through the unemployment rate$$.
 We also looked at age, gender, income demographics.
-Then, in addition to income demographics, we focused on percent of younger folks, as defined as those younger than 17 and older folks as defined as those older than 65 because of the nature of how COVID affects an individual.
+Then, in addition to income demographics, we focused on percent of younger folks, as defined as those younger than 17 and older folks as defined as those older than 65 because of the nature of how COVID-19 affects an individual.
 
@@
 
-\clearpage
-
 # Linear Regression Results
@@
 * In `Model 3` with age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17.
+
 * In terms of $R^2$ and RSE, we are improving with each model.
 * The addition of our chosen parameters for each following model is also statistically better than the previous, as indicated by the F statistic.
+
+$$RECHECK THIS SECTION FOR VALUES$$
+```{r calc. coeff percents}
+#Model 1
+#when you compare a coeff for non-log x to a log y:
+(1-exp(coef(model1)["is_rural"]))*100
+#For every percent increase in x 'is rural', you get a #% outcome in number of long trips.
+
+#Model 2
+(1-exp(coef(model2)["is_rural"]))*100
+#Model 3
+(1-exp(coef(model3)["is_rural"]))*100
+
+#Same for income
+
+#Model 2
+(1-exp(coef(model2)["log(median_income + 1)"]))*100
+#Model 3
+(1-exp(coef(model3)["log(median_income + 1)"]))*100
+
+#Same for density
+
+#Model 2
+(1-exp(coef(model2)["log(density)"]))*100
+#Model 3
+(1-exp(coef(model3)["log(density)"]))*100
+
+#Same for age 17 & 65
+#Model 3
+(1-exp(coef(model3)["log(lt17)"]))*100
+(1-exp(coef(model3)["log(gt65)"]))*100
+```
 

### 2021-08-06 src/data/DataCleaning.Rmd

@@
 d <- merge(x = age_f, y = d, by = c("county"))
+d <- d %>% rowwise() %>%
+  mutate(
+    lt17 = as.integer(0.01 * pct_lt17 * total_pop),
+    gt65 = as.integer(0.01 * pct_gt65 * total_pop)
+  )
 d <- d %>% drop_na()
@@
     density,
-    pct_lt17,
-    pct_gt65,
+    lt17,
+    gt65,
     in_poverty

### 2021-08-06 notebooks/Draft_Report.Rmd
final report

--- 2021-08-05 notebooks/Draft_Report.Rmd
@@
 ---
-title: "Section_8_Team1_Lab2_Report"
+title: "Causal Relationship Study: COVID-19 Caseload and June 2021 Vacation Travel in Texas"
 author: "Jonas Degnan, Autumn Rains, Lucy Wu"
@@
 output:
-  pdf_document: 
-    toc: yes
+  bookdown::pdf_document2: 
+    toc: true
+    toc_depth: 1
     fig_caption: yes
-    fig_crop: no
+    extra_dependencies: ["flafter"]
   word_document: default
@@
     theme: readable
+always_allow_html: true
 ---
@@
 # install.packages("tinytex")
+# webshot::install_phantomjs()
 ```
@@
 library(stargazer)
+library(plotly)
+library(bookdown)
 theme_set(theme_minimal())
@@
 # tinytex::install_tinytex()
-#knitr::knit(input = "./../notebooks/Draft_Report.Rmd", output = "./../reports/final_report.pdf")
 ```
@@
 Our hypothesis as to why this decline has occurred is due directly to the volume of COVID-19 number of cases.
-As virus infection oscillate regionally in the United States, our team of data scientists will look to answer the following primary research question using linear regression statistical techniques:
+As virus infection rates oscillate regionally in the United States, our team of data scientists will look to answer the following primary research question using linear regression statistical techniques:
 
-Do COVID-19 cases impact vacation travel trips for Texas for June, 2021?
+Do COVID-19 cases impact vacation travel trips for Texas for June 2021?
 
@@
 
-Does the median income, the population density, and/or the age by county type (rural or urban) impact vacation travel for Texas residents in June, 2021?
+Does the median income, the population density, and/or the age by county type (rural or urban) impact vacation travel for Texas residents in June 2021?
 
@@
 
----
-
-#### Operationalization of additional variables
-To enrich our model and limit OVB, we brainstormed a number of additional variables that could affect travel rates during COVID.
-For one, we looked at how COVID affected the county, $$through the unemployment rate$$.
-We also looked at age, gender, income demographics.
-Then, in addition to income demographics, we focused on percent of younger folks, as defined as those younger than 17 and older folks as defined as those older than 65 because of the nature of how COVID affects an individual.
-
-As indicated, each source will provide information on county population demographics such as income, rural vs.
-urban classification, etc.
-Together, these sources were filtered for the state of Texas and combined to create a dataframe for analysis.
----
-
 ```{r dataframe, echo=FALSE}
@@
 # head(df)
+```
+
+```{r operationalization-table, fig.cap="Operationalization Summary for Texas in June 2021", fig.align='center', echo=FALSE}
+knitr::include_graphics(path="./vartable.png",dpi=160)
 ```
@@
 
-```{r var-description}
-
-table(
-  source = c(1, 2, 3),
-  description = c("a", "b", "c"),
-  variable = c("A", "B", "C")
-)
+```{r var-description,echo=FALSE}
+fig <- plot_ly(
+  type = 'table',
+  header = list(
+    values = c("<br>Source</br>", "Description</br>", "<br>Variable Name</br>"),
+    line = list(color = 'black'),
+    # fill = list(color = '#119DFF'),
+    align = c('left','center'),
+    font = list(color = 'black', size = 10)
+  ),
+  cells = list(
+    values = rbind(
+      c("<br><i>Trips by Distance</i></br>Bureau of Transportation Statistics:</br>U.S. Dept. of Transportation</br>",
+        "<br><i>Coronavirus (COVID-19) Data<br>in the United States</i></br>The New York Times",
+        "<br><i>2010 Decennial Census</i></br>U.S. Census Bureau",
+        "<br><i>County Information Program</i></br>Texas Association of Counties"),
+      c("<br>Number of trips taken</br> (>100 miles)", "<br>Total cases</br>", "<br>Urban/rural county indicator</br>","<br>Count median income<br>County population density<br>Population <17</br>Population >65</br>"),
+      c("<br>long_trip</br>", "<br>cases</br>","<br>is_rural</br>","<br>median_income</br>density</br>pct_lt17</br>pct_gt65</br>")),
+    line = list(color = 'black'),
+    # fill = list(color = c('#25FEFD', 'white')),
+    align = c('left', 'center'),
+    font = list(color = c('black'), size = 10)
+    ))
 ```
 
-```{r var-summary, echo=FALSE, message=FALSE, fig.cap='Texas County Variable Characteristics', fig.pos='!b'}
+```{r var-summary, echo=FALSE, message=FALSE, fig.align='center'}
 df$rural_cat <- ifelse(df$is_rural == 1, "Rural", "Urban")
@@
 table1::label(df$gt65) <- "Age > 65 Years"
-table1::table1(~long_trip + density + cases + median_income + lt17 + gt65 | rural_cat, data = df, topclass="Rtable1-zebra", overall = "Total")
+table1::table1(~long_trip + density + cases + median_income + lt17 + gt65 | rural_cat, data = df, topclass="Rtable1-zebra", 
+               overall = "Total",
+               caption="Variable Summary Statistics")
 ```
@@
 Below are histograms of variables that did not follow normal distribution and required transformation.
-We applied a logarithmic transform to each of these variables, which can be seen in the collective Figure \@ref(tab:var-plots).
+We applied a logarithmic transform to each of these variables, which can be seen in the collective Figure \@ref(fig:var-plots).
 
-```{r var-plots, fig.align='center', message=FALSE, echo=FALSE}
+```{r var-plots, fig.cap= 'Distribution of raw variable data', fig.align='center', message=FALSE, echo=FALSE}
 attach(df)
@@
 
-```{r ggally, echo=FALSE, message=FALSE, fig.cap='Variable Characterstics', fig.pos='!b'}
+```{r ggally, echo=FALSE, message=FALSE, fig.cap='Variable Characterstics', fig.align='center'}
 County = as.factor(df$rural_cat)
@@
 To demonstrate the operationalization of the data, Table \@ref(tab:accounting-table) below display the original value counts from the dataset as well as how many values were removed as a part of filtering the dataset to remove all states except Texas.
-For the analysis, 230 responses were analyzed as shown in Table \@ref(tab:accounting-table).
+For the analysis, 230 observations were analyzed.
 
-```{r accounting-table, echo=FALSE, fig.cap='Variable Characterstics', fig.pos='!b'}
+```{r accounting-table, echo=FALSE, fig.align='center'}
 df2 <- df
@@
 colnames(accounting_df) <- c('Step','Number of Samples','Samples Removed', 'Reason')
-knitr::kable(accounting_df, align = "c")
+knitr::kable(accounting_df,
+             caption = "Accounting Table",
+             format = "simple")
 ```
@@
 
-```{r model creation, echo=FALSE, warning=FALSE}
+```{r model creation, echo=FALSE, warning=FALSE, fig.align='center', fig.cap='Stargazer'}
 model1 <- lm(log(long_trip) ~ log(cases+1) + is_rural, data = df) #limited model
@@
 
-First, we consider the models holistically. Here are some of our observations:
-* In `Model 1`, all of our variables are extremely significant, though in the opposite direction as expected.
-We would expect an inverse relationship between COVID and travel, though it seems that this model is producing a positive relationship.
-* In `Model 2`, when county density is factored in, the presence of COVID seems to have less positive of an effect on travel and that higher density counties see more travel.
-* In `Model 3` with age factored in, we see that actually the most statistically significant effect on long trips are median income of the household and presence of kids under 17.
-* In terms of $R^2$ and RSE, we are improving with each model.
-* The addition of our chosen parameters for each following model is also statistically better than the previous, as indicated by the F statistic.
+In `Model 1`, all of our variables are statistically significant, though the `cases` variables is in the opposite direction as expected: our primary hypothesis expected an inverse relationship between COVID-19 case counts.
+The `is_rural` variable coefficient directionality aligns our primary hypothesis that living in a rural county increases frequency of vacation trips.
 
-We will also interpret the coefficient values and what these mean in the context of our model.
-Almost all of our variables needed a log transform, including the output, making this a log-log regression.
-These regressions are most intuitive to understand in terms of percent increases.
-* First to note is our persistently significant constant in each model.
-Regardless of any variables, we are getting between ~300 (as with model3) and over 8000 (as with model2) trips over 100 miles.
-This is calculated as $e^(Constant)$.
-* In `Model 1`, for every percent increase in cases, we are getting almost half a percent increase in trips. The positive relationship doesnt make much sense, so we will look to the other models.
-* In `Model 2`, we see that density has a more positive relationship with trips, with every percent increase in county density modeled to a half percent increase in trips.
-Also the positive relationship between COVID cases and trips has decreased to resulting in 0.15 and 0.25% increase in trips respectively.
-* In `Model 3`, COVID cases become insignificant.
-The coefficients have become slightly negative as in increase in cases decreases trips, which is in line with our original hypothesis, but is not significant.
-Instead, for every percent median income of the county increases by, the amount of long trips decrease by 0.3%.
-This could indicate that more affluent counties trust science and err on the safe side in the midst of COVID.
-Finally, in our 3rd model, counties with more people under 17 years of age are more likely to take long trips.
-Due to the time frame of our analysis, this also makes sense because june is the start of summer break when many families like to take vacations together.
+In `Model 2`, when county density is factored in, the COVID-19 case counts have a direct relationship on vacation trip frequency.
+The relationship between median income and vacation trip frequency is not statistically significant.
+
+In `Model 3` with age factored in, the statistically significant relationships profile changes. Only age demographics and income now have significance on vacation trip frequency.
+
+Each successive models improves $R^2$ and residual standard errors, indicating that variance estimates improve with each model resulting in improve model performance.
+
+```{r calc. coeff percents, echo=FALSE}
+#Model 1
+#when you compare a coeff for non-log x to a log y:
+m1_is_rural = (1-exp(coef(model1)["is_rural"]))*100
+#For every percent increase in x 'is rural', you get a #% outcome in number of long trips.
+
+#Model 2
+m2_is_rural = (1-exp(coef(model2)["is_rural"]))*100
+
+#Model 3
+m3_is_rural = (1-exp(coef(model3)["is_rural"]))*100
+
+#Same for income
+
+#Model 2
+m2_income = round((coef(model2)["log(median_income + 1)"])/100,digits=4)
+#Model 3
+m3_income = round((coef(model3)["log(median_income + 1)"])/100,digits=4)
+
+#Same for density
+
+#Model 2
+m2_density = round((coef(model2)["log(density + 1)"])/100,digits=4)
+#Model 3
+m3_density = round((coef(model3)["log(density + 1)"])/100,digits=4)
+
+#Same for age 17 & 65
+#Model 3
+m3_lt17 = (1-exp(coef(model3)["log(lt17)"]))*100
+m3_gt65= (1-exp(coef(model3)["log(gt65)"]))*100
+```
+
+## Coefficient Interpretation
+
+In `Model 1`, for every percent increase in cases, there is a $0.0055 \%$ increase in trip frequency and being in a rural county increases trip frequency by $62.2 \%$.
+
+In `Model 2`, for every percent increase in county density, there is a $0.0062\%$ increase in trip frequency.
+This provides evidence to support the causal theory that more dense counties would have high rates of vacation travel.
+The strength of the causal relationship between the case count and `is_rural` indicator has decreased with this new covariate.
+
+In `Model 3`, the causal relationships between COVID-19 cases and county density are no longer statistically significant.
+The prevelance of COVID-19 cases no longer has a causal relationship with trips.
+This can be interpreted as Texans are experiencing pandemic fatigue and may no longer be influenced by health data when making travel plans.
+A county being rural or dense no longer a causal relationships with vacation trip counts and can be interpreted as age demographics have a accurately explain vacation trip counts.
+
+Median income's relationship is now statistically significant, as well as both age quantiles.
+Median income has a negative causal relationship ($0.003\%$) with vacation counts.
+This evidence does not support our hypothesis that higher median incomes result in an increase in the ability to take vacation trips.
+This is explored in our Classical Linear Model (CLM) assumptions related to collinearity.
+The population under age 17 now has the strongest causal relationship ($0.008\%$) with vacation trip counts and due to the time frame of our analysis, this makes sense because June is the start of summer break when many families take vacations together.
+The population over age 65 is typically retired and more available to take vacations and has a positive ($0.001\%$) causal relationship.
 
@@
 
-Below is an analysis of the Classin Linear Model assumptions for $Model 2$ and $Model 3$:
+Below is an analysis of the Classin Linear Model assumptions for `Model 2` and `Model 3`:
 
@@
 The results for the VIF analysis are seen in the table below.
-Running a VIF in R for $Model 2$ and $Model 3$, we see the VIF scores are relatively low (below 5) for the predictor variables in $Model 2$.
-However, upon review of $Model 3$, this is not the case.
+Running a VIF in R for `Model 2` and `Model 3`, we see the VIF scores are relatively low (below 5) for the predictor variables in `Model 2`.
+However, upon review of `Model 3`, this is not the case.
 With VIF factors greater than 5, the both age variables would be candidates for removal outright.
-When reviewing the population density variable in $Model 3$, it remains significant for $Model 2$.
-Additionally, when viewing the variable relationships between the two age variables and population in Figure 5 (GGally), strong linear relationships can be viewed which would also indicate almost near perfect colinearity and thus should be removed from the model.
+When reviewing the population density variable in `Model 3`, it remains significant for `Model 2`.
+Additionally, when viewing the variable relationships between the two age variables and population in Figure \@ref(fig:ggally), strong linear relationships can be viewed which would also indicate almost near perfect colinearity and thus should be removed from the model.
 
-```{r VIF & correlation, echo=FALSE}
+```{r model-2-VIF, echo=FALSE, fig.align='center'}
+knitr::kable(as.table(car::vif(model2)), 
+             digits = 2, 
+             caption = "Model 2 Covariate Variance Inflation Factors", 
+             col.names = c("Variable", "VIF"),
+             format = "simple")
+```
 
-#create vector of VIF values
-vif_values2 <- car::vif(model2)
-vif_values2
-vif_values3 <- car::vif(model3)
-vif_values3
+```{r model-3-VIF, echo=FALSE, fig.align='center'}
+knitr::kable(as.table(car::vif(model3)), 
+             digits = 2, caption = "Model 3 Covariate Variance Inflation Factors", 
+             col.names = c("Variable", "VIF"),
+             format="simple")
 
-#Put values into nicer format table
 ```
@@
 **Linear Conditional Expectation:** For this assumption, the residuals for linear relationships must be evaluated.
-By plotting the residuals, as seen below in the first scatter plot for both $Model 2$ and $Model 3$ ('Residual vs Fitted').
+By plotting the residuals, as seen below in the first scatter plot for both `Model 2` and `Model 3` ('Residual vs Fitted').
 We generally see a linear line as we move across predictor values with the residuals with no strong pattern.
@@
 By plotting fitted values vs.
-the square root of the standardized residuals for the $Model 2$ and model 3 ('Scale-Location' figure), there is an even or equal spread dispersion across the x values.
-Therefore $Model 2$ and $Model 3$ both meet this condition when including the transformed variables.
+the square root of the standardized residuals for the `Model 2` and `Model 3` ('Scale-Location' figure), there is an even or equal spread dispersion across the x values.
+Therefore `Model 2` and `Model 3` both meet this condition when including the transformed variables.
 
 5.
-**Normally Distributed Errors:** When investigating a normal distribution of the residuals, the QQ plot is helpful to investigate this assumption.
-For $Model 2$ and $Model 3$, a nice linear line can be seen between the theoretical quantities and the standardized residuals ('Normal Q-Q').
-In conclusion, we meet this assumption with both $Model 2$ and $Model 3$.
+**Normally Distributed Errors:** When investigating a normal distribution of the residuals, the Q-Q plot is helpful to investigate this assumption.
+For `Model 2` and `Model 3`, a linear relationship between theoretical quantities and the standardized residuals ('Normal Q-Q') can be observed.
+In conclusion, we meet this assumption with both `Model 2` and `Model 3`.
 
-```{r CLM assumptions check, echo=FALSE}
+```{r CLM-assumptions-check, echo=FALSE, fig.cap="Model 2 Q-Q", fig.align='center'}
 
@@
 plot(model2, col="blue")
-title("Model 2", line = -0.75, outer = TRUE)
-
-# Model 3
-# par(mfrow = c(2, 2))
-# plot(model3, col="green")
-# title("Model 3", line = -0.75, outer = TRUE)
+# title("Model 2 Q-Q", line = -0.75, outer = TRUE)
 
@@
 
-<!-- # Omitted Variable -->
+# Omitted Variable
 
-<!-- ### Vehicle Ownership -->
+## Vehicle Ownership
 
-<!-- One example of an omitted variable that may cause bias in our model is that of vehicle ownership among the population.
+One example of an omitted variable that may cause bias in our model is that of vehicle ownership among the population.
 It can certainly be expected that individuals that own a vehicle would be able to more easily travel short or long distances more frequently with relative ease.
@@
 From the data available, there would no immediate proxy for this omitted variable.
--->
 
-<!-- ### Fuel Prices -->
+## Fuel Prices
 
-<!-- Another example of an omitted variable that may also cause bias in our model is fuel prices.
+Another example of an omitted variable that may also cause bias in our model is fuel prices.
 Whether this is jet fuel or gasoline, fluctuations in this commodity would certainly impact travel costs.
@@
 Conversely, if fuel prices for the time period of the sample for the population were lower, the bias would would be farther from zero.
--->
 
-<!-- # Conclusion -->
+# Conclusion
 
-<!-- The goal of this analysis was to determine if there existed a causal relationship between distance traveled and COVID-19 infection rates for residents of the state of Texas.
-Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model, though with all variables considered, our results do not indicate vacation and COVID cases are necessarily related.
--->
+The goal of this analysis was to determine if there existed a causal relationship between distance traveled and COVID-19 infection rates for residents of the state of Texas.
+Based upon the linear regression analysis above, model 2 would indicate there may be a causal relationship between distance traveled (outcome variable) and the variables found within that model, though with all variables considered, our results do not indicate vacation and COVID-19 cases are necessarily related.
 
-<!-- ### Next Steps -->
+## Next Steps
 
-<!-- For future analysis, it is recommended that COVID-19 policy adherence data per county be incorporated into the model to increase the accuracy.
+For future analysis, it is recommended that COVID-19 policy adherence data per county be incorporated into the model to increase the accuracy.
 While the Governor of Texas issued detailed protocols or mandates for the state of Texas, the degree to which each county (and thus the individual residents) complied would be useful data to have if such data were to exist.
@@
 Last, it is highly recommended that any model generated in the future continue to use the latest data available given the oscillation in vaccination rates, government policy, and new or known variants of the COVID-19 virus.
--->
